### 2020-03-11 data_appendix.Rmd
Adding name

---
title: "Data Appendix to \"Name of your paper\""
author: "Katya Garcia-Israel"
output: 
  pdf_document:
    toc: true
    number_sections: true
---

```{r setup, echo = F, message = F}
knitr::opts_chunk$set(results = 'asis', cache = F)
library(tidyverse)
library(summarytools)
st_options(plain.ascii = F,
           style = "rmarkdown",
           footnote = NA,
           subtitle.emphasis = F,
           dfSummary.silent = T,
           dfSummary.valid.col = F,
           tmp.img.dir = "./tmp",
           dfSummary.style = "grid")

#The following custom function simplifies the process of writing dfSummaries to html files
export_summary_table <- function(dfSummary_output){
  data_info <- attr(dfSummary_output, "data_info")
  ds_name <- data_info$Data.frame
  print(dfSummary_output,
      file = str_c("output/", ds_name, "_summary.html"),
      method = "browser",
      report.title = ds_name)
}
```

```{r set dfSummary css style, echo = F, include = F}
st_css()
```
# Appendix description
*Your Data Appendix should begin with a brief statement explaining its purpose like the following one.*

This Data Appendix documents the data used in "Papter Title". It was prepared in a Rmarkdown document that contains both the documentation and the R code used to prepare the data used in the final estimation. It also includes descriptive statistics for both the original data and the final dataset, with a discussion of any issues of note.

The datasets used directly by the final analysis are saved in `processed-data/` at the end of this file.

*Note: this document structure will require you to re-run steps of your analysis multiple times. If your code takes a long time, please come talk with me about strategies to reduce run time or save earlier results.*

# Instructions for Use
This document includes instructions for how to create your Data Appendix. Outside of this section, instruction paragraphs are listed in *italics* (like the first paragraph above). Instructions should be removed before submission.

To start creating your own data appendix, follow these steps:

1. Replace the title and author in the section at the top of the file (called the YAML).
1. Commit your changes with a message like "customizing data appendix".
1. Delete this instruction section of the document.
1. Remove any other instructions in italics and examples from the completed sections of the document.

Remember that you will submit your assignment by committing and then pushing your versions to your repository. I encourage you to commit your changes often as you work, but there are three specific points at which you need to both submit and push changes, corresponding to course deadlines:

1. You must submit a version with the original data section completed by the Data Appendix 1 deadline. This will include the .Rmd file, the .pdf file, and the html data summary files stored in the output folder.
2. You must submit a version with all parts completed by the Data Appendix 2 deadline.  
3. You must submit a final version of this document that is consistent with your final paper by the final project deadline.

While creating your data appendix, refer regularly to the assignment descriptions posted on Moodle.

A few tips:

- When creating a list like this one, be sure to put an empty line above the list. If you don't do this, your entries won't be formatted a list.
- Make sure you have empty lines above and below section and subsection headings.
- When creating numbered lists, you can number all items in your list with 1. Rmarkdown will number them sequentially when it creates your final document.

# Raw data
*Each dataset you use will have its own documentation section. The next subsection in this document (Dataset description) is a template. You can copy this section and paste it into your document each time you need to add a section for a new dataset. Note that each line in the Dataset description section __must__ end with two spaces.* 
This section documents the datasets used in this analysis.

## Dataset description
**Citation:** Put citation here in APA or other consistent format that you will use throughout the project. Include a hyperlink if applicable.  
**DOI:** If the dataset has a documention identified (DOI) assigned put it here.  
**Date Downloaded:** Identify when you downloaded the dataset.  
**Filename(s):** raw_data/filename.csv *If you have a large number of files you can use a patten (see visit data below)*
**Unit of observation:** What distinguishes different rows in your dataset?  
**Dates covered:** What time frame does the data cover?

### To obtain a copy

Describe in a step-by-step fashion how an interested user could obtain the data.

### Importable version (if necessary)

**Filename(s):** importable-data/filename-importable.csv

In some cases the raw data is not directly importable. In this case, you should fully document every step you took to create the importable data in a subsection like this one.

### Variable descriptions

Create a bullet list with the name of each variable in the dataset followed by any information the user would need to understand it.

- **variable_name:** Variable description. 
- **variable_name2:** Description of second variable.

### Data import code and summary

*Once you've described the variables, enter an R chunk by selecting Code -> Insert Chunk, or Ctrl+Alt+I, give it a name to describe the dataset you are importing. After importing, export a dataframe summary using the command.*

`export_summary_table(dfSummary(dataset_name))`

*While it will make your resulting file long, you should not modify the chunk options to suppress printing of code and output. I would likely not include this in the documentation for an actual paper I was submitting, but including them here will let me read your code and the output message from R and may help identify data import concerns early in the process. Since these files will exist only electronically, their length is less of a concern. If you like to print out files to proofread and want me to help you shorten the printed versions, let me know. We can temporarily modify the chunk options for printing and restore them before you submit the assignment.*

# Data Processing and Combination
*This section should include a discussion of the processing and merging steps needed to create your basic data. The code to implement these steps should be included in chunks in this section. Once the final merged data has been created, you should use the dfSummary function again to summarize the data you will be using. You should also save a file containing all the objects you will use in your final analysis to the processed_data folder.*

# Analysis Variables

This section should include a description of all the variables that are used in your final analysis. At the end of the section, you should save all of these variables in the processed_data folder of your repository.

# Discussion of Data

*This section should include a discussion of any data patterns you notice based on the summaries created in the code above.*

### 2020-03-16 data_appendix.Rmd
Testing a change.
added raw data
Commit the changes

@@
 ---
-title: "Data Appendix to \"Name of your paper\""
+title: "Data Appendix to \"The impact of weather events on schooling outcomes\""
 author: "Katya Garcia-Israel"
@@
            dfSummary.style = "grid")
+# input Stata file
+library(foreign)
+data_2005 <- read.dta("C:/Users/katya/Documents/MyDocuments/A Smith College Spring Semester/Natural Resource Economics/seminar-paper-katyagi21/Raw data/seminar_paper_data/2005_hhsurvey.dta")
 
@@
 ## Dataset description
-**Citation:** Put citation here in APA or other consistent format that you will use throughout the project. Include a hyperlink if applicable.  
-**DOI:** If the dataset has a documention identified (DOI) assigned put it here.  
-**Date Downloaded:** Identify when you downloaded the dataset.  
+**Citation:** .  
+Instituto Nacional de Estadistica. (2005). Encuesta de Hogares (2005) [SAV file]. Retrieved from https://www.ine.gob.bo/index.php/banco/base-de-datos-sociales
+
+**Date Downloaded:** Downloaded September 2019.
 **Filename(s):** raw_data/filename.csv *If you have a large number of files you can use a patten (see visit data below)*
-**Unit of observation:** What distinguishes different rows in your dataset?  
-**Dates covered:** What time frame does the data cover?
+**Unit of observation:** Unit of observation at the individual level, but each individual is identified mainly by the household id. 
+**Dates covered:** 2005
 
@@
 
-Describe in a step-by-step fashion how an interested user could obtain the data.
+This household survey can be found on the Bolivian National Statistical Institute website, under the menu for 'Banco de Datos' and 'Base de datos de encuestas sociales.' Then the household surveys may be found for the years 2005-2018, excepting 2010.
 
@@
 
-In some cases the raw data is not directly importable. In this case, you should fully document every step you took to create the importable data in a subsection like this one.
+The original format of this data was from the sav file, but I imported it into Stata before importing the Stata format into R.
 
@@
 
-- **variable_name:** Variable description. 
+- **hh_id:** This variable is the household ID, which is the same for each member of the household. It is made up of letters and numbers. 
 - **variable_name2:** Description of second variable.

### 2020-03-19 data_appendix.Rmd
Commit data inputs and renaming changes

@@
            dfSummary.style = "grid")
-# input Stata file
+# input Stata file, add haven package to make stata 16 version readable 
 library(foreign)
-data_2005 <- read.dta("C:/Users/katya/Documents/MyDocuments/A Smith College Spring Semester/Natural Resource Economics/seminar-paper-katyagi21/Raw data/seminar_paper_data/2005_hhsurvey.dta")
+library(haven)
+data_2005 <- read_dta("C:\\Users\\katya\\Documents\\MyDocuments\\A Smith College Spring Semester\\Natural Resource Economics\\seminar-paper-katyagi21\\Raw data\\seminar_paper_data\\2005_hhsurvey.dta")
+#install.packages("reshape")
+require(reshape)
+data_2005 <- rename(data_2005, c(folio = "hh_id"))
+data_2005 <- rename(data_2005, c(s3_08 = "assist"))
+data_2005 <- rename(data_2005, c(s3_09 = "reason"))
+data_2005 <- rename(data_2005, c(nm_y = "members"))
+data_2005 <- rename(data_2005, c(s1_02 = "sex"))
+data_2005 <- rename(data_2005, c(s1_03 = "age"))
+data_2005 <- rename(data_2005, c(s1_05 = "relation"))
+data_2005 <- rename(data_2005, c(s3_01 = "literacy"))
+data_2005 <- rename(data_2005, c(s3_04 = "register"))
+data_2005 <- rename(data_2005, c(s3_05b = "grade"))
+data_2005 <- rename(data_2005, c(s3_07 = "repeated"))
 
@@
 
+```{r weather data set imported, echo = F, message = F}
+#Weather data set
+
+weather <- read_csv("C:\\Users\\katya\\Documents\\MyDocuments\\A Smith College Spring Semester\\Natural Resource Economics\\seminar-paper-katyagi21\\Raw data\\recurrencia_inundacion2002_2012bol.csv")
+
+#rename variables
+weather <- rename(weather, c(NOM_DEP = "dept"))
+weather <- rename(weather, c(A_2005 = "flood"))
+
+library(dplyr)
+weather %>% 
+  group_by(dept) %>% 
+  summarise(flood = sum(flood))
+
+```
+
 ```{r set dfSummary css style, echo = F, include = F}
@@
 
-This Data Appendix documents the data used in "Papter Title".
+This Data Appendix documents the data used in "Paper Title".
 It was prepared in a Rmarkdown document that contains both the documentation and the R code used to prepare the data used in the final estimation.
@@
 
+```{r renaming variables}
+
+```
+
 ### Variable descriptions
@@
 - **hh_id:** This variable is the household ID, which is the same for each member of the household. It is made up of letters and numbers. 
-- **variable_name2:** Description of second variable.
+- **nro1:** Description of second variable.
+- **members:** 
+- **urb_rur:** 
+- **urb_rur:** 
+- **urb_rur:**
 
@@
 After importing, export a dataframe summary using the command.*
+```{r selecting variables}
+# select variables 
+hh_vars <- c("hh_id", "assist", "reason", "members", "sex", "age", "relation", "literacy", "register", "grade", "repeated", "dept")
+hh_2005 <- data_2005[hh_vars]
+
+data_2005 <- rename(data_2005, c(folio = "hh_id"))
+data_2005 <- rename(data_2005, c(s3_08 = "assist"))
+data_2005 <- rename(data_2005, c(s3_09 = "reason"))
+data_2005 <- rename(data_2005, c(nm_y = "members"))
+data_2005 <- rename(data_2005, c(s1_02 = "sex"))
+data_2005 <- rename(data_2005, c(s1_03 = "age"))
+data_2005 <- rename(data_2005, c(s1_05 = "relation"))
+data_2005 <- rename(data_2005, c(s3_01 = "literacy"))
+data_2005 <- rename(data_2005, c(s3_04 = "register"))
+data_2005 <- rename(data_2005, c(s3_05b = "grade"))
+data_2005 <- rename(data_2005, c(s3_07 = "repeated"))
+data_2005 <- rename(data_2005, c(depto = "dept"))
+
+#select weather variables
+weather_vars <- c("dept", "flood")
+weather_2005 <- weather[weather_vars]
+
+```
+
+```{r merging datasets}
+total_2005 <- merge(hh_2005, weather_2005,by="dept")
+```
+
+```{r exporting data table}
+export_summary_table(dfSummary(hh_2005))
+```
 

### 2020-03-19 data_appendix.Rmd
*Reorganized code to match intended format. *Modified the file names to use relative paths so they will work on other machines *Switched the rename syntax to R style which can rename a whole long list of variables at once, doesn't use quotes and follows a new_name = old_name pattern. *Added notes which can be deleted later describing steps that need to be added.

--- 2020-03-16 data_appendix.Rmd
@@
            dfSummary.style = "grid")
-# input Stata file
+# add haven package to make stata 16 version readable 
 library(foreign)
-data_2005 <- read.dta("C:/Users/katya/Documents/MyDocuments/A Smith College Spring Semester/Natural Resource Economics/seminar-paper-katyagi21/Raw data/seminar_paper_data/2005_hhsurvey.dta")
+library(haven)
 
@@
 
-## Dataset description
+## Household Survey data
 **Citation:** .  
@@
 - **hh_id:** This variable is the household ID, which is the same for each member of the household. It is made up of letters and numbers. 
+- **nro1:** Description of second variable.
+- **members:** 
+- **urb_rur:** 
+- **urb_rur:** 
+- **urb_rur:**
+
+### Data import code and summary
+
+```{r import household data}
+data_2005 <- read_dta(file.path("Raw data","seminar_paper_data","2005_hhsurvey.dta")) %>% 
+  rename(hh_id = folio,
+        assist = s3_08,  
+        reason = s3_09, 
+        members = nm_y,
+        sex = s1_02,
+        age = s1_03,
+        relation = s1_05,
+        literacy = s3_01,
+        register = s3_04,
+        grade = s3_05b,
+        repeated = s3_07)
+# it looks like you need to add dept = xxx to the list above since dept is not showing up as a variable for me        
+hh_2005 <- data_2005 %>% 
+  select(hh_id, assist, reason, members, sex, age, relation, literacy, register, grade, repeated, dept)
+```
+
+```{r exporting data table}
+export_summary_table(dfSummary(hh_2005))
+```
+
+## Weather data
+**Citation:** Put citation here in APA or other consistent format that you will use throughout the project. Include a hyperlink if applicable.  
+**DOI:** If the dataset has a documention identified (DOI) assigned put it here.  
+**Date Downloaded:** Identify when you downloaded the dataset.  
+**Filename(s):** raw_data/filename.csv *If you have a large number of files you can use a patten (see visit data below)*
+**Unit of observation:** What distinguishes different rows in your dataset?  
+**Dates covered:** What time frame does the data cover?
+
+### To obtain a copy
+
+Describe in a step-by-step fashion how an interested user could obtain the data.
+
+### Importable version (if necessary)
+
+**Filename(s):** importable-data/filename-importable.csv
+
+In some cases the raw data is not directly importable. In this case, you should fully document every step you took to create the importable data in a subsection like this one.
+
+### Variable descriptions
+
+Create a bullet list with the name of each variable in the dataset followed by any information the user would need to understand it.
+
+- **variable_name:** Variable description. 
 - **variable_name2:** Description of second variable.
@@
 
-*Once you've described the variables, enter an R chunk by selecting Code -> Insert Chunk, or Ctrl+Alt+I, give it a name to describe the dataset you are importing.
-After importing, export a dataframe summary using the command.*
+```{r import weather data}
+weather <- read_csv(file.path("Raw data", "recurrencia_inundacion2002_2012bol.csv")) %>% 
+  rename(dept = NOM_DEP, flood = A_2005)
+
+##Note: I can't tell completely but it appears that the file you're reading in has all the weather data.
+Once you load the data onto GitHub, I can help you figure out how to use pivot_longer() to reshape the weather data.
+
+flood_summary %>% weather %>% 
+  group_by(dept) %>% 
+  summarise(flood = sum(flood))
 
 `export_summary_table(dfSummary(dataset_name))`
-
-*While it will make your resulting file long, you should not modify the chunk options to suppress printing of code and output.
-I would likely not include this in the documentation for an actual paper I was submitting, but including them here will let me read your code and the output message from R and may help identify data import concerns early in the process.
-Since these files will exist only electronically, their length is less of a concern.
-If you like to print out files to proofread and want me to help you shorten the printed versions, let me know.
-We can temporarily modify the chunk options for printing and restore them before you submit the assignment.*
+```
 
@@
 You should also save a file containing all the objects you will use in your final analysis to the processed_data folder.*
+
+```{r merging datasets}
+#I would recommend combining all the years of household data and then merging with the weather data
+
+total_2005 <- hh_2005 %>% 
+  left_join(flood_summary)
+```
+
+```
 

### 2020-03-19 data_appendix.Rmd
updated text

--- 2020-03-16 data_appendix.Rmd
@@
            dfSummary.style = "grid")
-# input Stata file
-library(foreign)
-data_2005 <- read.dta("C:/Users/katya/Documents/MyDocuments/A Smith College Spring Semester/Natural Resource Economics/seminar-paper-katyagi21/Raw data/seminar_paper_data/2005_hhsurvey.dta")
 
 #The following custom function simplifies the process of writing dfSummaries to html files
+
 export_summary_table <- function(dfSummary_output){
-  data_info <- attr(dfSummary_output, "data_info")
-  ds_name <- data_info$Data.frame
+  data_households <- attr(dfSummary_output, "data_info")
+  total_2005 <- data_households$hh_2005
   print(dfSummary_output,
-      file = str_c("output/", ds_name, "_summary.html"),
+      file = str_c("C:\\Users\\katya\\Documents\\MyDocuments\\A Smith College Spring Semester\\Natural Resource Economics\\seminar-paper-katyagi21\\", total_2005, "_summary.html"),
       method = "browser",
-      report.title = ds_name)
+      Weather.and.school = total_2005)
 }
+```
+
+# Raw data
+This section documents the datasets used in this analysis: household data, flood data, and drought data from 2005 in Bolivia.
+
+```{r input households data, echo = F, message = F}
+# input Stata file, add haven package to make stata 16 version readable 
+library(foreign)
+library(haven)
+data_2005 <- read_dta("C:\\Users\\katya\\Documents\\MyDocuments\\A Smith College Spring Semester\\Natural Resource Economics\\seminar-paper-katyagi21\\Raw data\\seminar_paper_data\\2005_hhsurvey.dta")
+#install.packages("reshape")
+require(reshape)
+data_2005 <- rename(data_2005, c(folio = "hh_id"))
+data_2005 <- rename(data_2005, c(s3_08 = "assist"))
+data_2005 <- rename(data_2005, c(s3_09 = "reason"))
+data_2005 <- rename(data_2005, c(nm_y = "members"))
+data_2005 <- rename(data_2005, c(s1_02 = "sex"))
+data_2005 <- rename(data_2005, c(s1_03 = "age"))
+data_2005 <- rename(data_2005, c(s1_05 = "relation"))
+data_2005 <- rename(data_2005, c(s3_01 = "literacy"))
+data_2005 <- rename(data_2005, c(s3_04 = "register"))
+data_2005 <- rename(data_2005, c(s3_05b = "grade"))
+data_2005 <- rename(data_2005, c(s3_07 = "repeated"))
+data_2005 <- rename(data_2005, c(depto = "dept"))
+```
+
+```{r floods data set imported, echo = F, message = F}
+#floods data set
+
+floods <- read_csv("C:\\Users\\katya\\Documents\\MyDocuments\\A Smith College Spring Semester\\Natural Resource Economics\\seminar-paper-katyagi21\\Raw data\\recurrencia_inundacion2002_2012bol.csv")
+
+#rename variables
+floods <- rename(floods, c(NOM_DEP = "dept"))
+floods <- rename(floods, c(A_2005 = "flood"))
+
+library(dplyr)
+floods %>% 
+  group_by(dept) %>% 
+  summarise(flood = sum(flood))
+
+```
+
+```{r importing droughts data, echo = F, message = F}
+#drought data set 
+drought <- read_csv("C:\\Users\\katya\\Documents\\MyDocuments\\A Smith College Spring Semester\\Natural Resource Economics\\seminar-paper-katyagi21\\Raw data\\recurrenciasequia_2002_2012bol.csv")
+
+#rename variables
+drought <- rename(weather, c(NOM_DEP = "dept"))
+drought <- rename(weather, c(A_2005 = "drought"))
+
+library(dplyr)
+drought %>% 
+  group_by(dept) %>% 
+  summarise(drought = sum(drought))
+
 ```
@@
 
-*Note: this document structure will require you to re-run steps of your analysis multiple times.
-If your code takes a long time, please come talk with me about strategies to reduce run time or save earlier results.*
-
 # Instructions for Use
-This document includes instructions for how to create your Data Appendix.
-Outside of this section, instruction paragraphs are listed in *italics* (like the first paragraph above).
-Instructions should be removed before submission.
 
@@
 
-# Raw data
-*Each dataset you use will have its own documentation section.
-The next subsection in this document (Dataset description) is a template.
-You can copy this section and paste it into your document each time you need to add a section for a new dataset.
-Note that each line in the Dataset description section __must__ end with two spaces.*
-This section documents the datasets used in this analysis.
-
 ## Dataset description
@@
 - **hh_id:** This variable is the household ID, which is the same for each member of the household. It is made up of letters and numbers. 
-- **variable_name2:** Description of second variable.
+- **nro1:** Description of second variable.
+- **assist:** 
+- **urb_rur:** 
+- **reason:** 
+- **members:** 
+- **sex:** 
+- **age:** 
+- **relation:** 
+- **literacy:** 
+- **register:**
+- **grade:** 
+- **repeated:** 
+- **dept:**
 
@@
 After importing, export a dataframe summary using the command.*
+```{r selecting variables}
+# select variables 
+hh_vars <- c("hh_id", "nro1", "urb_rur", "assist", "reason", "members", "sex", "age", "relation", "literacy", "register", "grade", "repeated", "dept")
+hh_2005 <- data_2005[hh_vars]
 
-`export_summary_table(dfSummary(dataset_name))`
+data_2005 <- rename(data_2005, c(folio = "hh_id"))
+data_2005 <- rename(data_2005, c(s3_08 = "assist"))
+data_2005 <- rename(data_2005, c(s3_09 = "reason"))
+data_2005 <- rename(data_2005, c(nm_y = "members"))
+data_2005 <- rename(data_2005, c(s1_02 = "sex"))
+data_2005 <- rename(data_2005, c(s1_03 = "age"))
+data_2005 <- rename(data_2005, c(s1_05 = "relation"))
+data_2005 <- rename(data_2005, c(s3_01 = "literacy"))
+data_2005 <- rename(data_2005, c(s3_04 = "register"))
+data_2005 <- rename(data_2005, c(s3_05b = "grade"))
+data_2005 <- rename(data_2005, c(s3_07 = "repeated"))
+data_2005 <- rename(data_2005, c(depto = "dept"))
 
-*While it will make your resulting file long, you should not modify the chunk options to suppress printing of code and output.
-I would likely not include this in the documentation for an actual paper I was submitting, but including them here will let me read your code and the output message from R and may help identify data import concerns early in the process.
-Since these files will exist only electronically, their length is less of a concern.
-If you like to print out files to proofread and want me to help you shorten the printed versions, let me know.
-We can temporarily modify the chunk options for printing and restore them before you submit the assignment.*
+#select weather variables
+flood_vars <- c("dept", "flood")
+flood_2005 <- floods[flood_vars]
+
+```
+
+```{r merging datasets}
+total_2005 <- merge(hh_2005, weather_2005,by="dept")
+```
+
+```{r exporting data table}
+export_summary_table(dfSummary(hh_2005))
+```
 

### 2020-03-19 data_appendix.Rmd
Commit merging changes

--- 2020-03-19 data_appendix.Rmd
@@
            dfSummary.style = "grid")
+
 # add haven package to make stata 16 version readable 
@@
 - **nro1:** Description of second variable.
+- **assist:** 
+- **urb_rur:** 
+- **reason:** 
 - **members:** 
-- **urb_rur:** 
-- **urb_rur:** 
-- **urb_rur:**
+- **sex:** 
+- **age:** 
+- **relation:** 
+- **literacy:** 
+- **register:**
+- **grade:** 
+- **repeated:** 
+- **dept:**
 
@@
         grade = s3_05b,
-        repeated = s3_07)
-# it looks like you need to add dept = xxx to the list above since dept is not showing up as a variable for me        
+        repeated = s3_07,
+        dept = depto)
 hh_2005 <- data_2005 %>% 
@@
 
-## Weather data
+## Flood data
 **Citation:** Put citation here in APA or other consistent format that you will use throughout the project. Include a hyperlink if applicable.  
@@
 
-```{r import weather data}
-weather <- read_csv(file.path("Raw data", "recurrencia_inundacion2002_2012bol.csv")) %>% 
+## Flood data
+**Citation:** Put citation here in APA or other consistent format that you will use throughout the project. Include a hyperlink if applicable.  
+**DOI:** If the dataset has a documention identified (DOI) assigned put it here.  
+**Date Downloaded:** Identify when you downloaded the dataset.  
+**Filename(s):** raw_data/filename.csv *If you have a large number of files you can use a patten (see visit data below)*
+**Unit of observation:** What distinguishes different rows in your dataset?  
+**Dates covered:** What time frame does the data cover?
+
+### To obtain a copy
+
+Describe in a step-by-step fashion how an interested user could obtain the data.
+
+### Importable version (if necessary)
+
+**Filename(s):** importable-data/filename-importable.csv
+
+In some cases the raw data is not directly importable. In this case, you should fully document every step you took to create the importable data in a subsection like this one.
+
+### Variable descriptions
+
+Create a bullet list with the name of each variable in the dataset followed by any information the user would need to understand it.
+
+- **variable_name:** Variable description. 
+- **variable_name2:** Description of second variable.
+
+### Data import code and summary
+```{r import flood data}
+flood <- read_csv(file.path("Raw data", "recurrencia_inundacion2002_2012bol.csv")) %>% 
   rename(dept = NOM_DEP, flood = A_2005)
@@
 
-`export_summary_table(dfSummary(dataset_name))`
+`export_summary_table(dfSummary(flood_summary))`
+```
+
+## Drough data
+**Citation:** Put citation here in APA or other consistent format that you will use throughout the project. Include a hyperlink if applicable.  
+**DOI:** If the dataset has a documention identified (DOI) assigned put it here.  
+**Date Downloaded:** Identify when you downloaded the dataset.  
+**Filename(s):** raw_data/filename.csv *If you have a large number of files you can use a patten (see visit data below)*
+**Unit of observation:** What distinguishes different rows in your dataset?  
+**Dates covered:** What time frame does the data cover?
+
+### To obtain a copy
+
+Describe in a step-by-step fashion how an interested user could obtain the data.
+
+### Importable version (if necessary)
+
+**Filename(s):** importable-data/filename-importable.csv
+
+In some cases the raw data is not directly importable. In this case, you should fully document every step you took to create the importable data in a subsection like this one.
+
+### Variable descriptions
+
+Create a bullet list with the name of each variable in the dataset followed by any information the user would need to understand it.
+
+- **variable_name:** Variable description. 
+- **variable_name2:** Description of second variable.
+
+```{r importing droughts data, echo = F, message = F}
+#drought data set 
+drought <- read_csv("Raw data\\recurrenciasequia_2002_2012bol.csv")
+# modify this section following the template with the flood data, but wait until we've gotten the pivot_longer() function working so you don't duplicate effort.
+#rename variables
+drought <- rename(weather, c(NOM_DEP = "dept"))
+drought <- rename(weather, c(A_2005 = "drought"))
+
+library(dplyr)
+drought %>% 
+  group_by(dept) %>% 
+  summarise(drought = sum(drought))
+
 ```

### 2020-03-23 data_appendix.Rmd
Commit coding changes

@@
 
-This Data Appendix documents the data used in "Paper Title".
-It was prepared in a Rmarkdown document that contains both the documentation and the R code used to prepare the data used in the final estimation.
-It also includes descriptive statistics for both the original data and the final dataset, with a discussion of any issues of note.
-
 The datasets used directly by the final analysis are saved in `processed-data/` at the end of this file.
 
-# Instructions for Use
-
-To start creating your own data appendix, follow these steps:
-
-1. Replace the title and author in the section at the top of the file (called the YAML).
-1. Commit your changes with a message like "customizing data appendix".
-1. Delete this instruction section of the document.
-1. Remove any other instructions in italics and examples from the completed sections of the document.
-
-Remember that you will submit your assignment by committing and then pushing your versions to your repository.
-I encourage you to commit your changes often as you work, but there are three specific points at which you need to both submit and push changes, corresponding to course deadlines:
-
-1.
-You must submit a version with the original data section completed by the Data Appendix 1 deadline.
-This will include the .Rmd file, the .pdf file, and the html data summary files stored in the output folder.
-2. You must submit a version with all parts completed by the Data Appendix 2 deadline.  
-3. You must submit a final version of this document that is consistent with your final paper by the final project deadline.
-
-While creating your data appendix, refer regularly to the assignment descriptions posted on Moodle.
-
-A few tips:
-
-- When creating a list like this one, be sure to put an empty line above the list. If you don't do this, your entries won't be formatted a list.
-- Make sure you have empty lines above and below section and subsection headings.
-- When creating numbered lists, you can number all items in your list with 1. Rmarkdown will number them sequentially when it creates your final document.
-
 # Raw data
-*Each dataset you use will have its own documentation section.
-The next subsection in this document (Dataset description) is a template.
-You can copy this section and paste it into your document each time you need to add a section for a new dataset.
-Note that each line in the Dataset description section __must__ end with two spaces.*
 This section documents the datasets used in this analysis.
@@
 **Date Downloaded:** Downloaded September 2019.
-**Filename(s):** raw_data/filename.csv *If you have a large number of files you can use a patten (see visit data below)*
+**Filename(s):** raw_data/filename.csv 
 **Unit of observation:** Unit of observation at the individual level, but each individual is identified mainly by the household id. 
@@
 
-### Importable version (if necessary)
-
-**Filename(s):** importable-data/filename-importable.csv
-
 The original format of this data was from the sav file, but I imported it into Stata before importing the Stata format into R.
@@
 
-Create a bullet list with the name of each variable in the dataset followed by any information the user would need to understand it.
-
-- **hh_id:** This variable is the household ID, which is the same for each member of the household. It is made up of letters and numbers. 
-- **nro1:** Description of second variable.
-- **assist:** 
-- **urb_rur:** 
-- **reason:** 
-- **members:** 
-- **sex:** 
-- **age:** 
-- **relation:** 
-- **literacy:** 
-- **register:**
-- **grade:** 
-- **repeated:** 
-- **dept:**
+- **hh_id:** The household ID, which is the same for each member of the household. It is made up of letters and numbers. 
+- **nro1:** The number of the person in each household.
+- **assist:** Binary variable, whether or not the individual assists school 
+- **urb_rur:** Binary variable, whether the household is in an urban or rural area
+- **reason:** the reason why the individual is not assisting school if they are not 
+- **members:** number of household members excluding servants 
+- **sex:** the sex of the individual, Male = 1, Female =2 
+- **age:** How old the individual is, in years 
+- **relation:** The relationship of the individual to the head of household 
+- **literacy:** Binary variable, whether a person can read and write 
+- **register:** Binary variable, whether an individual registered for school or not 
+- **grade:** For what grade is the person registered 
+- **repeated:** whether the person has repeated the grade they are currently registered in 
+- **dept:** the department in which the person lives
 
@@
 hh_2005 <- data_2005 %>% 
-  select(hh_id, assist, reason, members, sex, age, relation, literacy, register, grade, repeated, dept)
+  select(hh_id, assist, reason, members, sex, age, relation, literacy, register, grade, repeated, dept, urb_rur)
 ```
@@
 ## Flood data
-**Citation:** Put citation here in APA or other consistent format that you will use throughout the project. Include a hyperlink if applicable.  
+**Citation:** GeoSinager. (2016). Mapa Recurrencias de Inundaciones Gestiones 2002-2012. [CSV file]. Retrieved from http://geosinager.defensacivil.gob.bo/maps/226
+
+Instituto Nacional de Estadistica. (2005). Encuesta de Hogares (2005) [SAV file]. Retrieved from https://www.ine.gob.bo/index.php/banco/base-de-datos-sociales
+
 **DOI:** If the dataset has a documention identified (DOI) assigned put it here.  
@@
 
-- **variable_name:** Variable description. 
-- **variable_name2:** Description of second variable.
+- **dept:** department name 
+- **A_2005:** how many floods occurred in 2005 in that location 
+- **A_2006:** how many floods occurred in 2006 in that location 
+- **A_2007:** how many floods occurred in 2007 in that location 
+- **A_2008:** how many floods occurred in 2008 in that location 
+- **A_2009:** how many floods occurred in 2009 in that location 
+- **A_2010:** how many floods occurred in 2010 in that location 
+- **A_2011:** how many floods occurred in 2011 in that location 
+- **A_2012:** how many floods occurred in 2012 in that location
 
 ### Data import code and summary
+```{r import flood data}
+flood <- read_csv(file.path("Raw data", "recurrencia_inundacion2002_2012bol.csv")) %>% 
+  rename(dept = NOM_DEP, flood = A_2005)
 
-## Flood data
-**Citation:** Put citation here in APA or other consistent format that you will use throughout the project. Include a hyperlink if applicable.  
-**DOI:** If the dataset has a documention identified (DOI) assigned put it here.  
-**Date Downloaded:** Identify when you downloaded the dataset.  
-**Filename(s):** raw_data/filename.csv *If you have a large number of files you can use a patten (see visit data below)*
-**Unit of observation:** What distinguishes different rows in your dataset?  
-**Dates covered:** What time frame does the data cover?
+flood_summary %>% weather %>% 
+  group_by(dept) %>% 
+  summarise(flood = sum(flood))
+
+`export_summary_table(dfSummary(flood_summary))`
+```
+## Drought data
+**Citation:** GeoSinager. (2016). Mapa Recurrencias de Sequias Gestiones 2002-2012. [CSV file]. Retrieved from http://geosinager.defensacivil.gob.bo/maps/225
+
+**Date Downloaded:** October 2019.  
+**Filename(s):** Raw data/droughts.csv 
+**Unit of observation:** The unit of observation is the municipality 
+**Dates covered:** 2002-2012
 
@@
 
-Describe in a step-by-step fashion how an interested user could obtain the data.
+This data can be obtained from http://geosinager.defensacivil.gob.bo/maps/225. It is shown on a map with different layers.
 
@@
 
-- **variable_name:** Variable description. 
-- **variable_name2:** Description of second variable.
+- **dept:** department name 
+- **A_2005:** how many droughts occurred in 2005 in that location 
+- **A_2006:** how many droughts occurred in 2006 in that location 
+- **A_2007:** how many droughts occurred in 2007 in that location 
+- **A_2008:** how many droughts occurred in 2008 in that location 
+- **A_2009:** how many droughts occurred in 2009 in that location 
+- **A_2010:** how many droughts occurred in 2010 in that location 
+- **A_2011:** how many droughts occurred in 2011 in that location 
+- **A_2012:** how many droughts occurred in 2012 in that location
 
-### Data import code and summary
-```{r import flood data}
-flood <- read_csv(file.path("Raw data", "recurrencia_inundacion2002_2012bol.csv")) %>% 
-  rename(dept = NOM_DEP, flood = A_2005)
-
-##Note: I can't tell completely but it appears that the file you're reading in has all the weather data.
-Once you load the data onto GitHub, I can help you figure out how to use pivot_longer() to reshape the weather data.
-
-flood_summary %>% weather %>% 
+```{r importing droughts data, echo = F, message = F}
+#drought data set 
+library(readxl)
+drought <- read_excel(file.path("Raw data", "recurrenciasequia_2002_2012bol (1).xls")) %>% 
+#rename variables
+    rename(dept = NOM_DEP, drought = A_2005)
+drought_summary %>%
   group_by(dept) %>% 
@@
 
-`export_summary_table(dfSummary(flood_summary))`
-```
-
-## Drough data
-**Citation:** Put citation here in APA or other consistent format that you will use throughout the project. Include a hyperlink if applicable.  
-**DOI:** If the dataset has a documention identified (DOI) assigned put it here.  
-**Date Downloaded:** Identify when you downloaded the dataset.  
-**Filename(s):** raw_data/filename.csv *If you have a large number of files you can use a patten (see visit data below)*
-**Unit of observation:** What distinguishes different rows in your dataset?  
-**Dates covered:** What time frame does the data cover?
-
-### To obtain a copy
-
-Describe in a step-by-step fashion how an interested user could obtain the data.
-
-### Importable version (if necessary)
-
-**Filename(s):** importable-data/filename-importable.csv
-
-In some cases the raw data is not directly importable. In this case, you should fully document every step you took to create the importable data in a subsection like this one.
-
-### Variable descriptions
-
-Create a bullet list with the name of each variable in the dataset followed by any information the user would need to understand it.
-
-- **variable_name:** Variable description. 
-- **variable_name2:** Description of second variable.
-
-```{r importing droughts data, echo = F, message = F}
-#drought data set 
-drought <- read_csv("Raw data\\recurrenciasequia_2002_2012bol.csv")
-# modify this section following the template with the flood data, but wait until we've gotten the pivot_longer() function working so you don't duplicate effort.
-#rename variables
-drought <- rename(weather, c(NOM_DEP = "dept"))
-drought <- rename(weather, c(A_2005 = "drought"))
-
-library(dplyr)
-drought %>% 
-  group_by(dept) %>% 
-  summarise(drought = sum(drought))
+`export_summary_table(dfSummary(drought_summary))`
 
@@
 # Data Processing and Combination
+To create a complete dataset, I merged the household data with the floods data and the droughts data.
+Before merging the data, I cut down the datasets to include only the relevant variables needed for analysis.
 *This section should include a discussion of the processing and merging steps needed to create your basic data.
@@
 
-```
-
 # Analysis Variables
@@
 
+In the household data, the sample of children in 2005 was
+
 *This section should include a discussion of any data patterns you notice based on the summaries created in the code above.*

### 2020-03-23 data_appendix.Rmd
Added structure for creating a dept variable that has the name of the department instead of the number to facilitate matching with the flood and drought data
Added structure in the flood section that will read in all variables and flip the data. The same structure needs to be followed in the drought section.

@@
         grade = s3_05b,
-        repeated = s3_07,
-        dept = depto)
+        repeated = s3_07) %>% 
+  mutate(dept = case_when(depto == "1" ~ "Department 1",
+                          depto == "2" ~ "Department 2"))
+
 hh_2005 <- data_2005 %>% 
@@
 flood <- read_csv(file.path("Raw data", "recurrencia_inundacion2002_2012bol.csv")) %>% 
-  rename(dept = NOM_DEP, flood = A_2005)
+  rename(dept = NOM_DEP)
 
-flood_summary %>% weather %>% 
+flood_summary <- flood %>% 
   group_by(dept) %>% 
-  summarise(flood = sum(flood))
+  summarize_at(vars(starts_with("A_")), sum) %>% 
+  pivot_longer(-dept, names_to = "year", names_prefix = "A_", values_to = "floods")
 
-`export_summary_table(dfSummary(flood_summary))`
+export_summary_table(dfSummary(flood_summary))
 ```
@@
 #rename variables
-    rename(dept = NOM_DEP, drought = A_2005)
+    rename(dept = NOM_DEP)
+
 drought_summary %>%
@@
 
-`export_summary_table(dfSummary(drought_summary))`
+export_summary_table(dfSummary(drought_summary))
 

### 2020-03-23 data_appendix_2005.Rmd
merging again

--- 2020-03-23 data_appendix.Rmd
@@
         repeated = s3_07) %>% 
-  mutate(dept = case_when(depto == "1" ~ "Department 1",
-                          depto == "2" ~ "Department 2"))
+  mutate(dept = case_when(depto == "1" ~ "Chuquisaca",
+                          depto == "2" ~ "La Paz",
+                          depto == "3" ~ "Cochabamba",
+                          depto == "4" ~ "Oruro",
+                          depto == "5" ~ "Potosi",
+                          depto == "6" ~ "Tarija",
+                          depto == "7" ~ "Santa Cruz",
+                          depto == "8" ~ "Beni",
+                          depto == "9" ~ "Pando"))
 
@@
 
-drought_summary %>%
+drought_summary <- drought %>%
   group_by(dept) %>% 
-  summarise(flood = sum(flood))
+  summarize_at(vars(starts_with("A_")), sum) %>%
+  pivot_longer(-dept, names_to = "year", names_prefix = "A_", values_to = "droughts")
 
@@
 total_2005 <- hh_2005 %>% 
-  left_join(flood_summary)
+  left_join(flood_summary) %>%
+  left_join(drought_summary)
 ```

### 2020-03-23 data_appendix.Rmd
Moved mutate statements after select so variable list does not need to be updated every time new variables are added.

@@
         grade = s3_05b,
-        repeated = s3_07) %>% 
-  mutate(dept = case_when(depto == "1" ~ "Department 1",
-                          depto == "2" ~ "Department 2"))
+        repeated = s3_07)
 
 hh_2005 <- data_2005 %>% 
-  select(hh_id, assist, reason, members, sex, age, relation, literacy, register, grade, repeated, dept, urb_rur)
+  select(hh_id, assist, reason, members, sex, age,
+         relation, literacy, register, grade, repeated, dept0, urb_rur) %>% 
+  mutate(year = 2005,
+         dept = case_when(depto == "1" ~ "Chuquisaca",
+                          depto == "2" ~ "La Paz",
+                          depto == "3" ~ "Cochabamba",
+                          depto == "4" ~ "Oruro",
+                          depto == "5" ~ "Potosi",
+                          depto == "6" ~ "Tarija",
+                          depto == "7" ~ "Santa Cruz",
+                          depto == "8" ~ "Beni",
+                          depto == "9" ~ "Pando"))
 ```
@@
   summarize_at(vars(starts_with("A_")), sum) %>% 
-  pivot_longer(-dept, names_to = "year", names_prefix = "A_", values_to = "floods")
+  pivot_longer(-dept, names_to = "year", names_prefix = "A_", values_to = "floods") %>%
+  mutate(year = as.numeric(year))
 
@@
 
-drought_summary %>%
+drought_summary <- drought %>%
   group_by(dept) %>% 
-  summarise(flood = sum(flood))
+  summarize_at(vars(starts_with("A_")), sum) %>%
+  pivot_longer(-dept, names_to = "year", names_prefix = "A_", values_to = "droughts") %>%
+  mutate(year = as.numeric(year))
 

### 2020-03-23 data_appendix.Rmd
Fixed typo

@@
   select(hh_id, assist, reason, members, sex, age,
-         relation, literacy, register, grade, repeated, dept0, urb_rur) %>% 
+         relation, literacy, register, grade, repeated, depto, urb_rur) %>% 
   mutate(year = 2005,

### 2020-03-25 data_appendix.Rmd
adding to the data patterns part

@@
 # Appendix description
-*Your Data Appendix should begin with a brief statement explaining its purpose like the following one.*
+This data appendix details the steps of collecting, processing, and merging the data needed to create the complete data set showing information about schooling and weather events.
+It concludes with a discussion of patterns found in the data.
 
@@
 - **dept:** the department in which the person lives 
+- **income:** the per capita monthly income for each household, in bolivianos
 
@@
         grade = s3_05b,
-        repeated = s3_07)
+        repeated = s3_07,
+        income = y_perc1A)
 
@@
   select(hh_id, assist, reason, members, sex, age,
-         relation, literacy, register, grade, repeated, depto, urb_rur) %>% 
+         relation, literacy, register, grade, repeated, depto, urb_rur, income) %>% 
   mutate(year = 2005,
@@
 
-Instituto Nacional de Estadistica. (2005). Encuesta de Hogares (2005) [SAV file]. Retrieved from https://www.ine.gob.bo/index.php/banco/base-de-datos-sociales
-
-**DOI:** If the dataset has a documention identified (DOI) assigned put it here.  
-**Date Downloaded:** Identify when you downloaded the dataset.  
-**Filename(s):** raw_data/filename.csv *If you have a large number of files you can use a patten (see visit data below)*
-**Unit of observation:** What distinguishes different rows in your dataset?  
-**Dates covered:** What time frame does the data cover?
+**Date Downloaded:** October 2019
+**Filename(s):** raw_data/filename.csv 
+**Unit of observation:** Municipality  
+**Dates covered:** 2002-2012
 
@@
 
-Describe in a step-by-step fashion how an interested user could obtain the data.
-
-### Importable version (if necessary)
-
-**Filename(s):** importable-data/filename-importable.csv
-
-In some cases the raw data is not directly importable. In this case, you should fully document every step you took to create the importable data in a subsection like this one.
+This data can be found at the following website: http://geosinager.defensacivil.gob.bo/maps/226.
+It is associated with a map showing the flood occurrences information spatially, and to get the data in a non-spatial format, the flood layer should be downloaded.
+It may be downloaded in a variety of different formats, including csv.
 
@@
 
-This data can be obtained from http://geosinager.defensacivil.gob.bo/maps/225. It is shown on a map with different layers.
-
-### Importable version (if necessary)
-
-**Filename(s):** importable-data/filename-importable.csv
-
-In some cases the raw data is not directly importable. In this case, you should fully document every step you took to create the importable data in a subsection like this one.
+This data can be obtained from http://geosinager.defensacivil.gob.bo/maps/225.
+It is shown on a map with different layers, and the raw data may be downloaded in a variety of different forms, including csv.
 
@@
 # Data Processing and Combination
+
 To create a complete dataset, I merged the household data with the floods data and the droughts data.
-Before merging the data, I cut down the datasets to include only the relevant variables needed for analysis.
+Before merging the data, I cut down the datasets to include only the relevant variables needed for analysis in each file.
+I renamed several variables in the household survey data set to make them more understandable.
+I also created the year variable for the household data to ensure that it matched up correctly with the weather data.
+For the flood and drought data, I grouped the occurrences by department instead of by municipality so they would match the household data, and I made the department character numeric for the merge.
+
 *This section should include a discussion of the processing and merging steps needed to create your basic data.
@@
 
-```{r merging datasets}
-#I would recommend combining all the years of household data and then merging with the weather data
-
+``` {r merging datasets}
 total_2005 <- hh_2005 %>% 
@@
   left_join(drought_summary)
+
+export_summary_table(dfSummary(total_2005))
+
+total_2005 <- readRDS(file = analysis_variables.rds")
+
 ```
@@
 
-This section should include a description of all the variables that are used in your final analysis.
+The variables used in the final analysis are the variables from the flood and drought summaries, showing the number of occurrences in each department per year.
+The other variables used are from the household data, showing demographic information as well as the dependent variables related to school assistance and registration.
+The variables "assist" and "register" tell whether the child is currently registered and attending school, and the variables "grade" and "literacy" give further educational information showing what grade the child is in and whether they can read and write.
+The variables "urb_rur" and "dept" give geographic information, indicating whether the child's household is located in an urban area or a rural area and in what department of the country they are living.
+Demographic variables such as age, sex, relation to the head of household, number of household members, and income can be used in the analysis as controls.
+
 At the end of the section, you should save all of these variables in the processed_data folder of your repository.
@@
 
-In the household data, the sample of children in 2005 was
+The "assist" variable shows that around 60% of people who are registered for school currently assist, while 40% do not assist.
+The "reason" variable shows that two most common reasons for missing school are for work (31.3%) or vacation (20.5).
+Out of the people who are registered for school, 6.6% are repeating the grade they are currently in.
 
-*This section should include a discussion of any data patterns you notice based on the summaries created in the code above.*
+The patterns for households show that the average household has 5.3 members, and the average per capita monthly income per household is 258 bolivianos.
+Over half of the households in the sample are located in urban areas (56.3), while under half are located in rural areas (43.7).
+La Paz, Santa Cruz, and Cochabamba are the three departments with the most households in this sample, which corresponds to their overall population distribution.
+
+The patterns in the weather data show that every department experienced multiple floods in 2005, while many but not all experienced droughts.
+The place that experienced the highest number of flood occurrences overall was Cochabamba, while Beni experienced 13 droughts in 2005.

### 2020-03-25 data_appendix.Rmd
data appendix trying to get the variables saved in the processed data folder

@@
 
-Create a bullet list with the name of each variable in the dataset followed by any information the user would need to understand it.
-
 - **dept:** department name 
@@
 
-*This section should include a discussion of the processing and merging steps needed to create your basic data.
-The code to implement these steps should be included in chunks in this section.
-Once the final merged data has been created, you should use the dfSummary function again to summarize the data you will be using.
-You should also save a file containing all the objects you will use in your final analysis to the processed_data folder.*
-
 ``` {r merging datasets}
@@
 
-total_2005 <- readRDS(file = analysis_variables.rds")
+setwd(dir = "C:/Users/katya/Documents/MyDocuments/A Smith College Spring Semester/Natural Resource Economics/seminar-paper-katyagi21/processed_data/")
+
+total_2005 <- readRDS(file = "analysis_variables.rds")
 
@@
 Demographic variables such as age, sex, relation to the head of household, number of household members, and income can be used in the analysis as controls.
-
-At the end of the section, you should save all of these variables in the processed_data folder of your repository.
 

### 2020-03-28 data_appendix_all_years.Rmd
adding other household surveys

--- 2020-03-25 data_appendix.Rmd
@@
 
-```{r import household data}
-data_2005 <- read_dta(file.path("Raw data","seminar_paper_data","2005_hhsurvey.dta")) %>% 
+```{r import household data 2005}
+data_2005 <- read_dta("Raw data/seminar_paper_data/2005_hhsurvey.dta") %>% 
   rename(hh_id = folio,
@@
   mutate(year = 2005,
+         hh_id = as.numeric(hh_id),
+         dept = case_when(depto == "1" ~ "Chuquisaca",
+                          depto == "2" ~ "La Paz",
+                          depto == "3" ~ "Cochabamba",
+                          depto == "4" ~ "Oruro",
+                          depto == "5" ~ "Potosi",
+                          depto == "6" ~ "Tarija",
+                          depto == "7" ~ "Santa Cruz",
+                          depto == "8" ~ "Beni",
+                          depto == "9" ~ "Pando"))
+```
+```{r export summary table 2005}
+export_summary_table(dfSummary(hh_2005))
+```
+
+```{r import household data 2006}
+data_2006 <- read_dta("Raw data/seminar_paper_data/2006_hhsurvey.dta")  %>% 
+  rename(hh_id = Folio,
+        assist = s4_08,  
+        reason = S4_09, 
+        members = mhogar,
+        sex = s1_02,
+        age = s1_03,
+        relation = s1_05,
+        literacy = s4_01,
+        register = s4_04,
+        grade = s4_05b,
+        repeated = s4_07,
+        income = yhogpcf,
+        urb_rur = Urb_Rur)
+
+hh_2006 <- data_2006 %>% 
+  select(hh_id, assist, reason, members, sex, age,
+         relation, literacy, register, grade, repeated, ID01, urb_rur, income) %>% 
+  mutate(year = 2006,
+         hh_id = as.numeric(hh_id),
+         dept = case_when(ID01 == "1" ~ "Chuquisaca",
+                          ID01 == "2" ~ "La Paz",
+                          ID01 == "3" ~ "Cochabamba",
+                          ID01 == "4" ~ "Oruro",
+                          ID01 == "5" ~ "Potosi",
+                          ID01 == "6" ~ "Tarija",
+                          ID01 == "7" ~ "Santa Cruz",
+                          ID01 == "8" ~ "Beni",
+                          ID01 == "9" ~ "Pando"))
+```
+
+```{r exporting data table}
+export_summary_table(dfSummary(hh_2006))
+```
+
+```{r import household data 2007}
+data_2007 <- read_dta("Raw data/seminar_paper_data/2007_hhsurvey.dta")  %>% 
+  rename(hh_id = folio,
+        assist = s4_10,  
+        reason = s4_11, 
+        members = mhogar,
+        sex = s1_03,
+        age = s1_04,
+        relation = s1_06,
+        literacy = s4_01,
+        register = s4_06,
+        grade = s4_07b,
+        repeated = s4_09,
+        income = yhogpcf)
+
+hh_2007 <- data_2007 %>% 
+  select(hh_id, assist, reason, members, sex, age,
+         relation, literacy, register, grade, repeated, depto, urb_rur, income) %>% 
+  mutate(year = 2007,
+        hh_id = as.numeric(hh_id),
          dept = case_when(depto == "1" ~ "Chuquisaca",
@@
 
-```{r exporting data table}
-export_summary_table(dfSummary(hh_2005))
+```{r import household data 2011}
+data_2011 <- read_dta("Raw data/seminar_paper_data/2011_hhsurvey.dta")  %>% 
+  rename(hh_id = folio,
+        assist = s4_12,  
+        reason = s4_13,
+        sex = s1_03,
+        age = s1_04,
+        relation = s1_08,
+        literacy = s4_01,
+        register = s4_05,
+        grade = s4_06b,
+        repeated = s4_11a,
+        income = yhogpc,
+        school_level = s4_06a,
+        urb_rur = area)
+
+hh_2011 <- data_2011 %>% 
+  select(hh_id, assist, reason, sex, age,
+         relation, literacy, register, grade, repeated, depto, urb_rur, income, school_level) %>% 
+  mutate(year = 2011,
+         dept = case_when(depto == "1" ~ "Chuquisaca",
+                          depto == "2" ~ "La Paz",
+                          depto == "3" ~ "Cochabamba",
+                          depto == "4" ~ "Oruro",
+                          depto == "5" ~ "Potosi",
+                          depto == "6" ~ "Tarija",
+                          depto == "7" ~ "Santa Cruz",
+                          depto == "8" ~ "Beni",
+                          depto == "9" ~ "Pando"))
 ```
 
+```{r import household data 2012}
+data_2012 <- read_dta("Raw data/seminar_paper_data/2012_hhsurvey.dta")  %>% 
+  rename(hh_id = folio,
+        assist = s4_12,  
+        reason = s4_13,
+        sex = s1_03,
+        age = s1_04,
+        relation = s1_08,
+        literacy = s4_01,
+        register = s4_04,
+        grade = s4_05b,
+        repeated = s4_11a,
+        income = yhogpc,
+        school_level = s4_05a,
+        urb_rur = area) 
+hh_2012 <- data_2012 %>% 
+  select(hh_id, assist, reason, sex, age,
+         relation, literacy, register, grade, repeated, departamento, urb_rur, income, school_level) %>% 
+  mutate(year = 2012,
+         dept = case_when(departamento == "1" ~ "Chuquisaca",
+                          departamento == "2" ~ "La Paz",
+                          departamento == "3" ~ "Cochabamba",
+                          departamento == "4" ~ "Oruro",
+                          departamento == "5" ~ "Potosi",
+                          departamento == "6" ~ "Tarija",
+                          departamento == "7" ~ "Santa Cruz",
+                          departamento == "8" ~ "Beni",
+                          departamento == "9" ~ "Pando"))
+```
+
+```{r import household data 2013}
+data_2013 <- read_dta("Raw data/seminar_paper_data/2013_hhsurvey.dta")  %>% 
+  rename(hh_id = folio,
+        assist = s5_10,  
+        reason = s5_11,
+        sex = s2_02,
+        age = s2_03,
+        relation = s2_05,
+        literacy = s5_01,
+        register = s5_04,
+        grade = s5_05b,
+        income = yhogpc,
+        school_level = s5_05a,
+        urb_rur = area) 
+hh_2013 <- data_2013 %>% 
+  select(hh_id, assist, reason, sex, age,
+         relation, literacy, register, grade, id01, urb_rur, income, school_level) %>% 
+  mutate(year = 2013,
+         dept = case_when(id01 == "1" ~ "Chuquisaca",
+                          id01 == "2" ~ "La Paz",
+                          id01 == "3" ~ "Cochabamba",
+                          id01 == "4" ~ "Oruro",
+                          id01 == "5" ~ "Potosi",
+                          id01 == "6" ~ "Tarija",
+                          id01 == "7" ~ "Santa Cruz",
+                          id01 == "8" ~ "Beni",
+                          id01 == "9" ~ "Pando"))
+```
+
+```{r import household survey 2014}
+data_2014 <- read_dta("Raw data/seminar_paper_data/2014_hhsurvey.dta") %>%
+  rename(hh_id = folio,
+        assist = s5b_10,  
+        reason = s5b_11,
+        sex = s2a_02,
+        age = s2a_03,
+        relation = s2a_05,
+        literacy = s5a_01,
+        register = s5a_04,
+        grade = s5a_05a,
+        income = yhogpc,
+        school_level = s5a_05,
+        urb_rur = URBRUR) 
+hh_2014 <- data_2014 %>% 
+  select(hh_id, assist, reason, sex, age,
+         relation, literacy, register, grade, DEPTO, urb_rur, income, school_level) %>% 
+  mutate(year = 2014,
+         dept = case_when(DEPTO == "1" ~ "Chuquisaca",
+                          DEPTO == "2" ~ "La Paz",
+                          DEPTO == "3" ~ "Cochabamba",
+                          DEPTO == "4" ~ "Oruro",
+                          DEPTO == "5" ~ "Potosi",
+                          DEPTO == "6" ~ "Tarija",
+                          DEPTO == "7" ~ "Santa Cruz",
+                          DEPTO == "8" ~ "Beni",
+                          DEPTO == "9" ~ "Pando"))
+```
+
+```{r import household survey 2015}
+data_2015 <- read_dta("Raw data/seminar_paper_data/2015_hhsurvey.dta") %>%
+   rename(hh_id = folio,
+        assist = s5b_10,  
+        reason = s5b_11,
+        sex = s2a_02,
+        age = s2a_03,
+        relation = s2a_05,
+        literacy = s5a_1,
+        register = s5a_4,
+        grade = s5a_5b,
+        income = yhogpc,
+        school_level = s5a_5a,
+        urb_rur = area) 
+hh_2015 <- data_2015 %>% 
+  select(hh_id, assist, reason, sex, age,
+         relation, literacy, register, grade, departamento, urb_rur, income, school_level) %>% 
+  mutate(year = 2015,
+         dept = case_when(departamento == "1" ~ "Chuquisaca",
+                          departamento == "2" ~ "La Paz",
+                          departamento == "3" ~ "Cochabamba",
+                          departamento == "4" ~ "Oruro",
+                          departamento == "5" ~ "Potosi",
+                          departamento == "6" ~ "Tarija",
+                          departamento == "7" ~ "Santa Cruz",
+                          departamento == "8" ~ "Beni",
+                          departamento == "9" ~ "Pando"))
+
+```
+
+```{r import household data 2016}
+data_2016 <- read_dta("Raw data/seminar_paper_data/2016_hhsurvey.dta")  %>%
+   rename(hh_id = folio,
+        assist = s05b_10,  
+        reason = s05b_11,
+        sex = s02a_02,
+        age = s02a_03,
+        relation = s02a_05,
+        literacy = s05a_01,
+        register = s05a_05,
+        grade = s05a_06b,
+        income = yhogpc,
+        school_level = s05a_06a,
+        urb_rur = area) 
+hh_2016 <- data_2016 %>% 
+  select(hh_id, assist, reason, sex, age,
+         relation, literacy, register, grade, depto, urb_rur, income, school_level) %>% 
+  mutate(year = 2016,
+         dept = case_when(depto == "1" ~ "Chuquisaca",
+                          depto == "2" ~ "La Paz",
+                          depto == "3" ~ "Cochabamba",
+                          depto == "4" ~ "Oruro",
+                          depto == "5" ~ "Potosi",
+                          depto == "6" ~ "Tarija",
+                          depto == "7" ~ "Santa Cruz",
+                          depto == "8" ~ "Beni",
+                          depto == "9" ~ "Pando"))
+
+```
+
+```{r import household survey 2017}
+data_2017 <- read_dta("Raw data/seminar_paper_data/2017_hhsurvey.dta")  %>%
+  rename(hh_id = folio,
+        assist = s05b_10,  
+        reason = s05b_11,
+        sex = s02a_02,
+        age = s02a_03,
+        relation = s02a_05,
+        literacy = s05a_01,
+        register = s05a_05,
+        grade = s05a_06b,
+        income = yhogpc,
+        school_level = s05a_06a,
+        urb_rur = area) 
+hh_2017 <- data_2017 %>% 
+  select(hh_id, assist, reason, sex, age,
+         relation, literacy, register, grade, depto, urb_rur, income, school_level) %>% 
+  mutate(year = 2017,
+         dept = case_when(depto == "1" ~ "Chuquisaca",
+                          depto == "2" ~ "La Paz",
+                          depto == "3" ~ "Cochabamba",
+                          depto == "4" ~ "Oruro",
+                          depto == "5" ~ "Potosi",
+                          depto == "6" ~ "Tarija",
+                          depto == "7" ~ "Santa Cruz",
+                          depto == "8" ~ "Beni",
+                          depto == "9" ~ "Pando"))
+
+```
+
+```{r import household data 2018}
+data_2018 <- read_dta("Raw data/seminar_paper_data/2018_hhsurvey.dta")  %>%
+  rename(hh_id = folio,
+        assist = s05b_10,  
+        reason = s05b_11,
+        sex = s02a_02,
+        age = s02a_03,
+        relation = s02a_05,
+        literacy = s05a_01,
+        register = s05a_05,
+        grade = s05a_06c,
+        income = yhogpc,
+        school_level = s05a_06a,
+        urb_rur = area) 
+hh_2018 <- data_2018 %>% 
+  select(hh_id, assist, reason, sex, age,
+         relation, literacy, register, grade, depto, urb_rur, income, school_level) %>% 
+  mutate(year = 2018,
+         hh_id = as.numeric(hh_id),
+         dept = case_when(depto == "1" ~ "Chuquisaca",
+                          depto == "2" ~ "La Paz",
+                          depto == "3" ~ "Cochabamba",
+                          depto == "4" ~ "Oruro",
+                          depto == "5" ~ "Potosi",
+                          depto == "6" ~ "Tarija",
+                          depto == "7" ~ "Santa Cruz",
+                          depto == "8" ~ "Beni",
+                          depto == "9" ~ "Pando"))
+```
+
+```{r merge household datasets}
+```
+
+{r import household survey 2008}
+data_2008 <- read_dta("Raw data/seminar_paper_data/2008_hhsurvey.dta")  %>% 
+  rename(hh_id = FOLIO,
+        assist = S4_13,  
+        reason = S4_14, 
+        members = MHOGAR,
+        sex = S1_03,
+        age = S1_04,
+        relation = S1_06,
+        literacy = S4_01,
+        register = S4_05,
+        grade = S4_06B,
+        repeated = S4_12,
+        income = YHOGPCF)
+
+hh_2008 <- data_2008 %>% 
+  select(hh_id, assist, reason, members, sex, age,
+         relation, literacy, register, grade, repeated, depto, urb_rur, income) %>% 
+  mutate(year = 2008,
+         dept = case_when(depto == "1" ~ "Chuquisaca",
+                          depto == "2" ~ "La Paz",
+                          depto == "3" ~ "Cochabamba",
+                          depto == "4" ~ "Oruro",
+                          depto == "5" ~ "Potosi",
+                          depto == "6" ~ "Tarija",
+                          depto == "7" ~ "Santa Cruz",
+                          depto == "8" ~ "Beni",
+                          depto == "9" ~ "Pando"))
+```
 ## Flood data
@@
 ```
+
 ## Drought data
@@
 
-setwd(dir = "C:/Users/katya/Documents/MyDocuments/A Smith College Spring Semester/Natural Resource Economics/seminar-paper-katyagi21/processed_data/")
-
-total_2005 <- readRDS(file = "analysis_variables.rds")
+save("total_2005", file = "processed_data/analysis_data.RData")
 

### 2020-03-28 data_appendix_all_years.Rmd
categorical variable recode

@@
 ```
+
+```{r recoding categorical variables 2005}
+hh_2005
+reason <- as.factor(c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10))
+recode(reason, "1='vacation'; 2 = 'money'; 3 = 'work'; 4 = 'sickness/accident/disability'; 5 ='distance'; 6 = 'finished'; 7 = 'age'; 8 = 'not interested'; 9 = 'housework/childcare'; 10 = 'other'")
+
+```
+
 ```{r export summary table 2005}

### 2020-04-05 data_appendix_all_years.Rmd
Used case_when instead of recode to recode the reason variable.

--- 2020-03-28 data_appendix_all_years.Rmd
@@
                           depto == "8" ~ "Beni",
-                          depto == "9" ~ "Pando"))
+                          depto == "9" ~ "Pando"),
+         reason = case_when(reason == 1 ~ 'vacation',
+                            reason == 2 ~ 'money', 
+                            reason == 3 ~ 'work',
+                            reason == 4 ~ 'sickness/accident/disability',
+                            reason == 5 ~ 'distance', 
+                            reason == 6 ~ 'finished', 
+                            reason == 7 ~ 'age',
+                            reason == 8 ~ 'not interested',
+                            reason == 9 ~ 'housework/childcare',
+                            reason == 10 ~ 'other'))
 ```
+
+Note: remove this section after reading it. I added the code for recoding the variable in the above section, but for learning purposes I wanted to tell you what the code you had does:
+
+hh_2005 
+prints the head of the hh_2005 data to the console but doesn't do anything with the data
+
+reason <- as.factor(c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10))
+creates a new object called reason that is a factor vector containing the numbers 1-10.
+But this line doesn't know that this is connected to the hh_2005 dataset.
+Notice that above, I used the pipe operator to tell R to take the dataset we had been working with and then recode a variable from that dataset
+
+recode(reason, "1='vacation'; 2 = 'money'; 3 = 'work'; 4 = 'sickness/accident/disability'; 5 ='distance'; 6 = 'finished'; 7 = 'age'; 8 = 'not interested'; 9 = 'housework/childcare'; 10 = 'other'") 
+returns an error because the list of reasons and numbers are one long character string that R can't parse and because the haven package creates objects that don't work with some of the R functions (like recode, apparently)
+
+```
+
 ```{r export summary table 2005}

### 2020-04-14 data_appendix_all_years.Rmd
created household head and children data frames

--- 2020-03-28 data_appendix_all_years.Rmd
@@
         register = s3_04,
+        school_level = s3_05a,
         grade = s3_05b,
@@
 hh_2005 <- data_2005 %>% 
-  select(hh_id, assist, reason, members, sex, age,
-         relation, literacy, register, grade, repeated, depto, urb_rur, income) %>% 
+  select(hh_id, assist, reason, sex, age,
+         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income) %>% 
   mutate(year = 2005,
-         hh_id = as.numeric(hh_id),
          dept = case_when(depto == "1" ~ "Chuquisaca",
@@
                           depto == "8" ~ "Beni",
-                          depto == "9" ~ "Pando"))
+                          depto == "9" ~ "Pando"),
+         reason = case_when(reason == 1 ~ 'vacation',
+                            reason == 2 ~ 'money', 
+                            reason == 3 ~ 'work', 
+                            reason == 4 ~ 'sickness/accident/disability', 
+                            reason == 5 ~ 'distance',
+                            reason == 6 ~ 'finished', 
+                            reason == 7 ~ 'age',
+                            reason == 8 ~ 'not interested', 
+                            reason == 9 ~ 'housework/childcare',
+                            reason == 10 ~ 'other'))
+
 ```
+
 ```{r export summary table 2005}
 export_summary_table(dfSummary(hh_2005))
+
 ```
@@
         register = s4_04,
+        school_level = s4_05a,
         grade = s4_05b,
@@
         income = yhogpcf,
-        urb_rur = Urb_Rur)
+        urb_rur = Urb_Rur,
+        depto = ID01)
 
 hh_2006 <- data_2006 %>% 
-  select(hh_id, assist, reason, members, sex, age,
-         relation, literacy, register, grade, repeated, ID01, urb_rur, income) %>% 
+  select(hh_id, assist, reason, sex, age,
+         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income) %>% 
   mutate(year = 2006,
-         hh_id = as.numeric(hh_id),
-         dept = case_when(ID01 == "1" ~ "Chuquisaca",
-                          ID01 == "2" ~ "La Paz",
-                          ID01 == "3" ~ "Cochabamba",
-                          ID01 == "4" ~ "Oruro",
-                          ID01 == "5" ~ "Potosi",
-                          ID01 == "6" ~ "Tarija",
-                          ID01 == "7" ~ "Santa Cruz",
-                          ID01 == "8" ~ "Beni",
-                          ID01 == "9" ~ "Pando"))
+         dept = case_when(depto == "1" ~ "Chuquisaca",
+                          depto == "2" ~ "La Paz",
+                          depto == "3" ~ "Cochabamba",
+                          depto == "4" ~ "Oruro",
+                          depto == "5" ~ "Potosi",
+                          depto == "6" ~ "Tarija",
+                          depto == "7" ~ "Santa Cruz",
+                          depto == "8" ~ "Beni",
+                          depto == "9" ~ "Pando"),
+           reason = case_when(reason == 1 ~ 'vacation',
+                            reason == 2 ~ 'money', 
+                            reason == 3 ~ 'work', 
+                            reason == 4 ~ 'sickness/accident/disability', 
+                            reason == 5 ~ 'distance',
+                            reason == 6 ~ 'finished', 
+                            reason == 7 ~ 'age',
+                            reason == 8 ~ 'not interested', 
+                            reason == 9 ~ 'housework/childcare',
+                            reason == 10 ~ 'other'))
 ```
@@
         register = s4_06,
+        school_level = s4_07a,
         grade = s4_07b,
@@
 hh_2007 <- data_2007 %>% 
-  select(hh_id, assist, reason, members, sex, age,
-         relation, literacy, register, grade, repeated, depto, urb_rur, income) %>% 
+  select(hh_id, assist, reason, sex, age,
+         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income) %>% 
   mutate(year = 2007,
-        hh_id = as.numeric(hh_id),
          dept = case_when(depto == "1" ~ "Chuquisaca",
@@
                           depto == "8" ~ "Beni",
-                          depto == "9" ~ "Pando"))
+                          depto == "9" ~ "Pando"),
+           reason = case_when(reason == 1 ~ 'vacation',
+                            reason == 2 ~ 'money', 
+                            reason == 3 ~ 'work', 
+                            reason == 4 ~ 'sickness/accident/disability', 
+                            reason == 5 ~ 'distance',
+                            reason == 6 ~ 'finished', 
+                            reason == 7 ~ 'age',
+                            reason == 8 ~ 'not interested', 
+                            reason == 9 ~ 'housework/childcare',
+                            reason == 10 ~ 'other'))
 ```
@@
   select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, grade, repeated, depto, urb_rur, income, school_level) %>% 
+         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income) %>% 
   mutate(year = 2011,
@@
                           depto == "8" ~ "Beni",
-                          depto == "9" ~ "Pando"))
+                          depto == "9" ~ "Pando"),
+           reason = case_when(reason == 1 ~ 'vacation',
+                            reason == 2 ~ 'money', 
+                            reason == 3 ~ 'work', 
+                            reason == 4 ~ 'sickness/accident/disability', 
+                            reason == 5 ~ 'distance',
+                            reason == 6 ~ 'finished', 
+                            reason == 7 ~ 'age',
+                            reason == 8 ~ 'not interested', 
+                            reason == 9 ~ 'housework/childcare',
+                            reason == 10 ~ 'other'))
 ```
 
-```{r import household data 2012}
+```{r household data 2012}
 data_2012 <- read_dta("Raw data/seminar_paper_data/2012_hhsurvey.dta")  %>% 
@@
         school_level = s4_05a,
-        urb_rur = area) 
+        urb_rur = area,
+        depto = departamento)
+
 hh_2012 <- data_2012 %>% 
   select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, grade, repeated, departamento, urb_rur, income, school_level) %>% 
+         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income) %>% 
   mutate(year = 2012,
-         dept = case_when(departamento == "1" ~ "Chuquisaca",
-                          departamento == "2" ~ "La Paz",
-                          departamento == "3" ~ "Cochabamba",
-                          departamento == "4" ~ "Oruro",
-                          departamento == "5" ~ "Potosi",
-                          departamento == "6" ~ "Tarija",
-                          departamento == "7" ~ "Santa Cruz",
-                          departamento == "8" ~ "Beni",
-                          departamento == "9" ~ "Pando"))
-```
-
-```{r import household data 2013}
-data_2013 <- read_dta("Raw data/seminar_paper_data/2013_hhsurvey.dta")  %>% 
-  rename(hh_id = folio,
-        assist = s5_10,  
-        reason = s5_11,
-        sex = s2_02,
-        age = s2_03,
-        relation = s2_05,
-        literacy = s5_01,
-        register = s5_04,
-        grade = s5_05b,
-        income = yhogpc,
-        school_level = s5_05a,
-        urb_rur = area) 
-hh_2013 <- data_2013 %>% 
-  select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, grade, id01, urb_rur, income, school_level) %>% 
-  mutate(year = 2013,
-         dept = case_when(id01 == "1" ~ "Chuquisaca",
-                          id01 == "2" ~ "La Paz",
-                          id01 == "3" ~ "Cochabamba",
-                          id01 == "4" ~ "Oruro",
-                          id01 == "5" ~ "Potosi",
-                          id01 == "6" ~ "Tarija",
-                          id01 == "7" ~ "Santa Cruz",
-                          id01 == "8" ~ "Beni",
-                          id01 == "9" ~ "Pando"))
-```
-
-```{r import household survey 2014}
-data_2014 <- read_dta("Raw data/seminar_paper_data/2014_hhsurvey.dta") %>%
-  rename(hh_id = folio,
-        assist = s5b_10,  
-        reason = s5b_11,
-        sex = s2a_02,
-        age = s2a_03,
-        relation = s2a_05,
-        literacy = s5a_01,
-        register = s5a_04,
-        grade = s5a_05a,
-        income = yhogpc,
-        school_level = s5a_05,
-        urb_rur = URBRUR) 
-hh_2014 <- data_2014 %>% 
-  select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, grade, DEPTO, urb_rur, income, school_level) %>% 
-  mutate(year = 2014,
-         dept = case_when(DEPTO == "1" ~ "Chuquisaca",
-                          DEPTO == "2" ~ "La Paz",
-                          DEPTO == "3" ~ "Cochabamba",
-                          DEPTO == "4" ~ "Oruro",
-                          DEPTO == "5" ~ "Potosi",
-                          DEPTO == "6" ~ "Tarija",
-                          DEPTO == "7" ~ "Santa Cruz",
-                          DEPTO == "8" ~ "Beni",
-                          DEPTO == "9" ~ "Pando"))
-```
-
-```{r import household survey 2015}
-data_2015 <- read_dta("Raw data/seminar_paper_data/2015_hhsurvey.dta") %>%
-   rename(hh_id = folio,
-        assist = s5b_10,  
-        reason = s5b_11,
-        sex = s2a_02,
-        age = s2a_03,
-        relation = s2a_05,
-        literacy = s5a_1,
-        register = s5a_4,
-        grade = s5a_5b,
-        income = yhogpc,
-        school_level = s5a_5a,
-        urb_rur = area) 
-hh_2015 <- data_2015 %>% 
-  select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, grade, departamento, urb_rur, income, school_level) %>% 
-  mutate(year = 2015,
-         dept = case_when(departamento == "1" ~ "Chuquisaca",
-                          departamento == "2" ~ "La Paz",
-                          departamento == "3" ~ "Cochabamba",
-                          departamento == "4" ~ "Oruro",
-                          departamento == "5" ~ "Potosi",
-                          departamento == "6" ~ "Tarija",
-                          departamento == "7" ~ "Santa Cruz",
-                          departamento == "8" ~ "Beni",
-                          departamento == "9" ~ "Pando"))
-
-```
-
-```{r import household data 2016}
-data_2016 <- read_dta("Raw data/seminar_paper_data/2016_hhsurvey.dta")  %>%
-   rename(hh_id = folio,
-        assist = s05b_10,  
-        reason = s05b_11,
-        sex = s02a_02,
-        age = s02a_03,
-        relation = s02a_05,
-        literacy = s05a_01,
-        register = s05a_05,
-        grade = s05a_06b,
-        income = yhogpc,
-        school_level = s05a_06a,
-        urb_rur = area) 
-hh_2016 <- data_2016 %>% 
-  select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, grade, depto, urb_rur, income, school_level) %>% 
-  mutate(year = 2016,
-         dept = case_when(depto == "1" ~ "Chuquisaca",
+         dept = case_when(depto== "1" ~ "Chuquisaca",
                           depto == "2" ~ "La Paz",
@@
                           depto == "8" ~ "Beni",
-                          depto == "9" ~ "Pando"))
+                          depto == "9" ~ "Pando"),
+           reason = case_when(reason == 1 ~ 'vacation',
+                            reason == 2 ~ 'money', 
+                            reason == 3 ~ 'work', 
+                            reason == 4 ~ 'sickness/accident/disability', 
+                            reason == 5 ~ 'distance',
+                            reason == 6 ~ 'finished', 
+                            reason == 7 ~ 'age',
+                            reason == 8 ~ 'not interested', 
+                            reason == 9 ~ 'housework/childcare',
+                            reason == 10 ~ 'other'))
+```
+
+```{r combine household datasets}
+all_years <- hh_2005 %>% 
+  #bind_rows(hh_2005, hh_2006, hh_2007, hh_2011, hh_2012)
+  rbind(hh_2005, hh_2006)
 
@@
 
-```{r import household survey 2017}
-data_2017 <- read_dta("Raw data/seminar_paper_data/2017_hhsurvey.dta")  %>%
-  rename(hh_id = folio,
-        assist = s05b_10,  
-        reason = s05b_11,
-        sex = s02a_02,
-        age = s02a_03,
-        relation = s02a_05,
-        literacy = s05a_01,
-        register = s05a_05,
-        grade = s05a_06b,
-        income = yhogpc,
-        school_level = s05a_06a,
-        urb_rur = area) 
-hh_2017 <- data_2017 %>% 
-  select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, grade, depto, urb_rur, income, school_level) %>% 
-  mutate(year = 2017,
-         dept = case_when(depto == "1" ~ "Chuquisaca",
-                          depto == "2" ~ "La Paz",
-                          depto == "3" ~ "Cochabamba",
-                          depto == "4" ~ "Oruro",
-                          depto == "5" ~ "Potosi",
-                          depto == "6" ~ "Tarija",
-                          depto == "7" ~ "Santa Cruz",
-                          depto == "8" ~ "Beni",
-                          depto == "9" ~ "Pando"))
-
-```
-
-```{r import household data 2018}
-data_2018 <- read_dta("Raw data/seminar_paper_data/2018_hhsurvey.dta")  %>%
-  rename(hh_id = folio,
-        assist = s05b_10,  
-        reason = s05b_11,
-        sex = s02a_02,
-        age = s02a_03,
-        relation = s02a_05,
-        literacy = s05a_01,
-        register = s05a_05,
-        grade = s05a_06c,
-        income = yhogpc,
-        school_level = s05a_06a,
-        urb_rur = area) 
-hh_2018 <- data_2018 %>% 
-  select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, grade, depto, urb_rur, income, school_level) %>% 
-  mutate(year = 2018,
-         hh_id = as.numeric(hh_id),
-         dept = case_when(depto == "1" ~ "Chuquisaca",
-                          depto == "2" ~ "La Paz",
-                          depto == "3" ~ "Cochabamba",
-                          depto == "4" ~ "Oruro",
-                          depto == "5" ~ "Potosi",
-                          depto == "6" ~ "Tarija",
-                          depto == "7" ~ "Santa Cruz",
-                          depto == "8" ~ "Beni",
-                          depto == "9" ~ "Pando"))
-```
-
-```{r merge household datasets}
-```
-
-{r import household survey 2008}
-data_2008 <- read_dta("Raw data/seminar_paper_data/2008_hhsurvey.dta")  %>% 
-  rename(hh_id = FOLIO,
-        assist = S4_13,  
-        reason = S4_14, 
-        members = MHOGAR,
-        sex = S1_03,
-        age = S1_04,
-        relation = S1_06,
-        literacy = S4_01,
-        register = S4_05,
-        grade = S4_06B,
-        repeated = S4_12,
-        income = YHOGPCF)
-
-hh_2008 <- data_2008 %>% 
-  select(hh_id, assist, reason, members, sex, age,
-         relation, literacy, register, grade, repeated, depto, urb_rur, income) %>% 
-  mutate(year = 2008,
-         dept = case_when(depto == "1" ~ "Chuquisaca",
-                          depto == "2" ~ "La Paz",
-                          depto == "3" ~ "Cochabamba",
-                          depto == "4" ~ "Oruro",
-                          depto == "5" ~ "Potosi",
-                          depto == "6" ~ "Tarija",
-                          depto == "7" ~ "Santa Cruz",
-                          depto == "8" ~ "Beni",
-                          depto == "9" ~ "Pando"))
-```
 ## Flood data
@@
 ```
-
 ## Drought data
@@
 ``` {r merging datasets}
-total_2005 <- hh_2005 %>% 
+total_data<- all_years %>% 
   left_join(flood_summary) %>%
@@
 
-export_summary_table(dfSummary(total_2005))
+export_summary_table(dfSummary(total_data))
 
-save("total_2005", file = "processed_data/analysis_data.RData")
+save("total_data", file = "processed_data/analysis_data.RData")
 
@@
 Demographic variables such as age, sex, relation to the head of household, number of household members, and income can be used in the analysis as controls.
+
+```{r creating head of household subset}
+
+hh_head <- subset(total_data, relation == 1, select = c("hh_id", "sex", "age", "literacy", "dept", "urb_rur", "income", "year"))
+
+```
+
+```{r creating children subset}
+
+children <- subset(total_data, age < 18 & age > 4, select = c("hh_id", "age", "sex", "literacy", "dept", "urb_rur", "income", "assist", "reason", "register", "school_level", "grade", "repeated", "floods", "droughts"))
+
+```
 

### 2020-04-14 data_appendix_all_years.Rmd
got the final data set

@@
 all_years <- hh_2005 %>% 
+  rbind(hh_2005, hh_2006, hh_2007, hh_2011, hh_2012)
   #bind_rows(hh_2005, hh_2006, hh_2007, hh_2011, hh_2012)
-  rbind(hh_2005, hh_2006)
+  #rbind(hh_2005, hh_2006)
 
@@
 
-children <- subset(total_data, age < 18 & age > 4, select = c("hh_id", "age", "sex", "literacy", "dept", "urb_rur", "income", "assist", "reason", "register", "school_level", "grade", "repeated", "floods", "droughts"))
+children <- subset(total_data, age < 18 & age > 4, select = c("hh_id", "age", "sex", "relation", "literacy", "dept", "urb_rur", "income", "assist", "reason", "register", "school_level", "grade", "repeated", "floods", "droughts", "year"))
+
+```
+
+```{r combining household head characteristics with children dataset}
+hh_head_2 <- hh_head %>%
+  select(sex, age, literacy, hh_id) %>%
+  rename (female_parent = sex,
+          parent_age = age,
+          parent_literacy = literacy)
+
+total_children <- children %>%
+  left_join(hh_head_2)
+
+```
+```{r creating summary statistics}
+
+export_summary_table(dfSummary(total_children))
 

### 2020-04-24 data_appendix_all_years.Rmd
new regression changes

--- 2020-04-14 data_appendix_all_years.Rmd
@@
         repeated = s3_07,
-        income = y_perc1A)
+        income = y_perc1A,
+        educ = aoesc,
+        job = s4_16cod,
+        tasks = s4_16b,
+        pov_line = z,
+        extr_pov = z_ext)
 
@@
   select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income) %>% 
+         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
   mutate(year = 2005,
@@
         urb_rur = Urb_Rur,
-        depto = ID01)
+        depto = ID01,
+        educ = añoe,
+        job = s5_17cod,
+        tasks = s5_16b,
+        pov_line = z,
+        extr_pov = zext)
 
@@
   select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income) %>% 
+         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
   mutate(year = 2006,
@@
         repeated = s4_09,
-        income = yhogpcf)
+        income = yhogpcf,
+        voucher = s4_05,
+        educ = a_oe,
+        job = s5_16co,
+        tasks = s5_16b,
+        pov_line = z,
+        extr_pov = zext)
 
@@
   select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income) %>% 
+         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
   mutate(year = 2007,
@@
         school_level = s4_06a,
-        urb_rur = area)
+        urb_rur = area,
+        breakfast = s4_07,
+        voucher = s4_08,
+        voucher_spent = s4_09,
+        educ = e,
+        job = cod_16a,
+        tasks = s5_16b,
+        pov_line = z,
+        extr_pov = zext)
 
@@
   select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income) %>% 
+         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
   mutate(year = 2011,
@@
         urb_rur = area,
-        depto = departamento)
+        depto = departamento,
+        breakfast = s4_07,
+        voucher = s4_08,
+        voucher_spent = s4_09,
+        educ = e,
+        job = cods5_16a,
+        tasks = s5_16b,
+        pov_line = z,
+        extr_pov = zext)
 
@@
   select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income) %>% 
+         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
   mutate(year = 2012,
@@
 all_years <- hh_2005 %>% 
-  #bind_rows(hh_2005, hh_2006, hh_2007, hh_2011, hh_2012)
-  rbind(hh_2005, hh_2006)
+  rbind(hh_2005, hh_2006, hh_2007, hh_2011, hh_2012)
 
@@
 
-hh_head <- subset(total_data, relation == 1, select = c("hh_id", "sex", "age", "literacy", "dept", "urb_rur", "income", "year"))
+hh_head <- subset(total_data, relation == 1, select = c("hh_id", "sex", "age", "literacy", "dept", "urb_rur", "income", "year", "educ", "job", "tasks", "pov_line", "extr_pov"))
 
@@
 
-children <- subset(total_data, age < 18 & age > 4, select = c("hh_id", "age", "sex", "literacy", "dept", "urb_rur", "income", "assist", "reason", "register", "school_level", "grade", "repeated", "floods", "droughts"))
+children <- subset(total_data, age < 18 & age > 4, select = c("hh_id", "age", "sex", "relation", "literacy", "dept", "urb_rur", "income", "assist", "reason", "register", "school_level", "grade", "repeated", "floods", "droughts", "year", "job", "tasks"))
+
+total_children <- mutate(children, id = rownames(children))
+
+```
+
+```{r combining household head characteristics with children dataset}
+hh_head_2 <- hh_head %>%
+  select(sex, age, literacy, educ, job, tasks, pov_line, extr_pov, hh_id) %>%
+  rename (female_hhhead = sex,
+          hhhead_age = age,
+          hhhead_literacy = literacy,
+          hhhead_educ = educ,
+          hhhead_job = job,
+          hhhead_tasks = tasks)
+
+total <- total_children %>%
+  left_join(hh_head_2)
+
+#total <- merge(children,hh_head_2, by="hh_id")
+
+```
+
+```{r creating summary statistics}
+
+export_summary_table(dfSummary(total_children))
+
+```
+
+```{r regression variable specifications}
+reg_data <- total %>%
+  mutate(lny = log(income))
+```
+
+```{r regressions total sample}
+install.packages("plm")
+library(plm)
+#example code:
+#model_levels <- lm(recreation_visits ~ svr_median + svr_p10 + month_code*unit_code, data = regression_data)
+#grun.fe <- plm(inv~value+capital, data = Grunfeld, model = "within")
+
+model_simple <- plm(assist ~ floods + droughts, data = reg_data, model = "within")
+summary(model_simple)
+model_all <- plm(assist ~ floods + droughts + sex + age + lny + hhhead_educ + female_hhhead + urb_rur, data = reg_data, model = "within")
+summary(model_all)
+model_register <- plm(register ~ floods + droughts + sex + age + lny + hhhead_educ + female_hhhead + urb_rur, data = reg_data, model = "within")
+summary(model_register)
+```
+
+```{r regression rural sample}
+
+```
+
+```{r regression urban sample}
 

### 2020-04-24 data_appendix_all_years.Rmd
if this were a panel but it's not

@@
 
+#set as panel data
+panel <- plm.data(reg_data, index = c("id", "year"))
 model_simple <- plm(assist ~ floods + droughts, data = reg_data, model = "within")

### 2020-04-24 data_appendix_all_years.Rmd
linear regressions

--- 2020-04-24 data_appendix_all_years.Rmd
@@
 ```{r regressions total sample}
-install.packages("plm")
-library(plm)
+
 #example code:
 #model_levels <- lm(recreation_visits ~ svr_median + svr_p10 + month_code*unit_code, data = regression_data)
-#grun.fe <- plm(inv~value+capital, data = Grunfeld, model = "within")
 
-model_simple <- plm(assist ~ floods + droughts, data = reg_data, model = "within")
+model_simple <- lm(assist ~ floods + droughts, data = reg_data)
 summary(model_simple)
-model_all <- plm(assist ~ floods + droughts + sex + age + lny + hhhead_educ + female_hhhead + urb_rur, data = reg_data, model = "within")
+model_all <- lm(assist ~ floods + droughts + sex + age + lny + hhhead_educ + female_hhhead + urb_rur, data = reg_data)
 summary(model_all)
-model_register <- plm(register ~ floods + droughts + sex + age + lny + hhhead_educ + female_hhhead + urb_rur, data = reg_data, model = "within")
+model_register <- lm(register ~ floods + droughts + sex + age + lny + hhhead_educ + female_hhhead + urb_rur, data = reg_data)
 summary(model_register)
+
 ```

### 2020-04-24 data_appendix_all_years.Rmd
trying to fix duplicates...still 406 extra observations somehow between the children dataset and the merge with the heads of households' characteristics

@@
 
-total_children <- mutate(children, id = rownames(children))
+household_counts <- children %>%
+group_by(hh_id) %>%
+summarize(num_children = n())
+
+children_id <- mutate(children, id = rownames(children))
+
+total_children <- merge(children_id,household_counts, by="hh_id")
 
@@
           hhhead_tasks = tasks)
+total_heads_subset <- hh_head_2 %>% distinct()
 
-total <- total_children %>%
-  left_join(hh_head_2)
+#total <- total_children %>%
+  #left_join(hh_head_2)
+#left join command creating the duplicates problem it seems 
+total <- merge(total_children,hh_head_2, by="hh_id")
 
-#total <- merge(children,hh_head_2, by="hh_id")
+total_subset <- total %>% distinct()
 
@@
 ```{r regression variable specifications}
-reg_data <- total %>%
+reg_data <- total_subset %>%
   mutate(lny = log(income))

### 2020-04-24 data_appendix_all_years.Rmd
working on summary stats

@@
----
-title: "Data Appendix to \"The impact of weather events on schooling outcomes\""
+title: "Data Appendix to "The impact of weather events on schooling outcomes""
 author: "Katya Garcia-Israel"
 output:
+  html_document:
+    toc: yes
+    df_print: paged
   pdf_document:
-    toc: true
-    number_sections: true
+    toc: yes
+    number_sections: yes
 ---
@@
 
-export_summary_table(dfSummary(total_children))
+export_summary_table(dfSummary(total_subset))
+
+install.packages("stargazer")
+library(stargazer)
+stargazer(attitude)
 
@@
 ```{r regression variable specifications}
+total_subset %>%
+  mutate(ln_income = log(income))
 reg_data <- total_subset %>%
-  mutate(lny = log(income))
+  mutate(year_factor = as.factor(year))
 ```
@@
 
-model_simple <- lm(assist ~ floods + droughts, data = reg_data)
+model_simple <- lm(assist ~ floods + droughts + year, data = reg_data)
 summary(model_simple)
-model_all <- lm(assist ~ floods + droughts + sex + age + lny + hhhead_educ + female_hhhead + urb_rur, data = reg_data)
-summary(model_all)
-model_register <- lm(register ~ floods + droughts + sex + age + lny + hhhead_educ + female_hhhead + urb_rur, data = reg_data)
-summary(model_register)
+model_assist <- lm(assist ~ floods + droughts + age + sex + income + hhhead_educ + female_hhhead + urb_rur + year, data = reg_data)
+print(model_assist)
+model_register <- lm(register ~ floods + droughts + age + sex + income + hhhead_educ + female_hhhead + urb_rur + year, data = reg_data)
+print(model_register)
 
+```
+```{r generate regression table, echo = F, results = "asis"}
+stargazer::stargazer(list(model_assist, model_register), 
+               title = "Full sample regression results", 
+               label = "fullResults",
+               header = F,
+               keep = c("floods", "droughts"))
 ```

### 2020-04-24 data_appendix_all_years.Rmd
commit

--- 2020-04-24 data_appendix_all_years.Rmd
@@
 all_years <- hh_2005 %>% 
-  rbind(hh_2005, hh_2006, hh_2007, hh_2011, hh_2012)
+  rbind(hh_2005, hh_2006, hh_2007, hh_2011, hh_2012) %>%
 
@@
 
-```{r creating summary statistics}
+{r creating summary statistics}
+export_summary_table(dfSummary(total_subset))
 
-export_summary_table(dfSummary(total_children))
-
-```
+install.packages("devtools")
+devtools::install_github("ewenharrison/finalfit")
+library(finalfit)
+library(dplyr)
 
 ```{r regression variable specifications}
+total_subset %>%
+  mutate(ln_income = log(income))
 reg_data <- total_subset %>%
-  mutate(lny = log(income))
+  mutate(year_factor = as.factor(year))
+
 ```
@@
 
-model_simple <- lm(assist ~ floods + droughts, data = reg_data)
-summary(model_simple)
-model_all <- lm(assist ~ floods + droughts + sex + age + lny + hhhead_educ + female_hhhead + urb_rur, data = reg_data)
-summary(model_all)
-model_register <- lm(register ~ floods + droughts + sex + age + lny + hhhead_educ + female_hhhead + urb_rur, data = reg_data)
+#model_simple <- lm(assist ~ floods + droughts + year, data = reg_data)
+#summary(model_simple)
+#model_assist <- lm(assist ~ floods + droughts + age + sex + income + hhhead_educ + female_hhhead + year, data = reg_data)
+#summary(model_assist)
+model_register <- lm(register ~ floods + droughts + age + sex + income + hhhead_educ + female_hhhead + year, data = reg_data)
 summary(model_register)
@@
 
+```{r generate regression table, echo = F, results = "asis"}
+stargazer::stargazer(list(model_assist, model_register), 
+               title = "Full sample regression results", 
+               label = "fullResults",
+               header = F,
+               keep = c("floods", "droughts"))
+```
+
 ```{r regression rural sample}
+
+rural <- subset(reg_data, urb_rur == 2, select = c("hh_id", "id", "age", "sex", "literacy", "dept", "urb_rur", "income", "assist", "register", "floods", "droughts", "year", "job", "tasks", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor"))
+
+model_assist_r <- lm(assist ~ floods + droughts + age + sex + income + hhhead_educ + female_hhhead + urb_rur + year, data = reg_data)
+summary(model_assist_r)
+model_register_r <- lm(register ~ floods + droughts + age + sex + income + hhhead_educ + female_hhhead + urb_rur + year, data = reg_data)
+summary(model_register_r)
 

### 2020-04-24 data_appendix_all_years.Rmd
commit

@@
 total_subset %>%
-  mutate(ln_income = log(income))
+  mutate(ln_income = log(income),
+         literate = literacy-1,
+         female = sex-1,
+         no_assist = assist-1,
+         not_registered = register-1)
 reg_data <- total_subset %>%

### 2020-04-24 data_appendix_all_years.Rmd
commit

--- 2020-04-24 data_appendix_all_years.Rmd
@@
 total_subset %>%
-  mutate(ln_income = log(income))
+  mutate(ln_income = log(income),
+         literate = literacy-1,
+         female = sex-1,
+         no_assist = assist-1,
+         not_registered = register-1)
+#trying to make the assist, register, sex, and literacy variables into 0 1 binary variables, with assist and register = 1, sex = 1 if female, and literacy = 1 if literate.
+The current coding from the original data is all in 1 2 format, with the opposites.
+Trying to change it by subtracting doesn't seem to be working.
 reg_data <- total_subset %>%

### 2020-04-25 data_appendix_all_years.Rmd
Fixing a typo and replacing the not working code with female with working code. Problem appears to be related to the "haven labelled" data category.

--- 2020-04-24 data_appendix_all_years.Rmd
@@
 all_years <- hh_2005 %>% 
-  rbind(hh_2005, hh_2006, hh_2007, hh_2011, hh_2012) %>%
+  rbind(hh_2005, hh_2006, hh_2007, hh_2011, hh_2012)
 
@@
 ```{r regression variable specifications}
-total_subset %>%
-  mutate(ln_income = log(income))
 reg_data <- total_subset %>%
-  mutate(year_factor = as.factor(year))
+  mutate(ln_income = log(income),
+         literate = literacy - 1,
+         female = case_when(sex == 2 ~ 1,
+                            sex == 1 ~ 0),
+         no_assist = assist - 1,
+         not_registered = register - 1,
+         year_factor = as.factor(year))
 

### 2020-04-26 data_appendix_all_years.Rmd
commit spacing
another subset of regressions

@@
   rename(hh_id = folio,
-        assist = s3_08,  
+        assist1 = s3_08,  
         reason = s3_09, 
@@
         relation = s1_05,
-        literacy = s3_01,
-        register = s3_04,
+        literacy1 = s3_01,
+        register1 = s3_04,
         school_level = s3_05a,
@@
 hh_2005 <- data_2005 %>% 
-  select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
+  select(hh_id, assist1, reason, sex, age,
+         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
   mutate(year = 2005,
@@
                             reason == 9 ~ 'housework/childcare',
-                            reason == 10 ~ 'other'))
+                            reason == 10 ~ 'other'),
+          literacy = case_when(literacy1 == 1 ~ 1,
+                              literacy1 == 2 ~ 0), 
+         female = case_when(sex == 2 ~ 1,
+                            sex == 1 ~ 0),
+         rural = case_when(urb_rur == 1 ~ 0,
+                           urb_rur == 2 ~ 1), 
+         assist = case_when(assist1 == 1 ~ 1,
+                            assist1 == 2 ~ 0),
+         register = case_when(register1 == 1 ~ 1,
+                              register1 == 2 ~ 0),
+         primary = case_when(school_level == 17 ~ 1,
+                             school_level != 17 ~ 0),
+         secondary = case_when(school_level == 18 ~ 1,
+                               school_level != 18 ~ 0))
 
@@
   rename(hh_id = Folio,
-        assist = s4_08,  
+        assist1 = s4_08,  
         reason = S4_09, 
@@
         relation = s1_05,
-        literacy = s4_01,
-        register = s4_04,
+        literacy1 = s4_01,
+        register1 = s4_04,
         school_level = s4_05a,
@@
 hh_2006 <- data_2006 %>% 
-  select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
+  select(hh_id, assist1, reason, sex, age,
+         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
   mutate(year = 2006,
@@
                             reason == 9 ~ 'housework/childcare',
-                            reason == 10 ~ 'other'))
+                            reason == 10 ~ 'other'),
+          literacy = case_when(literacy1 == 1 ~ 1,
+                              literacy1 == 2 ~ 0), 
+         female = case_when(sex == 2 ~ 1,
+                            sex == 1 ~ 0),
+         rural = case_when(urb_rur == 1 ~ 0,
+                           urb_rur == 2 ~ 1), 
+         assist = case_when(assist1 == 1 ~ 1,
+                            assist1 == 2 ~ 0),
+         register = case_when(register1 == 1 ~ 1,
+                              register1 == 2 ~ 0),
+         primary = case_when(school_level == 17 ~ 1,
+                             school_level != 17 ~ 0),
+         secondary = case_when(school_level == 18 ~ 1,
+                               school_level != 18 ~ 0))
 ```
@@
   rename(hh_id = folio,
-        assist = s4_10,  
+        assist1 = s4_10,  
         reason = s4_11, 
@@
         relation = s1_06,
-        literacy = s4_01,
-        register = s4_06,
+        literacy1 = s4_01,
+        register1 = s4_06,
         school_level = s4_07a,
@@
 hh_2007 <- data_2007 %>% 
-  select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
+  select(hh_id, assist1, reason, sex, age,
+         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
   mutate(year = 2007,
@@
   rename(hh_id = folio,
-        assist = s4_12,  
+        assist1 = s4_12,  
         reason = s4_13,
@@
         relation = s1_08,
-        literacy = s4_01,
-        register = s4_05,
+        literacy1 = s4_01,
+        register1 = s4_05,
         grade = s4_06b,
@@
 hh_2011 <- data_2011 %>% 
-  select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
+  select(hh_id, assist1, reason, sex, age,
+         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
   mutate(year = 2011,
@@
                             reason == 9 ~ 'housework/childcare',
-                            reason == 10 ~ 'other'))
+                            reason == 10 ~ 'other'),
+          literacy = case_when(literacy1 == 1 ~ 1,
+                               literacy1 == 2 ~ 0), 
+          female = case_when(sex == 2 ~ 1,
+                             sex == 1 ~ 0),
+           rural = case_when(urb_rur == 1 ~ 0,
+                             urb_rur == 2 ~ 1), 
+           assist = case_when(assist1 == 1 ~ 1,
+                              assist1 == 2 ~ 0),
+           register = case_when(register1 == 1 ~ 1,
+                                register1 == 2 ~ 0),
+         primary = case_when(school_level == 16 ~ 1,
+                             school_level != 16 ~ 0),
+         secondary = case_when(school_level == 17 ~ 1,
+                               school_level != 17 ~ 0))
 ```
@@
         relation = s1_08,
-        literacy = s4_01,
-        register = s4_04,
+        literacy1 = s4_01,
+        register1 = s4_04,
         grade = s4_05b,
@@
 hh_2012 <- data_2012 %>% 
-  select(hh_id, assist, reason, sex, age,
-         relation, literacy, register, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
+  select(hh_id, assist1, reason, sex, age,
+         relation, literacy1, register1, school_level, grade, repeated, depto, urb_rur, income, educ, job, tasks, pov_line, extr_pov) %>% 
   mutate(year = 2012,
@@
                             reason == 9 ~ 'housework/childcare',
-                            reason == 10 ~ 'other'))
+                            reason == 10 ~ 'other'),
+         literacy = case_when(literacy1 == 1 ~ 1,
+                              literacy1 == 2 ~ 0), 
+         female = case_when(sex == 2 ~ 1,
+                            sex == 1 ~ 0),
+         rural = case_when(urb_rur == 1 ~ 0,
+                           urb_rur == 2 ~ 1), 
+         assist = case_when(assist1 == 1 ~ 1,
+                            assist1 == 2 ~ 0),
+         register = case_when(register1 == 1 ~ 1,
+                              register1 == 2 ~ 0),
+         primary = case_when(school_level == 19 ~ 1,
+                             school_level != 19 ~ 0),
+         secondary = case_when(school_level == 20 ~ 1,
+                               school_level != 20 ~ 0))
+
 ```
@@
   left_join(flood_summary) %>%
-  left_join(drought_summary)
+  left_join(drought_summary) %>%
+  mutate(child = case_when(relation == 3 ~ 1,
+                           relation != 3 ~ 0),
+         grandchild = case_when(relation == 8 ~ 1,
+                                relation != 8 ~ 0),
+         pov_diff = income - pov_line,
+         pov = case_when(pov_diff > 0 ~ 0,
+                         pov_diff < 0 ~ 1),
+         extr_diff = income - extr_pov,
+         extr = case_when (extr_diff > 0 ~ 0,
+                           extr_diff < 0 ~ 1),
+                  work_reason = case_when(reason == "work" ~ 1,
+                                 reason != "work" ~ 0),
+         money_reason = case_when(reason == "money" ~ 1, 
+                                  reason != "money" ~ 0),
+         lny = log(income))
+
+total_data %>%
+  group_by(work_reason, money_reason) %>%
+  summarise(n = n())
 
@@
 
-hh_head <- subset(total_data, relation == 1, select = c("hh_id", "sex", "age", "literacy", "dept", "urb_rur", "income", "year", "educ", "job", "tasks", "pov_line", "extr_pov"))
+hh_head <- subset(total_data, relation == 1, select = c("hh_id", "female", "age", "literacy", "dept", "rural", "income", "year", "educ", "job", "tasks", "pov_diff", "extr_diff", "pov", "extr"))
 
@@
 
-children <- subset(total_data, age < 18 & age > 4, select = c("hh_id", "age", "sex", "relation", "literacy", "dept", "urb_rur", "income", "assist", "reason", "register", "school_level", "grade", "repeated", "floods", "droughts", "year", "job", "tasks"))
+children <- subset(total_data, age < 18 & age > 4, select = c("hh_id", "age", "female", "child", "grandchild", "literacy", "dept", "rural", "income", "assist", "reason", "register", "primary", "secondary", "floods", "droughts", "year", "job", "tasks", "money_reason", "work_reason"))
 
@@
 
+total_children %>%
+      group_by(child, grandchild) %>%
+  summarise(n = n())
+
+total_children %>%
+  group_by(work_reason, money_reason) %>%
+  summarise(n = n())
+
 ```
@@
 hh_head_2 <- hh_head %>%
-  select(sex, age, literacy, educ, job, tasks, pov_line, extr_pov, hh_id) %>%
-  rename (female_hhhead = sex,
+  select(female, age, literacy, educ, job, tasks, pov, extr, hh_id) %>%
+  rename (female_hhhead = female,
           hhhead_age = age,
@@
 ```{r regression variable specifications}
+
 reg_data <- total_subset %>%
-  mutate(ln_income = log(income),
-         literate = literacy - 1,
-         female = case_when(sex == 2 ~ 1,
-                            sex == 1 ~ 0),
-         no_assist = assist - 1,
-         not_registered = register - 1,
-         year_factor = as.factor(year))
+  mutate(year_factor = as.factor(year))
 
@@
 #summary(model_simple)
-#model_assist <- lm(assist ~ floods + droughts + age + sex + income + hhhead_educ + female_hhhead + year, data = reg_data)
-#summary(model_assist)
-model_register <- lm(register ~ floods + droughts + age + sex + income + hhhead_educ + female_hhhead + year, data = reg_data)
-summary(model_register)
+model_assist <- lm(assist ~ floods + droughts + age + female + income + child + grandchild + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
+summary(model_assist)
+#model_register <- lm(register ~ floods + droughts + age + female + income + child + grandchild + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
+#summary(model_register)
 
@@
 ```{r generate regression table, echo = F, results = "asis"}
-stargazer::stargazer(list(model_assist, model_register), 
+stargazer::stargazer(list(model_assist_r, model_register_r), 
                title = "Full sample regression results", 
@@
 
-rural <- subset(reg_data, urb_rur == 2, select = c("hh_id", "id", "age", "sex", "literacy", "dept", "urb_rur", "income", "assist", "register", "floods", "droughts", "year", "job", "tasks", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor"))
+rural <- subset(reg_data, rural == 1, select = c("hh_id", "id", "age", "female", "literacy", "dept", "rural", "income", "assist", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor"))
 
-model_assist_r <- lm(assist ~ floods + droughts + age + sex + income + hhhead_educ + female_hhhead + urb_rur + year, data = reg_data)
+model_assist_r <- (lm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data))
 summary(model_assist_r)
-model_register_r <- lm(register ~ floods + droughts + age + sex + income + hhhead_educ + female_hhhead + urb_rur + year, data = reg_data)
+model_register_r <- lm(register ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data)
 summary(model_register_r)
+model_reason <- (lm(work_reason ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data))
+summary(model_reason)
 
@@
 
+urban <- subset(reg_data, rural == 0, select = c("hh_id", "id", "age", "female", "literacy", "dept", "rural", "income", "assist", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor"))
+
+model_assist_u <- lm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data)
+summary(model_assist_u)
+model_register_u <- lm(register ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data)
+summary(model_register_u)
 ```

### 2020-04-27 data_appendix_all_years.Rmd
commit

@@
 
+export_summary_table(dfSummary(total_children))
+
 total_children %>%
@@
           hhhead_tasks = tasks)
+
 total_heads_subset <- hh_head_2 %>% distinct()
@@
 
+reg_data %>%
+  group_by(rural) %>%
+  summarise(n = n())
+
 ```
@@
 #summary(model_simple)
-model_assist <- lm(assist ~ floods + droughts + age + female + income + child + grandchild + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
+model_assist <- lm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
 summary(model_assist)
-#model_register <- lm(register ~ floods + droughts + age + female + income + child + grandchild + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
-#summary(model_register)
+model_register <- lm(register ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
+summary(model_register)
+model_reason_t <- (lm(work_reason ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data))
+summary(model_reason_t)
 

### 2020-04-30 data_appendix_all_years.Rmd
added probit regression

@@
 #summary(model_simple)
-model_assist <- lm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
-summary(model_assist)
-model_register <- lm(register ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
-summary(model_register)
+#model_assist <- lm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
+#summary(model_assist)
+#model_register <- lm(register ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
+#summary(model_register)
 model_reason_t <- (lm(work_reason ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data))
@@
 ```
+```{r probit regression total sample}
+
+probitMod_assist <- glm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data, family=binomial(link="probit"))  # build the logit model
+
+summary(probitMod_assist)  # model summary
+```
 
@@
 
-rural <- subset(reg_data, rural == 1, select = c("hh_id", "id", "age", "female", "literacy", "dept", "rural", "income", "assist", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor"))
+rural <- subset(reg_data, rural == 1, select = c("hh_id", "id", "age", "female", "literacy", "dept", "rural", "income", "assist", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason"))
 
-model_assist_r <- (lm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data))
+model_assist_r <- (lm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = rural))
 summary(model_assist_r)
-model_register_r <- lm(register ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data)
+model_register_r <- lm(register ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = rural)
 summary(model_register_r)
-model_reason <- (lm(work_reason ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data))
+model_reason <- (lm(work_reason ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = rural))
 summary(model_reason)

### 2020-04-30 data_appendix_all_years.Rmd
made a subset without the vacation kids

@@
          money_reason = case_when(reason == "money" ~ 1, 
-                                  reason != "money" ~ 0),
-         lny = log(income))
+                                  reason != "money" ~ 0))
 
@@
 #left join command creating the duplicates problem it seems 
-total <- merge(total_children,hh_head_2, by="hh_id")
+total <- merge(total_children,total_heads_subset, by="hh_id")
 
-total_subset <- total %>% distinct()
+#total_2 <- total[!(total$reason == "vacation" & total$assist == 0]
+
+total %>%
+  group_by(reason) %>%
+  summarise(n = n())
+
+total_2 <- subset(total, !(reason == "vacation" & assist == 0)) #taking out the observations with vacation kids
+
+total_subset <- total_2 %>% distinct()
 
@@
 reg_data <- total_subset %>%
-  mutate(year_factor = as.factor(year))
+  mutate(year_factor = as.factor(year),
+         lny = log(income))
 
@@
 
-probitMod_assist <- glm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data, family=binomial(link="probit"))  # build the logit model
+probitMod_assist <- glm(assist ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data, family=binomial(link="probit"))
+
+probitMod_register <- glm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data, family=binomial(link="probit"))
+
+probitMod_work <- glm(work_reason ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data, family=binomial(link="probit"))
 
 summary(probitMod_assist)  # model summary
+summary(probitMod_register)
+summary(probitMod_work)
 ```

### 2020-05-01 data_appendix_all_years.Rmd
commiting final changes, intermediate draft to go back to

@@
+---
+output:
+  pdf_document: default
+  html_document: default
+---
+
 title: "Data Appendix to "The impact of weather events on schooling outcomes""
@@
   mutate(year_factor = as.factor(year),
-         lny = log(income))
+         income2 = case_when(income == 0 ~ 1), 
+         lny = log(income2))
 
@@
 #summary(model_simple)
-#model_assist <- lm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
-#summary(model_assist)
-#model_register <- lm(register ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
-#summary(model_register)
-model_reason_t <- (lm(work_reason ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data))
+model_assist <- lm(assist ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
+summary(model_assist)
+model_register <- lm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
+summary(model_register)
+model_reason_t <- (lm(work_reason ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data))
 summary(model_reason_t)
@@
 ```{r generate regression table, echo = F, results = "asis"}
-stargazer::stargazer(list(model_assist_r, model_register_r), 
+stargazer::stargazer(list(model_assist, model_register), 
                title = "Full sample regression results", 
@@
 summary(probitMod_work)
+```
+
+```{r generate probit regression table, echo = F, results = "asis"}
+stargazer::stargazer(list(probitMod_assist, probitMod_register, probitMod_work), 
+               title = "Full sample probit regression results", 
+               label = "fullResults",
+               header = F,
+               keep = c("floods", "droughts"))
+
 ```

### 2020-05-01 paper_draft.Rmd
paper coding draft separate from the data appendix

--- 2020-05-01 data_appendix_all_years.Rmd
@@
 
-title: "Data Appendix to "The impact of weather events on schooling outcomes""
+title: ""The impact of environmental shocks on children's schooling in Bolivia""
 author: "Katya Garcia-Israel"
@@
 ```
-# Appendix description
-This data appendix details the steps of collecting, processing, and merging the data needed to create the complete data set showing information about schooling and weather events.
-It concludes with a discussion of patterns found in the data.
+# Introduction
 
-The datasets used directly by the final analysis are saved in `processed-data/` at the end of this file.
+# Literature Review
 
-# Raw data
-This section documents the datasets used in this analysis.
+# Data and Methods
 
-## Household Survey data
-**Citation:** .  
+## Data
+
+The household data was taken from the social surveys database of the Bolivia National Statistical Institute, using household surveys from the years 2005-2007, 2011-2012.
 Instituto Nacional de Estadistica. (2005). Encuesta de Hogares (2005) [SAV file]. Retrieved from https://www.ine.gob.bo/index.php/banco/base-de-datos-sociales
 
-**Date Downloaded:** Downloaded September 2019.
-**Filename(s):** raw_data/filename.csv 
-**Unit of observation:** Unit of observation at the individual level, but each individual is identified mainly by the household id. 
-**Dates covered:** 2005
-
-### To obtain a copy
-
-This household survey can be found on the Bolivian National Statistical Institute website, under the menu for 'Banco de Datos' and 'Base de datos de encuestas sociales.' Then the household surveys may be found for the years 2005-2018, excepting 2010.
-
-The original format of this data was from the sav file, but I imported it into Stata before importing the Stata format into R.
+The unit of observation at the individual level, but each individual is identified mainly by the household id.
 
@@
 
-### Data import code and summary
-
-```{r import household data 2005}
+```{r import household data 2005, echo = F, message = F}
 data_2005 <- read_dta("Raw data/seminar_paper_data/2005_hhsurvey.dta") %>% 
@@
 
-```{r export summary table 2005}
+```{r export summary table 2005, echo = F, message = F}
 export_summary_table(dfSummary(hh_2005))
@@
 
-```{r import household data 2006}
+```{r import household data 2006, echo = F, message = F}
 data_2006 <- read_dta("Raw data/seminar_paper_data/2006_hhsurvey.dta")  %>% 
@@
 
-```{r exporting data table}
+```{r exporting data table, echo = F, message = F}
 export_summary_table(dfSummary(hh_2006))
@@
 
-```{r import household data 2007}
+```{r import household data 2007, echo = F, message = F}
 data_2007 <- read_dta("Raw data/seminar_paper_data/2007_hhsurvey.dta")  %>% 
@@
 
-```{r import household data 2011}
+```{r import household data 2011, echo = F, message = F}
 data_2011 <- read_dta("Raw data/seminar_paper_data/2011_hhsurvey.dta")  %>% 
@@
 
-```{r household data 2012}
+```{r household data 2012, echo = F, message = F}
 data_2012 <- read_dta("Raw data/seminar_paper_data/2012_hhsurvey.dta")  %>% 
@@
 
-```{r combine household datasets}
+```{r combine household datasets, echo = F, message = F}
 all_years <- hh_2005 %>% 
@@
 
-## Flood data
-**Citation:** GeoSinager. (2016). Mapa Recurrencias de Inundaciones Gestiones 2002-2012. [CSV file]. Retrieved from http://geosinager.defensacivil.gob.bo/maps/226
-
-**Date Downloaded:** October 2019
-**Filename(s):** raw_data/filename.csv 
-**Unit of observation:** Municipality  
-**Dates covered:** 2002-2012
-
-### To obtain a copy
+The flood data is from the years 2005-2012.
+The unit of observation is municipality, but the data is later combined to make the unit of observation at the department level to make it comparable with the household survey data.
 
@@
 
-### Variable descriptions
-
-- **dept:** department name 
-- **A_2005:** how many floods occurred in 2005 in that location 
-- **A_2006:** how many floods occurred in 2006 in that location 
-- **A_2007:** how many floods occurred in 2007 in that location 
-- **A_2008:** how many floods occurred in 2008 in that location 
-- **A_2009:** how many floods occurred in 2009 in that location 
-- **A_2010:** how many floods occurred in 2010 in that location 
-- **A_2011:** how many floods occurred in 2011 in that location 
-- **A_2012:** how many floods occurred in 2012 in that location
-
-### Data import code and summary
-```{r import flood data}
+```{r import flood data, echo = F, message = F}
 flood <- read_csv(file.path("Raw data", "recurrencia_inundacion2002_2012bol.csv")) %>% 
@@
 ```
-## Drought data
-**Citation:** GeoSinager. (2016). Mapa Recurrencias de Sequias Gestiones 2002-2012. [CSV file]. Retrieved from http://geosinager.defensacivil.gob.bo/maps/225
-
-**Date Downloaded:** October 2019.  
-**Filename(s):** Raw data/droughts.csv 
-**Unit of observation:** The unit of observation is the municipality 
-**Dates covered:** 2002-2012
-
-### To obtain a copy
+The drought data also spans the years 2005-2012 and shows the frequency of droughts in each municipality by year.
 
@@
 It is shown on a map with different layers, and the raw data may be downloaded in a variety of different forms, including csv.
-
-### Variable descriptions
-
-- **dept:** department name 
-- **A_2005:** how many droughts occurred in 2005 in that location 
-- **A_2006:** how many droughts occurred in 2006 in that location 
-- **A_2007:** how many droughts occurred in 2007 in that location 
-- **A_2008:** how many droughts occurred in 2008 in that location 
-- **A_2009:** how many droughts occurred in 2009 in that location 
-- **A_2010:** how many droughts occurred in 2010 in that location 
-- **A_2011:** how many droughts occurred in 2011 in that location 
-- **A_2012:** how many droughts occurred in 2012 in that location
 
@@
 
-``` {r merging datasets}
+``` {r merging datasets, echo = F, message = F}
 total_data<- all_years %>% 
@@
 
-# Analysis Variables
+## Analysis Variables
 
@@
 
-```{r creating head of household subset}
+```{r creating head of household subset, echo = F, message = F}
 
@@
 
-```{r creating children subset}
+```{r creating children subset, echo = F, message = F}
 
@@
 
-```{r combining household head characteristics with children dataset}
+```{r combining household head characteristics with children dataset, echo = F, message = F}
 hh_head_2 <- hh_head %>%
@@
 ```
+```{r summary statistic tables with total_subset data, echo = F, message = F}
+#creating a table with the main explanatory variables
 
-{r creating summary statistics}
-export_summary_table(dfSummary(total_subset))
+#creating a table dividing the characteristics by urban/rural area
+```
+#Empirical Model
 
-install.packages("devtools")
-devtools::install_github("ewenharrison/finalfit")
-library(finalfit)
-library(dplyr)
-
-```{r regression variable specifications}
+```{r regression variable specifications, echo = F, message = F}
 
@@
   mutate(year_factor = as.factor(year),
+         dept_factor = as.factor(dept),
          income2 = case_when(income == 0 ~ 1), 
@@
 
-```{r regressions total sample}
+```{r linear regressions total sample, echo = F, message = F}
 
@@
 #summary(model_simple)
-model_assist <- lm(assist ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
+model_assist <- lm(assist ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + dept_factor + year_factor, data = reg_data)
 summary(model_assist)
-model_register <- lm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
+model_register <- lm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + dept_factor + year_factor, data = reg_data)
 summary(model_register)
-model_reason_t <- (lm(work_reason ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data))
+model_reason_t <- (lm(work_reason ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data))
 summary(model_reason_t)
@@
 ```
-```{r probit regression total sample}
+```{r probit regression total sample, echo = F, message = F}
 
-probitMod_assist <- glm(assist ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_assist <- glm(assist ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
-probitMod_register <- glm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_register <- glm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
-probitMod_work <- glm(work_reason ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_work <- glm(work_reason ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
@@
 
-```{r regression rural sample}
+```{r regression rural sample, echo = F, message = F}
 
-rural <- subset(reg_data, rural == 1, select = c("hh_id", "id", "age", "female", "literacy", "dept", "rural", "income", "assist", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason"))
+rural <- subset(reg_data, rural == 1, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "income", "assist", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason"))
 
-model_assist_r <- (lm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = rural))
+model_assist_r <- (lm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
 summary(model_assist_r)
-model_register_r <- lm(register ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = rural)
+model_register_r <- lm(register ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural)
 summary(model_register_r)
-model_reason <- (lm(work_reason ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = rural))
+model_reason <- (lm(work_reason ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
 summary(model_reason)
@@
 
-```{r regression urban sample}
+```{r linear regression urban sample, echo = F, message = F}
 
-urban <- subset(reg_data, rural == 0, select = c("hh_id", "id", "age", "female", "literacy", "dept", "rural", "income", "assist", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor"))
+urban <- subset(reg_data, rural == 0, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "income", "assist", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor"))
 
-model_assist_u <- lm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data)
+model_assist_u <- lm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
 summary(model_assist_u)
-model_register_u <- lm(register ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data)
+model_register_u <- lm(register ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
 summary(model_register_u)

### 2020-05-01 paper_draft.Rmd
added other probit regressions

@@
 
+```{r generate linear regression table rural sample, echo = F, results = "asis"}
+stargazer::stargazer(list(model_assist, model_register), 
+               title = "Rural sample linear regression results", 
+               label = "fullResults",
+               header = F,
+               keep = c("floods", "droughts"))
+```
+
+```{r probit regression rural sample, echo = F, message = F}
+
+probitMod_assist_r <- glm(assist ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+
+probitMod_register_r <- glm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+
+probitMod_work_r <- glm(work_reason ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+
+summary(probitMod_assist_r)  # model summary
+summary(probitMod_register_r)
+summary(probitMod_work_r)
+```
+
 ```{r linear regression urban sample, echo = F, message = F}
@@
 summary(model_register_u)
+```
+
+```{r linear regression generate table urban sample, echo = F, results = "asis"}
+stargazer::stargazer(list(model_assist, model_register), 
+               title = "Urban sample linear regression results", 
+               label = "urbanlinearResults",
+               header = F,
+               keep = c("floods", "droughts"))
+
+```
+
+```{r probit regression urban sample, echo = F, message = F}
+
+probitMod_assist_u <- glm(assist ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+
+probitMod_register_u <- glm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+
+probitMod_work_u <- glm(work_reason ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+
+summary(probitMod_assist_r)  # model summary
+summary(probitMod_register_r)
+summary(probitMod_work_r)
 ```

### 2020-05-06 paper_draft.Rmd
natural log specifications line 571

@@
 
-children <- subset(total_data, age < 18 & age > 4, select = c("hh_id", "age", "female", "child", "grandchild", "literacy", "dept", "rural", "income", "assist", "reason", "register", "primary", "secondary", "floods", "droughts", "year", "job", "tasks", "money_reason", "work_reason"))
+children <- subset(total_data, age < 18 & age > 4, select = c("hh_id", "age", "female", "child", "grandchild", "literacy", "dept", "rural", "assist", "reason", "register", "primary", "secondary", "floods", "droughts", "year", "job", "tasks", "money_reason", "work_reason"))
 
@@
 hh_head_2 <- hh_head %>%
-  select(female, age, literacy, educ, job, tasks, pov, extr, hh_id) %>%
+  select(female, age, literacy, educ, job, tasks, income, pov, extr, hh_id) %>%
   rename (female_hhhead = female,
@@
          dept_factor = as.factor(dept),
-         income2 = case_when(income == 0 ~ 1), 
+         income2 = case_when(income == 0.0000 ~ 1.0000), 
          lny = log(income2))
@@
 
-probitMod_assist <- glm(assist ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_assist <- glm(assist ~ floods + droughts + age + female + lny + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
-probitMod_register <- glm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_register <- glm(register ~ floods + droughts + age + female + lny + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
-probitMod_work <- glm(work_reason ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_work <- glm(work_reason ~ floods + droughts + age + female + lny + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
@@
 
-rural <- subset(reg_data, rural == 1, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "income", "assist", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason"))
+rural <- subset(reg_data, rural == 1, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "income", "assist", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason", "primary"))
 
-model_assist_r <- (lm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
+model_assist_r <- (lm(assist ~ floods + droughts + age + female + income + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
 summary(model_assist_r)
-model_register_r <- lm(register ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural)
+model_register_r <- lm(register ~ floods + droughts + age + female + income + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural)
 summary(model_register_r)
-model_reason <- (lm(work_reason ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
+model_reason <- (lm(work_reason ~ floods + droughts + age + female + income + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
 summary(model_reason)

### 2020-05-06 paper_draft.Rmd
adding the lny variable that works

@@
 #Empirical Model
+see the link for math equation help.
 
@@
          dept_factor = as.factor(dept),
-         income2 = case_when(income == 0.0000 ~ 1.0000), 
-         lny = log(income2))
+         lny = log(income + 1),
+         valley = case_when(dept == "Chuquisaca", "Cochabamba", "Tarija" ~ 1,
+                            dept != "Chuquisaca", "Cochabamba", "Tarija" ~ 0))
+
+#adding the region variables for the region fixed-effects option
 
 reg_data %>%
-  group_by(rural) %>%
+  group_by(dept_factor) %>%
   summarise(n = n())

### 2020-05-06 paper_draft.Rmd
working on the region variable code for valleys, line 592

@@
 ```
-```{r summary statistic tables with total_subset data, echo = F, message = F}
+
+```{r summary stats example}
+install.packages("stargazer")
+library(stargazer)
+
+stargazer(total_subset)
+
+```
+
+```{r create summary statistic table, echo = F, results = 'asis'}
 #creating a table with the main explanatory variables
+
+load("processed-data/analysis_data.RData")
+
+stargazer::stargazer(as.data.frame(stat_data),
+                     label = "summaryStats",
+                     title = "Summary statistics",
+                     header = F,
+                     type = "latex",
+                     digits = 0)
 
@@
 ```
+\begin{table}
+\centering
+\caption{Variable definitions}
+\label{variableDefinitions}
+\begin{tabular}{ll}
+\hline
+\hline
+Variable name      & Description                                   \\
+\hline
+svr\_mean          & Monthly mean of SVR at a park                 \\
+svr\_median        & Monthly median of SVR at a park               \\
+svr\_p10           & Monthly 10th percentile of SVR at a park      \\
+svr\_p25           & Monthly 25th percentile of SVR at a park      \\
+svr\_p75           & Monthly 75th percentile of SVR at a park      \\
+svr\_p90           & Monthly 90th percentile of SVR at a park      \\
+recreation\_visits & Monthly number of recreation visits at a park \\
+\hline
+\end{tabular}
+\end{table}
 

### 2020-05-06 paper_draft.Rmd
working on the variables table

@@
          lny = log(income + 1),
-         valley = case_when(dept == "Chuquisaca", "Cochabamba", "Tarija" ~ 1,
-                            dept != "Chuquisaca", "Cochabamba", "Tarija" ~ 0))
+         valley = case_when(dept %in% c("Chuquisaca", "Cochahamba", "Tarija") ~ 1, TRUE ~ 0),
+         mountain = case_when(dept %in% c("La Paz", "Oruro", "Potosi") ~ 1, TRUE ~ 0),
+         plains = case_when(dept %in% c("Beni", "Pando", "Santa Cruz") ~ 1, TRUE ~ 0),
+         lagged_flood =
 
@@
 \hline
-svr\_mean          & Monthly mean of SVR at a park                 \\
-svr\_median        & Monthly median of SVR at a park               \\
-svr\_p10           & Monthly 10th percentile of SVR at a park      \\
-svr\_p25           & Monthly 25th percentile of SVR at a park      \\
-svr\_p75           & Monthly 75th percentile of SVR at a park      \\
-svr\_p90           & Monthly 90th percentile of SVR at a park      \\
-recreation\_visits & Monthly number of recreation visits at a park \\
+Attend          & Binary variable, 1 = attends school                          \\
+Register        & Binary variable, 1 = registered for school                   \\
+Work_reason     & Binary variable, 1 = does not attend school for work reasons \\
+Lny             & Log of income per capita (in Bolivianos/month)
+Female          & Binary variable, 1 = female      \\
+Age             & Age in years       \\
+Child           & 
+Grandchild
+Num_children
+Secondary
+hhhead_educ
+female_hhhead
+Floods          & Number of flood occurrences per dept per year      \\
+Droughts        & Number of drought occurrences per dept per year \\
+Dept
+Valleys
+Mountains
+Plains
+
 \hline

### 2020-05-07 paper_draft.Rmd
added interaction terms

--- 2020-05-06 paper_draft.Rmd
@@
 
-### Variable descriptions
-
-- **hh_id:** The household ID, which is the same for each member of the household. It is made up of letters and numbers. 
-- **nro1:** The number of the person in each household.
-- **assist:** Binary variable, whether or not the individual assists school 
-- **urb_rur:** Binary variable, whether the household is in an urban or rural area
-- **reason:** the reason why the individual is not assisting school if they are not 
-- **members:** number of household members excluding servants 
-- **sex:** the sex of the individual, Male = 1, Female =2 
-- **age:** How old the individual is, in years 
-- **relation:** The relationship of the individual to the head of household 
-- **literacy:** Binary variable, whether a person can read and write 
-- **register:** Binary variable, whether an individual registered for school or not 
-- **grade:** For what grade is the person registered 
-- **repeated:** whether the person has repeated the grade they are currently registered in 
-- **dept:** the department in which the person lives 
-- **income:** the per capita monthly income for each household, in bolivianos
-
 ```{r import household data 2005, echo = F, message = F}
@@
                            urb_rur == 2 ~ 1), 
-         assist = case_when(assist1 == 1 ~ 1,
+         attend = case_when(assist1 == 1 ~ 1,
                             assist1 == 2 ~ 0),
@@
                              urb_rur == 2 ~ 1), 
-           assist = case_when(assist1 == 1 ~ 1,
+           attend = case_when(assist1 == 1 ~ 1,
                               assist1 == 2 ~ 0),
@@
 
-# Data Processing and Combination
-
-To create a complete dataset, I merged the household data with the floods data and the droughts data.
-Before merging the data, I cut down the datasets to include only the relevant variables needed for analysis in each file.
-I renamed several variables in the household survey data set to make them more understandable.
-I also created the year variable for the household data to ensure that it matched up correctly with the weather data.
-For the flood and drought data, I grouped the occurrences by department instead of by municipality so they would match the household data, and I made the department character numeric for the merge.
-
 ``` {r merging datasets, echo = F, message = F}
@@
 
-total_data %>%
-  group_by(work_reason, money_reason) %>%
-  summarise(n = n())
+#total_data %>%
+  #group_by(work_reason, money_reason) %>%
+  #summarise(n = n())
 
@@
 
-children <- subset(total_data, age < 18 & age > 4, select = c("hh_id", "age", "female", "child", "grandchild", "literacy", "dept", "rural", "assist", "reason", "register", "primary", "secondary", "floods", "droughts", "year", "job", "tasks", "money_reason", "work_reason"))
+children <- subset(total_data, age < 18 & age > 4, select = c("hh_id", "age", "female", "child", "grandchild", "literacy", "dept", "rural", "attend", "reason", "register", "primary", "secondary", "floods", "droughts", "year", "job", "tasks", "money_reason", "work_reason"))
 
@@
 
-#total_2 <- total[!(total$reason == "vacation" & total$assist == 0]
-
 total %>%
@@
 
-total_2 <- subset(total, !(reason == "vacation" & assist == 0)) #taking out the observations with vacation kids
+total_2 <- subset(total, !(reason == "vacation" & attend == 0)) #taking out the observations with vacation kids
 
@@
 ```
-```{r summary statistic tables with total_subset data, echo = F, message = F}
-#creating a table with the main explanatory variables
-
-#creating a table dividing the characteristics by urban/rural area
-```
-#Empirical Model
-see the link for math equation help.
 
@@
          lny = log(income + 1),
-         valley = case_when(dept == "Chuquisaca", "Cochabamba", "Tarija" ~ 1,
-                            dept != "Chuquisaca", "Cochabamba", "Tarija" ~ 0))
+         valleys = case_when(dept %in% c("Chuquisaca", "Cochahamba", "Tarija") ~ 1, TRUE ~ 0),
+         mountains = case_when(dept %in% c("La Paz", "Oruro", "Potosi") ~ 1, TRUE ~ 0),
+         plains = case_when(dept %in% c("Beni", "Pando", "Santa Cruz") ~ 1, TRUE ~ 0),
+         flood_secondary = floods * secondary,
+         drought_secondary = droughts * secondary,
+         flood_pov = floods * extr,
+         drought_pov = droughts * extr)
 
 #adding the region variables for the region fixed-effects option
+#adding interaction terms
+#adding lagged flood and drought variables
+
+#creating rural sample
+rural <- subset(reg_data, rural == 1, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "lny", "attend", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason", "secondary", "valleys", "mountains", "plains", "dept", "extr", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov"))
 
@@
   group_by(dept_factor) %>%
-  summarise(n = n())
+  #summarise(n = n())
 
 ```
+##Variable Descriptions
+
+\begin{table}
+\centering
+\caption{Variable definitions}
+\label{variableDefinitions}
+\begin{tabular}{ll}
+\hline
+\hline
+Variable name      & Description                                               \\
+\hline
+Attend          & Binary variable, 1 = attends school                          \\
+Register        & Binary variable, 1 = registered for school                   \\
+Work reason     & Binary variable, 1 = missing school for work                 \\
+Lny             & Log of income per capita (in Bolivianos/month)               \\
+Extr            & Binary variable, 1 = below extreme poverty line              \\
+Female          & Binary variable, 1 = female                                  \\
+Age             & Age in years                                                 \\
+Child           & Binary variable, 1 = is child of head of household           \\
+Grandchild      & Binary variable, 1 = is grandchild of head of household      \\
+Num children    & Number of children in the household                          \\
+Secondary       & Binary variable, 1 = in secondary school group               \\
+hhhead educ     & Years of education of head of household                      \\
+female hhhead   & Binary variable, 1 = female head of household                \\
+Floods          & Number of flood occurrences per dept per year                \\
+Droughts        & Number of drought occurrences per dept per year              \\
+Dept            & Department                                                   \\
+Valleys         & Binary variable, 1 = department is in valleys region         \\
+Mountains       & Binary variable, 1 = department is in mountains region       \\
+Plains          & Binary variable, 1 = department is in plains region          \\
+
+\hline
+\end{tabular}
+\end{table}
+
+```{r create summary statistic table, echo = F, results = 'asis'}
+#creating a table with the main explanatory variables
+
+summary_data <- reg_data %>%
+  select(attend, register, work_reason, lny, extr, female, age, child, grandchild, num_children, secondary, hhhead_educ, female_hhhead, floods, droughts, dept, valleys, plains, mountains)%>%
+  summarise_all(list(mean = mean, sd = sd))
+
+stargazer::stargazer(as.data.frame(summary_data),
+                     label = "summaryStats",
+                     title = "Summary statistics (full sample)",
+                     header = F,
+                     type = "latex",
+                     digits = 0)
+
+#creating a table dividing the characteristics by urban/rural area
+summary_rural <- rural %>%
+  select(attend, register, work_reason, lny, extr, female, age, child, grandchild, num_children, secondary, hhhead_educ, female_hhhead, floods, droughts, dept, valleys, plains, mountains) %>%
+  summarise_all(list(mean = mean, sd = sd))
+
+stargazer::stargazer(as.data.frame(summary_data),
+                     label = "summaryStatsR",
+                     title = "Summary statistics (rural sample)",
+                     header = F,
+                     type = "latex",
+                     digits = 0)
+
+```
+
+#Empirical Model
+see the link for math equation help.
 
@@
 #summary(model_simple)
-model_assist <- lm(assist ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + dept_factor + year_factor, data = reg_data)
+model_assist <- lm(attend ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + dept_factor + year_factor, data = reg_data)
 summary(model_assist)
@@
 ```
+
 ```{r probit regression total sample, echo = F, message = F}
 
-probitMod_assist <- glm(assist ~ floods + droughts + age + female + lny + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_assist <- glm(attend ~ floods + droughts + age + female + lny + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
@@
                title = "Full sample probit regression results", 
-               label = "fullResults",
+               label = "fullResults1",
                header = F,
@@
 
-rural <- subset(reg_data, rural == 1, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "income", "assist", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason", "primary"))
-
-model_assist_r <- (lm(assist ~ floods + droughts + age + female + income + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
+model_assist_r <- (lm(attend ~ floods + droughts + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
 summary(model_assist_r)
-model_register_r <- lm(register ~ floods + droughts + age + female + income + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural)
+model_register_r <- lm(register ~ floods + droughts + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural)
 summary(model_register_r)
-model_reason <- (lm(work_reason ~ floods + droughts + age + female + income + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
+model_reason <- (lm(work_reason ~ floods + droughts + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
 summary(model_reason)
@@
                title = "Rural sample linear regression results", 
-               label = "fullResults",
+               label = "fullResults2",
                header = F,
@@
 
-probitMod_assist_r <- glm(assist ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+probitMod_assist_r <- glm(attend ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
 
@@
 
-urban <- subset(reg_data, rural == 0, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "income", "assist", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor"))
+urban <- subset(reg_data, rural == 0, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "lny", "attend", "register", "work_reason", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "extr", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov"))
 
-model_assist_u <- lm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
+model_assist_u <- lm(attend ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
 summary(model_assist_u)
-model_register_u <- lm(register ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
+model_register_u <- lm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
 summary(model_register_u)
@@
 
-probitMod_assist_u <- glm(assist ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+probitMod_assist_u <- glm(attend ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
 

### 2020-05-07 paper_draft.Rmd
trying to get lagged data in

@@
 
+#lagged flood data created in excel
+library(xlsx)
+write.xlsx(flood_summary, "C:\Users\katya\Documents\MyDocuments\A Smith College Spring Semester\Natural Resource Economics\seminar-paper-katyagi21/lagflood.xlsx")
+#library(DataCombine)
+#flood_lags <- slide(flood_summary, Var = "floods", slideBy = -1)
+
 export_summary_table(dfSummary(flood_summary))
+
 ```

### 2020-05-07 paper_draft.Rmd
created lagged variables

--- 2020-05-07 paper_draft.Rmd
@@
 
-export_summary_table(dfSummary(flood_summary))
+#lagged flood data created in excel
+flood_lag <- read_csv(file.path("Raw data", "lagflood.csv"))
+#library(DataCombine)
+#flood_lags <- slide(flood_summary, Var = "floods", slideBy = -1)
+
+flood_total <- flood_summary %>% 
+  left_join(flood_lag)
+
+export_summary_table(dfSummary(flood_total))
+
 ```
@@
   mutate(year = as.numeric(year))
+
+#lagged drought data created in excel
+#library(xlsx)
+#write.xlsx(drought_summary, "C:\\Users\\katya\\Documents\\MyDocuments\\A Smith College Spring Semester\\Natural Resource Economics\\seminar-paper-katyagi21\\lagdrought.xlsx")
+
+drought_lag <- read_csv(file.path("Raw data", "lagdrought.csv"))
+
+drought_total <- drought_summary %>% 
+  left_join(drought_lag)
 

### 2020-05-07 paper_draft.Rmd
summary stats line 621

@@
 
-export_summary_table(dfSummary(drought_summary))
+export_summary_table(dfSummary(drought_total))
 
@@
 total_data<- all_years %>% 
-  left_join(flood_summary) %>%
-  left_join(drought_summary) %>%
+  left_join(flood_total) %>%
+  left_join(drought_total) %>%
   mutate(child = case_when(relation == 3 ~ 1,
@@
   group_by(dept_factor) %>%
-  #summarise(n = n())
+  summarise(n = n())
 
@@
 
-summary_data <- reg_data %>%
+summary_total <- reg_data %>%
   select(attend, register, work_reason, lny, extr, female, age, child, grandchild, num_children, secondary, hhhead_educ, female_hhhead, floods, droughts, dept, valleys, plains, mountains)%>%
@@
 
-stargazer::stargazer(as.data.frame(summary_data),
+stargazer::stargazer(as.data.frame(summary_total), 
+                     summary = T,
                      label = "summaryStats",
@@
                      type = "latex",
-                     digits = 0)
+                     digits = 3)
 
@@
 
-stargazer::stargazer(as.data.frame(summary_data),
+stargazer::stargazer(as.data.frame(summary_rural),
                      label = "summaryStatsR",
                      title = "Summary statistics (rural sample)",
+                     summary = T,
                      header = F,
                      type = "latex",
-                     digits = 0)
+                     digits = 3)
 
@@
 
-```{r linear regressions total sample, echo = F, message = F}
-
-#example code:
-#model_levels <- lm(recreation_visits ~ svr_median + svr_p10 + month_code*unit_code, data = regression_data)
-
-#model_simple <- lm(assist ~ floods + droughts + year, data = reg_data)
-#summary(model_simple)
-model_assist <- lm(attend ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + dept_factor + year_factor, data = reg_data)
-summary(model_assist)
-model_register <- lm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + dept_factor + year_factor, data = reg_data)
-summary(model_register)
-model_reason_t <- (lm(work_reason ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data))
-summary(model_reason_t)
-
-```
-
-```{r generate regression table, echo = F, results = "asis"}
-stargazer::stargazer(list(model_assist, model_register), 
-               title = "Full sample regression results", 
-               label = "fullResults",
-               header = F,
-               keep = c("floods", "droughts"))
-```
+#Results
 
@@
 
-probitMod_assist <- glm(attend ~ floods + droughts + age + female + lny + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_assist <- glm(attend ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
-probitMod_register <- glm(register ~ floods + droughts + age + female + lny + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_register <- glm(register ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
-probitMod_work <- glm(work_reason ~ floods + droughts + age + female + lny + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_work <- glm(work_reason ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
@@
 
+```{r linear regressions total sample, echo = F, message = F}
+
+#example code:
+#model_levels <- lm(recreation_visits ~ svr_median + svr_p10 + month_code*unit_code, data = regression_data)
+
+#model_simple <- lm(assist ~ floods + droughts + year, data = reg_data)
+#summary(model_simple)
+model_assist <- lm(attend ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + dept_factor + year_factor, data = reg_data)
+summary(model_assist)
+model_register <- lm(register ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + dept_factor + year_factor, data = reg_data)
+summary(model_register)
+model_reason_t <- (lm(work_reason ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data))
+summary(model_reason_t)
+
+```
+
+```{r generate regression table, echo = F, results = "asis"}
+stargazer::stargazer(list(model_assist, model_register), 
+               title = "Full sample linear regression results", 
+               label = "fullResults",
+               header = F,
+               keep = c("floods", "droughts"))
+```
+
 ```{r regression rural sample, echo = F, message = F}
 
-model_assist_r <- (lm(attend ~ floods + droughts + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
+model_assist_r <- (lm(attend ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
 summary(model_assist_r)
-model_register_r <- lm(register ~ floods + droughts + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural)
+model_register_r <- lm(register ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural)
 summary(model_register_r)
-model_reason <- (lm(work_reason ~ floods + droughts + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
+model_reason <- (lm(work_reason ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
 summary(model_reason)
@@
 
-probitMod_assist_r <- glm(attend ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+probitMod_assist_r <- glm(attend ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
 
-probitMod_register_r <- glm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+probitMod_register_r <- glm(register ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
 
-probitMod_work_r <- glm(work_reason ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+probitMod_work_r <- glm(work_reason ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
 
@@
 
-urban <- subset(reg_data, rural == 0, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "lny", "attend", "register", "work_reason", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "extr", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov"))
+urban <- subset(reg_data, rural == 0, select = c("hh_id", "id", "age", "female", "literacy", "secondary", "dept_factor", "rural", "lny", "attend", "register", "work_reason", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "extr", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov"))
 
-model_assist_u <- lm(attend ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
+model_assist_u <- lm(attend ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
 summary(model_assist_u)
-model_register_u <- lm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
+model_register_u <- lm(register ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
 summary(model_register_u)
@@
 
-probitMod_assist_u <- glm(attend ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+probitMod_assist_u <- glm(attend ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
 
-probitMod_register_u <- glm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+probitMod_register_u <- glm(register ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
 
-probitMod_work_u <- glm(work_reason ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+probitMod_work_u <- glm(work_reason ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
 

### 2020-05-07 paper_draft.Rmd
added lagged variables to descriptions

@@
 Floods          & Number of flood occurrences per dept per year                \\
-Droughts        & Number of drought occurrences per dept per year              \\
-Dept            & Department                                                   \\
+Droughts        & Number of drought occurrences per dept in current year       \\
+Lag floods      & Number of flood occurrences per dept in the previous year    \\
+Lag droughts    & Number of drought occurrences per dept in the previous year  \\
+Dept            & Department (nine options)                                    \\
 Valleys         & Binary variable, 1 = department is in valleys region         \\
@@
 
-summary_total <- reg_data %>%
-  select(attend, register, work_reason, lny, extr, female, age, child, grandchild, num_children, secondary, hhhead_educ, female_hhhead, floods, droughts, dept, valleys, plains, mountains)%>%
-    summarise_all(list(mean = mean, sd = sd))
+table_total <- reg_data %>%
+  select(attend, register, work_reason, lny, extr, female, age, child, grandchild, num_children, secondary, hhhead_educ, female_hhhead, floods, droughts, dept, valleys, plains, mountains)
 
-stargazer::stargazer(as.data.frame(summary_total), 
+stargazer::stargazer(as.data.frame(table_total), 
                      summary = T,
@@
 #creating a table dividing the characteristics by urban/rural area
-summary_rural <- rural %>%
-  select(attend, register, work_reason, lny, extr, female, age, child, grandchild, num_children, secondary, hhhead_educ, female_hhhead, floods, droughts, dept, valleys, plains, mountains) %>%
-    summarise_all(list(mean = mean, sd = sd))
+table_rural <- rural %>%
+  select(attend, register, work_reason, lny, extr, female, age, child, grandchild, num_children, secondary, hhhead_educ, female_hhhead, floods, droughts, dept, valleys, plains, mountains)
 
-stargazer::stargazer(as.data.frame(summary_rural),
+stargazer::stargazer(as.data.frame(table_rural),
                      label = "summaryStatsR",

### 2020-05-07 full_paper_draft.Rmd
adding all of the regressions and robusteness checks

--- 2020-05-07 paper_draft.Rmd
@@
 
-children <- subset(total_data, age < 18 & age > 4, select = c("hh_id", "age", "female", "child", "grandchild", "literacy", "dept", "rural", "attend", "reason", "register", "primary", "secondary", "floods", "droughts", "year", "job", "tasks", "money_reason", "work_reason"))
+children <- subset(total_data, age < 18 & age > 4, select = c("hh_id", "age", "female", "child", "grandchild", "literacy", "dept", "rural", "attend", "reason", "register", "primary", "secondary", "floods", "droughts", "lag_floods", "lag_droughts", "year", "money_reason", "work_reason"))
 
@@
          flood_pov = floods * extr,
-         drought_pov = droughts * extr)
+         drought_pov = droughts * extr,
+         flood_lny = floods * lny,
+         drought_lny = floods * lny,
+         lag_flood_lny = lag_floods * lny,
+         lag_drought_lny = lag_droughts * lny,
+         lag_f_secondary = lag_floods * secondary,
+         lag_d_secondary = lag_droughts * secondary)
 
@@
 #creating rural sample
-rural <- subset(reg_data, rural == 1, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "lny", "attend", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason", "secondary", "valleys", "mountains", "plains", "dept", "extr", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov"))
+rural <- subset(reg_data, rural == 1, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "lny", "attend", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason", "secondary", "valleys", "mountains", "plains", "dept", "extr", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov", "flood_lny", "drought_lny", "lag_flood_lny", "lag_drought_lny", "lag_f_secondary", "lag_d_secondary"))
 
@@
 
-probitMod_assist <- glm(attend ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_assist <- glm(attend ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
-probitMod_register <- glm(register ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_register <- glm(register ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
-probitMod_work <- glm(work_reason ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + primary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_work <- glm(work_reason ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
-summary(probitMod_assist)  # model summary
-summary(probitMod_register)
-summary(probitMod_work)
 ```
@@
                header = F,
-               keep = c("floods", "droughts"))
+               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
 
@@
 
+```{r probit regression rural sample, echo = F, message = F}
+
+probitMod_assist_r <- glm(attend ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+
+probitMod_register_r <- glm(register ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+
+probitMod_work_r <- glm(work_reason ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+
+```
+
+```{r generate probit regression table rural sample, echo = F, results = "asis"}
+stargazer::stargazer(list(probitMod_assist_r, probitMod_register_r, probitMod_work_r), 
+               title = "Rural sample probit regression results", 
+               label = "fullResults3",
+               header = F,
+               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
+
+```
+
+```{r probit regression urban sample, echo = F, message = F}
+
+urban <- subset(reg_data, rural == 0, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "lny", "attend", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason", "secondary", "valleys", "mountains", "plains", "dept", "extr", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov", "flood_lny", "drought_lny", "lag_flood_lny", "lag_drought_lny", "lag_f_secondary", "lag_d_secondary"))
+
+probitMod_assist_u <- glm(attend ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+
+probitMod_register_u <- glm(register ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+
+probitMod_work_u <- glm(work_reason ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+
+```
+
+```{r generate probit regression table urban sample, echo = F, results = "asis"}
+stargazer::stargazer(list(probitMod_assist_u, probitMod_register_u, probitMod_work_u), 
+               title = "Urban sample probit regression results", 
+               label = "fullResults4",
+               header = F,
+               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
+
+```
+
+The following tables show the results with the lagged variables for floods and droughts.
+```{r probit regression total sample lagged variables, echo = F, message = F}
+
+probitMod_assist_lt <- glm(attend ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+
+probitMod_register_lt <- glm(register ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+
+probitMod_work_lt <- glm(work_reason ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+
+```
+
+```{r generate probit regression table lagged variables, echo = F, results = "asis"}
+stargazer::stargazer(list(probitMod_assist_lt, probitMod_register_lt, probitMod_work_lt), 
+               title = "Full sample probit regression results (lagged)", 
+               label = "fullResults5",
+               header = F,
+               keep = c("floods", "droughts", "lag_f_secondary", "drought_f_secondary", "lag_flood_lny", "lag_drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
+
+```
+
+```{r probit regression rural sample lagged, echo = F, message = F}
+
+probitMod_assist_lr <- glm(attend ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+
+probitMod_register_lr <- glm(register ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+
+probitMod_work_lr <- glm(work_reason ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+
+```
+
+```{r generate probit regression table rural sample lagged, echo = F, results = "asis"}
+stargazer::stargazer(list(probitMod_assist_lr, probitMod_register_lr, probitMod_work_lr), 
+               title = "Rural sample probit regression results (lagged)", 
+               label = "fullResults6",
+               header = F,
+               keep = c("floods", "droughts", "lag_f_secondary", "drought_f_secondary", "lag_flood_lny", "lag_drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
+
+```
+
+```{r probit regression urban sample lagged, echo = F, message = F}
+
+probitMod_assist_lu <- glm(attend ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+
+probitMod_register_lu <- glm(register ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+
+probitMod_work_lu <- glm(work_reason ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+
+```
+
+```{r generate probit regression table urban sample lagged, echo = F, results = "asis"}
+stargazer::stargazer(list(probitMod_assist_lu, probitMod_register_lu, probitMod_work_lu), 
+               title = "Urban sample probit regression results (lagged)", 
+               label = "fullResults7",
+               header = F,
+               keep = c("floods", "droughts", "lag_f_secondary", "drought_f_secondary", "lag_flood_lny", "lag_drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
+
+```
+
+##Robustness Checks
+
+The following results show a different income specification, the inclusion of the extreme poverty variable instead of the log of income.
+
+```{r probit regression total sample extr, echo = F, message = F}
+
+probitMod_assistx <- glm(attend ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+
+probitMod_registerx <- glm(register ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+
+probitMod_workx <- glm(work_reason ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+
+```
+
+```{r generate probit regression table extr, echo = F, results = "asis"}
+stargazer::stargazer(list(probitMod_assistx, probitMod_registerx, probitMod_workx), 
+               title = "Full sample probit regression results (poverty measure)", 
+               label = "fullResults8",
+               header = F,
+               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov","age", "female", "extr", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
+
+```
+
+```{r probit regression rural sample extr, echo = F, message = F}
+
+probitMod_assist_rx <- glm(attend ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+
+probitMod_register_rx <- glm(register ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+
+probitMod_work_rx <- glm(work_reason ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+
+```
+
+```{r generate probit regression table rural sample extr, echo = F, results = "asis"}
+stargazer::stargazer(list(probitMod_assist_rx, probitMod_register_rx, probitMod_work_rx), 
+               title = "Rural sample probit regression results (poverty measure)", 
+               label = "fullResults9",
+               header = F,
+               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov","age", "female", "extr", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
+
+```
+
+```{r probit regression urban sample extr, echo = F, message = F}
+
+probitMod_assist_ux <- glm(attend ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+
+probitMod_register_ux <- glm(register ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+
+probitMod_work_ux <- glm(work_reason ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+
+```
+
+```{r generate probit regression table urban sample extr, echo = F, results = "asis"}
+stargazer::stargazer(list(probitMod_assist_ux, probitMod_register_ux, probitMod_work_ux), 
+               title = "Urban sample probit regression results (poverty measure)", 
+               label = "fullResults10",
+               header = F,
+               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov","age", "female", "extr", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
+
+```
+
+The following results show the linear regression models.
+
 ```{r linear regressions total sample, echo = F, message = F}
 
-#example code:
-#model_levels <- lm(recreation_visits ~ svr_median + svr_p10 + month_code*unit_code, data = regression_data)
+model_assist_t <- lm(attend ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
 
-#model_simple <- lm(assist ~ floods + droughts + year, data = reg_data)
-#summary(model_simple)
-model_assist <- lm(attend ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + dept_factor + year_factor, data = reg_data)
-summary(model_assist)
-model_register <- lm(register ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + dept_factor + year_factor, data = reg_data)
-summary(model_register)
-model_reason_t <- (lm(work_reason ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data))
-summary(model_reason_t)
+model_register_t <- lm(register ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
+
+model_reason_t <- (lm(work_reason ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data))
 
@@
 ```{r generate regression table, echo = F, results = "asis"}
-stargazer::stargazer(list(model_assist, model_register), 
+stargazer::stargazer(list(model_assist_t, model_register_t, model_reason_t), 
                title = "Full sample linear regression results", 
-               label = "fullResults",
+               label = "fullResults11",
                header = F,
-               keep = c("floods", "droughts"))
+               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
 ```
@@
 summary(model_register_r)
-model_reason <- (lm(work_reason ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
+model_reason_r <- (lm(work_reason ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
 summary(model_reason)
@@
 ```{r generate linear regression table rural sample, echo = F, results = "asis"}
-stargazer::stargazer(list(model_assist, model_register), 
+stargazer::stargazer(list(model_assist_r, model_register_r, model_reason_r), 
                title = "Rural sample linear regression results", 
-               label = "fullResults2",
+               label = "fullResults12",
                header = F,
-               keep = c("floods", "droughts"))
-```
-
-```{r probit regression rural sample, echo = F, message = F}
-
-probitMod_assist_r <- glm(attend ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
-
-probitMod_register_r <- glm(register ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
-
-probitMod_work_r <- glm(work_reason ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
-
-summary(probitMod_assist_r)  # model summary
-summary(probitMod_register_r)
-summary(probitMod_work_r)
+               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
 ```
@@
 
-urban <- subset(reg_data, rural == 0, select = c("hh_id", "id", "age", "female", "literacy", "secondary", "dept_factor", "rural", "lny", "attend", "register", "work_reason", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "extr", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov"))
+model_assist_u <- lm(attend ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban)
 
-model_assist_u <- lm(attend ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
-summary(model_assist_u)
-model_register_u <- lm(register ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
-summary(model_register_u)
+model_register_u <- lm(register ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban)
+
+model_reason_u <- lm(work_reason ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban)
+
 ```
@@
 ```{r linear regression generate table urban sample, echo = F, results = "asis"}
-stargazer::stargazer(list(model_assist, model_register), 
+stargazer::stargazer(list(model_assist_u, model_register_u, model_reason_u), 
                title = "Urban sample linear regression results", 
@@
                header = F,
-               keep = c("floods", "droughts"))
+               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
 
-```
-
-```{r probit regression urban sample, echo = F, message = F}
-
-probitMod_assist_u <- glm(attend ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
-
-probitMod_register_u <- glm(register ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
-
-probitMod_work_u <- glm(work_reason ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
-
-summary(probitMod_assist_r)  # model summary
-summary(probitMod_register_r)
-summary(probitMod_work_r)
 ```

### 2020-05-07 full_paper_draft.Rmd
added region fixed effects in robustness checks

@@
 
+The following results show the region fixed-effects instead of department fixed-effects.
+
+```{r probit regression total sample region, echo = F, message = F}
+
+probitMod_assist_g <- glm(attend ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = reg_data, family=binomial(link="probit"))
+
+probitMod_register_g <- glm(register ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = reg_data, family=binomial(link="probit"))
+
+probitMod_work_g <- glm(work_reason ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = reg_data, family=binomial(link="probit"))
+
+```
+
+```{r generate probit regression table region, echo = F, results = "asis"}
+stargazer::stargazer(list(probitMod_assist_g, probitMod_register_g, probitMod_work_g), 
+               title = "Full sample probit regression results", 
+               label = "fullResults13",
+               header = F,
+               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead", "mountains", "plains"))
+
+```
+
+```{r probit regression rural sample region, echo = F, message = F}
+
+probitMod_assist_rg <- glm(attend ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = rural, family=binomial(link="probit"))
+
+probitMod_register_rg <- glm(register ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = rural, family=binomial(link="probit"))
+
+probitMod_work_rg <- glm(work_reason ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = rural, family=binomial(link="probit"))
+
+```
+
+```{r generate probit regression table rural sample region, echo = F, results = "asis"}
+stargazer::stargazer(list(probitMod_assist_rg, probitMod_register_rg, probitMod_work_rg), 
+               title = "Rural sample probit regression results", 
+               label = "fullResults14",
+               header = F,
+               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead", "mountains", "plains"))
+
+```
+
+```{r probit regression urban sample region, echo = F, message = F}
+
+probitMod_assist_ug <- glm(attend ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = urban, family=binomial(link="probit"))
+
+probitMod_register_ug <- glm(register ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = urban, family=binomial(link="probit"))
+
+probitMod_work_ug <- glm(work_reason ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = urban, family=binomial(link="probit"))
+
+```
+
+```{r generate probit regression table urban sample region, echo = F, results = "asis"}
+stargazer::stargazer(list(probitMod_assist_ug, probitMod_register_ug, probitMod_work_ug), 
+               title = "Urban sample probit regression results", 
+               label = "fullResults15",
+               header = F,
+               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead", "mountains", "plains"))
+
+```
+
 The following results show the linear regression models.

### 2020-05-07 full_paper_draft.Rmd
fixing regression tables

@@
 
-export_summary_table(dfSummary(flood_total))
-
 ```
@@
          flood_lny = floods * lny,
-         drought_lny = floods * lny,
+         drought_lny = droughts * lny,
          lag_flood_lny = lag_floods * lny,
@@
 #creating rural sample
-rural <- subset(reg_data, rural == 1, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "lny", "attend", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason", "secondary", "valleys", "mountains", "plains", "dept", "extr", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov", "flood_lny", "drought_lny", "lag_flood_lny", "lag_drought_lny", "lag_f_secondary", "lag_d_secondary"))
+rural <- subset(reg_data, rural == 1, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "lny", "attend", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason", "secondary", "valleys", "mountains", "plains", "dept", "extr", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov", "flood_lny", "drought_lny", "lag_flood_lny", "lag_drought_lny", "lag_f_secondary", "lag_d_secondary", "lag_floods", "lag_droughts"))
 
@@
 
-urban <- subset(reg_data, rural == 0, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "lny", "attend", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason", "secondary", "valleys", "mountains", "plains", "dept", "extr", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov", "flood_lny", "drought_lny", "lag_flood_lny", "lag_drought_lny", "lag_f_secondary", "lag_d_secondary"))
+urban <- subset(reg_data, rural == 0, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "lny", "attend", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason", "secondary", "valleys", "mountains", "plains", "dept", "extr", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov", "flood_lny", "drought_lny", "lag_flood_lny", "lag_drought_lny", "lag_f_secondary", "lag_d_secondary", "lag_floods", "lag_droughts"))
 
@@
 stargazer::stargazer(list(probitMod_assist_g, probitMod_register_g, probitMod_work_g), 
-               title = "Full sample probit regression results", 
+               title = "Full sample probit regression results (region fixed-effects)", 
                label = "fullResults13",
@@
 stargazer::stargazer(list(probitMod_assist_rg, probitMod_register_rg, probitMod_work_rg), 
-               title = "Rural sample probit regression results", 
+               title = "Rural sample probit regression results (region fixed-effects)", 
                label = "fullResults14",
@@
 stargazer::stargazer(list(probitMod_assist_ug, probitMod_register_ug, probitMod_work_ug), 
-               title = "Urban sample probit regression results", 
+               title = "Urban sample probit regression results (region fixed-effects)", 
                label = "fullResults15",
@@
 model_assist_r <- (lm(attend ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
-summary(model_assist_r)
+
 model_register_r <- lm(register ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural)
-summary(model_register_r)
+
 model_reason_r <- (lm(work_reason ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
-summary(model_reason)
 
@@
 
-The "assist" variable shows that around 60% of people who are registered for school currently assist, while 40% do not assist.
-The "reason" variable shows that two most common reasons for missing school are for work (31.3%) or vacation (20.5).
-Out of the people who are registered for school, 6.6% are repeating the grade they are currently in.
-
 The patterns for households show that the average household has 5.3 members, and the average per capita monthly income per household is 258 bolivianos.

### 2020-05-08 full_paper_draft.Rmd
made line graphs for variation

@@
 ```
+
 The drought data also spans the years 2005-2012 and shows the frequency of droughts in each municipality by year.
@@
 
+reg_data %>%
+   group_by(attend, register) %>%
+   summarise(n = n())
+
 ```
@@
 
+```{r creating line graphs for variation within variables, echo = F, message = F}
+
+Beni <- subset(flood_summary, dept == "Beni" & year > 2003, select = c("year", "floods")) 
+Chuq <- subset(flood_summary, dept == "Chuquisaca" & year > 2003, select = c("year", "floods"))
+Pando <- subset(flood_summary, dept == "Pando" & year > 2003, select = c("year", "floods"))
+SantaCruz <- subset(flood_summary, dept == "Santa Cruz" & year > 2003, select = c("year", "floods"))
+Oruro <- subset(flood_summary, dept == "Oruro" & year > 2003, select = c("year", "floods"))
+Potosi <- subset(flood_summary, dept == "Potosi" & year > 2003, select = c("year", "floods"))
+LaPaz <- subset(flood_summary, dept == "La Paz" & year > 2003, select = c("year", "floods"))
+Tarija <- subset(flood_summary, dept == "Tarija" & year > 2003, select = c("year", "floods"))
+Cocha <- subset(flood_summary, dept == "Cochabamba" & year > 2003, select = c("year", "floods"))
+# Create a first line
+plot(LaPaz, type = "b", frame = FALSE, pch = 19, 
+     col = "red", xlab = "year", ylab = "flood counts")
+# Add lines
+lines(Potosi, pch = 18, col = "orange", type = "b", lty = 1)
+lines(Oruro, pch = 18, col = "orangered4", type = "b", lty = 1)
+lines(Beni, pch = 18, col = "green", type = "b", lty = 1)
+lines(Pando, pch = 18, col = "yellow", type = "b", lty = 1)
+lines(SantaCruz, pch = 18, col = "green4", type = "b", lty = 1)
+lines(Cocha, pch = 18, col = "midnightblue", type = "b", lty = 1)
+lines(Tarija, pch = 18, col = "mediumorchid1", type = "b", lty = 1)
+lines(Chuq, pch = 18, col = "blue", type = "b", lty = 1)
+# Add a legend to the plot
+legend("topleft", legend=c("LaPaz", "Potosi", "Oruro", "Beni", "Pando", "SantaCruz", "Cochabamba", "Tarija", "Chuquisaca"), col=c("red", "orange", "orangered4", "green", "yellow", "green4", "midnightblue", "mediumorchid1", "blue"), lty = 1:2, cex=0.8)
+
+```
+
 #Empirical Model
@@
 
-probitMod_register <- glm(register ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_register <- glm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
@@
 
-probitMod_register_r <- glm(register ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+probitMod_register_r <- glm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
 
@@
 
-probitMod_register_u <- glm(register ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+probitMod_register_u <- glm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
 
@@
 
-probitMod_register_lt <- glm(register ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_register_lt <- glm(register ~ lag_floods + lag_flood_lny + lag_droughts + lag_drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
@@
 
-probitMod_register_lr <- glm(register ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+probitMod_register_lr <- glm(register ~ lag_floods + lag_flood_lny + lag_droughts + lag_drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
 
@@
 
-probitMod_register_lu <- glm(register ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+probitMod_register_lu <- glm(register ~ lag_floods + lag_flood_lny + lag_droughts + lag_drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
 
@@
 
-probitMod_registerx <- glm(register ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_registerx <- glm(register ~ floods + flood_pov + droughts + drought_pov + age + female + extr + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
@@
 
-probitMod_register_rx <- glm(register ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+probitMod_register_rx <- glm(register ~ floods + flood_pov + droughts + drought_pov + age + female + extr + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
 
@@
 
-probitMod_register_ux <- glm(register ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+probitMod_register_ux <- glm(register ~ floods + flood_pov + droughts + drought_pov + age + female + extr + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
 
@@
 
-probitMod_register_g <- glm(register ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_register_g <- glm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = reg_data, family=binomial(link="probit"))
 
@@
 
-probitMod_register_rg <- glm(register ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = rural, family=binomial(link="probit"))
+probitMod_register_rg <- glm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = rural, family=binomial(link="probit"))
 
@@
 
-probitMod_register_ug <- glm(register ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = urban, family=binomial(link="probit"))
+probitMod_register_ug <- glm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = urban, family=binomial(link="probit"))
 
@@
 
-model_register_t <- lm(register ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
+model_register_t <- lm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
 
@@
 
-model_register_r <- lm(register ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural)
+model_register_r <- lm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural)
 
@@
 
-model_register_u <- lm(register ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban)
+model_register_u <- lm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban)
 

### 2020-05-08 final_seminar_paper.Rmd
finishing model specification

--- 2020-05-08 full_paper_draft.Rmd
@@
         breakfast = s4_07,
-        voucher = s4_08,
-        voucher_spent = s4_09,
         educ = e,
@@
         tasks = s5_16b,
+        tasks = 
         pov_line = z,
-        extr_pov = zext)
+        extr_pov = zext) %>%
+  mutate(school_meal = case_when(breakfast == 1 ~ 1,
+                                 breakfast == 2 ~ 0))
 
@@
         pov_line = z,
-        extr_pov = zext)
+        extr_pov = zext) %>%
+  mutate(school_meal = case_when(breakfast == 1 ~ 1,
+                                 breakfast == 2 ~ 1,
+                                 breakfast == 3 ~ 0))
 
@@
 
+```
+
+```{r merging 2011 and 2012 subsample for school lunch analysis}
+meal_data <- data_2012 %>% 
+  rbind(data_2012, data_2011)
+  left_join(flood_total) %>%
+  left_join(drought_total) %>%
 ```
@@
 
-```{r creating line graphs for variation within variables, echo = F, message = F}
+```{r creating line graphs for variation within flood variables, echo = F, message = F}
 
@@
 # Create a first line
-plot(LaPaz, type = "b", frame = FALSE, pch = 19, 
-     col = "red", xlab = "year", ylab = "flood counts")
+plot(LaPaz, type = "b", frame = FALSE, pch = 19, main = "Figure 2. Floods Over Time", cex.main = 1, col = "red", xlab = "year", ylab = "flood counts")
 # Add lines
@@
 # Add a legend to the plot
-legend("topleft", legend=c("LaPaz", "Potosi", "Oruro", "Beni", "Pando", "SantaCruz", "Cochabamba", "Tarija", "Chuquisaca"), col=c("red", "orange", "orangered4", "green", "yellow", "green4", "midnightblue", "mediumorchid1", "blue"), lty = 1:2, cex=0.8)
+legend("topleft", legend=c("La Paz", "Potosi", "Oruro", "Beni", "Pando", "Santa Cruz", "Cochabamba", "Tarija", "Chuquisaca"), col=c("red", "orange", "orangered4", "green", "yellow", "green4", "midnightblue", "mediumorchid1", "blue"), lty = 1:2, cex=0.8)
+
+```
+
+```{r creating line graphs for variation within drought variables, echo = F, message = F}
+
+Benid <- subset(drought_summary, dept == "Beni" & year > 2003, select = c("year", "droughts")) 
+Chuqd <- subset(drought_summary, dept == "Chuquisaca" & year > 2003, select = c("year", "droughts"))
+Pandod <- subset(drought_summary, dept == "Pando" & year > 2003, select = c("year", "droughts"))
+SantaCruzd <- subset(drought_summary, dept == "Santa Cruz" & year > 2003, select = c("year", "droughts"))
+Orurod <- subset(drought_summary, dept == "Oruro" & year > 2003, select = c("year", "droughts"))
+Potosid <- subset(drought_summary, dept == "Potosi" & year > 2003, select = c("year", "droughts"))
+LaPazd <- subset(drought_summary, dept == "La Paz" & year > 2003, select = c("year", "droughts"))
+Tarijad <- subset(drought_summary, dept == "Tarija" & year > 2003, select = c("year", "droughts"))
+Cochad <- subset(drought_summary, dept == "Cochabamba" & year > 2003, select = c("year", "droughts"))
+# Create a first line
+plot(LaPazd, type = "b", frame = FALSE, pch = 19, main = "Figure 3. Droughts Over Time", cex.main = 1, col = "red", xlab = "year", ylab = "drought counts")
+# Add lines
+lines(Potosid, pch = 18, col = "orange", type = "b", lty = 1)
+lines(Orurod, pch = 18, col = "orangered4", type = "b", lty = 1)
+lines(Benid, pch = 18, col = "green", type = "b", lty = 1)
+lines(Pandod, pch = 18, col = "yellow", type = "b", lty = 1)
+lines(SantaCruzd, pch = 18, col = "green4", type = "b", lty = 1)
+lines(Cochad, pch = 18, col = "midnightblue", type = "b", lty = 1)
+lines(Tarijad, pch = 18, col = "mediumorchid1", type = "b", lty = 1)
+lines(Chuqd, pch = 18, col = "blue", type = "b", lty = 1)
+# Add a legend to the plot
+legend("topleft", legend=c("La Paz", "Potosi", "Oruro", "Beni", "Pando", "Santa Cruz", "Cochabamba", "Tarija", "Chuquisaca"), col=c("red", "orange", "orangered4", "green", "yellow", "green4", "midnightblue", "mediumorchid1", "blue"), lty = 1:2, cex=0.8)
 
@@
 #Empirical Model
-see the link for math equation help.
+
+The subscript 'd' refers to the department and 't' to the year, as the probit regression utilizes both department-level and year fixed-effects.
+
+$attend_{dt} = \beta_{1}floods_{dt} + \beta_{2}droughts_{dt} + \beta_{3}lny_{dt} + \beta_{i} + \gamma_{t} + \epsilon_{dt}$
 

### 2020-05-08 final_seminar_paper.Rmd
still have a problem reading down to the discussion line

@@
 
-$attend_{dt} = \beta_{1}floods_{dt} + \beta_{2}droughts_{dt} + \beta_{3}lny_{dt} + \beta_{i} + \gamma_{t} + \epsilon_{dt}$
+$attend_{dt} = \beta_{1}floods_{dt} + \beta_{2}droughts_{dt} + \beta_{3}floods*lny_{dt} + \beta_{4}droughts*lny_{dt} + \beta_{5}age_{dt} + \beta_{6}female_{dt} + \beta_{7}lny_{dt} + \beta_{8}secondary_{dt} + \beta_{9}child_{dt} + \beta_{10}grandchild_{dt} + \beta_{11}num\_children_{dt} + \beta_{12}hhhead\_educ_{dt} + \beta_{13}female\_hhhead_{dt} + \beta_{i} + \gamma_{t} + \epsilon_{dt}$
+
+$register_{dt} = \beta_{1}floods_{dt} + \beta_{2}droughts_{dt} + \beta_{3}floods*lny_{dt} + \beta_{4}droughts*lny_{dt} + \beta_{5}age_{dt} + \beta_{6}female_{dt} + \beta_{7}lny_{dt} + \beta_{8}secondary_{dt} + \beta_{9}child_{dt} + \beta_{10}grandchild_{dt} + \beta_{11}num\_children_{dt} + \beta_{12}hhhead\_educ_{dt} + \beta_{13}female\_hhhead_{dt} + \beta_{i} + \gamma_{t} + \epsilon_{dt}$
+
+$work\_reason_{dt} = \beta_{1}floods_{dt} + \beta_{2}droughts_{dt} + \beta_{3}floods*lny_{dt} + \beta_{4}droughts*lny_{dt} + \beta_{5}age_{dt} + \beta_{6}female_{dt} + \beta_{7}lny_{dt} + \beta_{8}secondary_{dt} + \beta_{9}child_{dt} + \beta_{10}grandchild_{dt} + \beta_{11}num\_children_{dt} + \beta_{12}hhhead\_educ_{dt} + \beta_{13}female\_hhhead_{dt} + \beta_{i} + \gamma_{t} + \epsilon_{dt}$
 
@@
 
-probitMod_assist <- glm(attend ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_assist <- glm(attend ~ floods + flood_lny + droughts + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
@@
 
-probitMod_work <- glm(work_reason ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_work <- glm(work_reason ~ floods + flood_lny + droughts + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
@@
                header = F,
-               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
+               keep = c("floods", "droughts", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
 
@@
 
-probitMod_assist_r <- glm(attend ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+probitMod_assist_r <- glm(attend ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
 
@@
 
-probitMod_work_r <- glm(work_reason ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+probitMod_work_r <- glm(work_reason ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
 
@@
 
-probitMod_assist_u <- glm(attend ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
+probitMod_assist_u <- glm(attend ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
 
@@
 
-probitMod_work_u <- glm(work_reason ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
-
-```
-
-```{r generate probit regression table urban sample, echo = F, results = "asis"}
-stargazer::stargazer(list(probitMod_assist_u, probitMod_register_u, probitMod_work_u), 
-               title = "Urban sample probit regression results", 
-               label = "fullResults4",
-               header = F,
-               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
+probitMod_work_u <- glm(work_reason ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
 
@@
 
-probitMod_assist_lt <- glm(attend ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_assist_lt <- glm(attend ~ lag_floods + lag_flood_lny + lag_droughts + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
@@
 
-probitMod_work_lt <- glm(work_reason ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+probitMod_work_lt <- glm(work_reason ~ lag_floods + lag_flood_lny + lag_droughts + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
 
@@
                header = F,
-               keep = c("floods", "droughts", "lag_f_secondary", "drought_f_secondary", "lag_flood_lny", "lag_drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
+               keep = c("floods", "droughts", "lag_flood_lny", "lag_drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
 
@@
 
-probitMod_assist_lr <- glm(attend ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+probitMod_assist_lr <- glm(attend ~ lag_floods + lag_flood_lny + lag_droughts + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
 
@@
 
-probitMod_work_lr <- glm(work_reason ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+probitMod_work_lr <- glm(work_reason ~ lag_floods + lag_flood_lny + lag_droughts + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
 
@@
                header = F,
-               keep = c("floods", "droughts", "lag_f_secondary", "drought_f_secondary", "lag_flood_lny", "lag_drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
-
-```
-
-```{r probit regression urban sample lagged, echo = F, message = F}
-
-probitMod_assist_lu <- glm(attend ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
-
-probitMod_register_lu <- glm(register ~ lag_floods + lag_flood_lny + lag_droughts + lag_drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
-
-probitMod_work_lu <- glm(work_reason ~ lag_floods + lag_f_secondary + lag_flood_lny + lag_droughts + lag_d_secondary + lag_drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
-
-```
-
-```{r generate probit regression table urban sample lagged, echo = F, results = "asis"}
-stargazer::stargazer(list(probitMod_assist_lu, probitMod_register_lu, probitMod_work_lu), 
-               title = "Urban sample probit regression results (lagged)", 
-               label = "fullResults7",
-               header = F,
-               keep = c("floods", "droughts", "lag_f_secondary", "drought_f_secondary", "lag_flood_lny", "lag_drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
+               keep = c("floods", "droughts", "lag_flood_lny", "lag_drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
 
@@
                keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov","age", "female", "extr", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
-
-```
-
-```{r probit regression rural sample extr, echo = F, message = F}
-
-probitMod_assist_rx <- glm(attend ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
-
-probitMod_register_rx <- glm(register ~ floods + flood_pov + droughts + drought_pov + age + female + extr + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
-
-probitMod_work_rx <- glm(work_reason ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
-
-```
-
-```{r generate probit regression table rural sample extr, echo = F, results = "asis"}
-stargazer::stargazer(list(probitMod_assist_rx, probitMod_register_rx, probitMod_work_rx), 
-               title = "Rural sample probit regression results (poverty measure)", 
-               label = "fullResults9",
-               header = F,
-               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov","age", "female", "extr", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
-
-```
-
-```{r probit regression urban sample extr, echo = F, message = F}
-
-probitMod_assist_ux <- glm(attend ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
-
-probitMod_register_ux <- glm(register ~ floods + flood_pov + droughts + drought_pov + age + female + extr + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
-
-probitMod_work_ux <- glm(work_reason ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
-
-```
-
-```{r generate probit regression table urban sample extr, echo = F, results = "asis"}
-stargazer::stargazer(list(probitMod_assist_ux, probitMod_register_ux, probitMod_work_ux), 
-               title = "Urban sample probit regression results (poverty measure)", 
-               label = "fullResults10",
-               header = F,
-               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov","age", "female", "extr", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
-
-```
-
-The following results show the region fixed-effects instead of department fixed-effects.
-
-```{r probit regression total sample region, echo = F, message = F}
-
-probitMod_assist_g <- glm(attend ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = reg_data, family=binomial(link="probit"))
-
-probitMod_register_g <- glm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = reg_data, family=binomial(link="probit"))
-
-probitMod_work_g <- glm(work_reason ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = reg_data, family=binomial(link="probit"))
-
-```
-
-```{r generate probit regression table region, echo = F, results = "asis"}
-stargazer::stargazer(list(probitMod_assist_g, probitMod_register_g, probitMod_work_g), 
-               title = "Full sample probit regression results (region fixed-effects)", 
-               label = "fullResults13",
-               header = F,
-               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead", "mountains", "plains"))
-
-```
-
-```{r probit regression rural sample region, echo = F, message = F}
-
-probitMod_assist_rg <- glm(attend ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = rural, family=binomial(link="probit"))
-
-probitMod_register_rg <- glm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = rural, family=binomial(link="probit"))
-
-probitMod_work_rg <- glm(work_reason ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = rural, family=binomial(link="probit"))
-
-```
-
-```{r generate probit regression table rural sample region, echo = F, results = "asis"}
-stargazer::stargazer(list(probitMod_assist_rg, probitMod_register_rg, probitMod_work_rg), 
-               title = "Rural sample probit regression results (region fixed-effects)", 
-               label = "fullResults14",
-               header = F,
-               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead", "mountains", "plains"))
-
-```
-
-```{r probit regression urban sample region, echo = F, message = F}
-
-probitMod_assist_ug <- glm(attend ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = urban, family=binomial(link="probit"))
-
-probitMod_register_ug <- glm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = urban, family=binomial(link="probit"))
-
-probitMod_work_ug <- glm(work_reason ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + mountains + plains + year_factor, data = urban, family=binomial(link="probit"))
-
-```
-
-```{r generate probit regression table urban sample region, echo = F, results = "asis"}
-stargazer::stargazer(list(probitMod_assist_ug, probitMod_register_ug, probitMod_work_ug), 
-               title = "Urban sample probit regression results (region fixed-effects)", 
-               label = "fullResults15",
-               header = F,
-               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead", "mountains", "plains"))
 
@@
 
-```{r linear regression urban sample, echo = F, message = F}
-
-model_assist_u <- lm(attend ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban)
-
-model_register_u <- lm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban)
-
-model_reason_u <- lm(work_reason ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban)
-
-```
-
-```{r linear regression generate table urban sample, echo = F, results = "asis"}
-stargazer::stargazer(list(model_assist_u, model_register_u, model_reason_u), 
-               title = "Urban sample linear regression results", 
-               label = "urbanlinearResults",
-               header = F,
-               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
-
-```
-
 # Discussion of Data
 
-The patterns for households show that the average household has 5.3 members, and the average per capita monthly income per household is 258 bolivianos.
-Over half of the households in the sample are located in urban areas (56.3), while under half are located in rural areas (43.7).
-La Paz, Santa Cruz, and Cochabamba are the three departments with the most households in this sample, which corresponds to their overall population distribution.
-
-The patterns in the weather data show that every department experienced multiple floods in 2005, while many but not all experienced droughts.
-The place that experienced the highest number of flood occurrences overall was Cochabamba, while Beni experienced 13 droughts in 2005.
+T

### 2020-05-08 final_seminar_paper.Rmd
commit

@@
 ---
+title: "The Effect of Weather Shocks on Children's Education in Bolivia"
+author: "Katya Garcia-Israel"
+abstract: "Weather shocks have the potential to dramatically affect households' livelihoods and income levels, especially in developing countries.
+A growing subset of developmental and environmental literature focuses on investigating the relationship between environment and weather factors and household changes in behavior.
+One frequently cited response to negative weather shocks that decrease income is an increase in child labor to supplement the household income and a decrease in school attendance.
+This mechanism is especially relevant in agricultural areas, but can also be seen in more urban areas if the weather events cause a loss of property, injury, or displace people.
+This study explores the relationship between flood and drought occurrences and school attendance in Bolivia.
+I find that ..."
+header-includes:
+   - \usepackage{booktabs}
+   - \usepackage{setspace}
+   - \doublespacing
+   - \usepackage[labelfont=bf]{caption}
 output: 
-  pdf_document: default
-  html_document: default
----
-
-title: ""The impact of environmental shocks on children's schooling in Bolivia""
-author: "Katya Garcia-Israel"
-output:
-  html_document:
-    toc: yes
-    df_print: paged
   pdf_document:
-    toc: yes
-    number_sections: yes
----
+    number_sections: true
+    keep_tex: true
 
@@
 ```
-
-```{r set dfSummary css style, echo = F, include = F}
-st_css()
-```
 # Introduction
@@
 
+Several papers have researched the effects of weather events on general household well-being, examining household responses to shocks on income.
+Baez et al (2017) find that rural households in Guatemala responded to income shocks after a tropical storm by increasing child labor.
+Such responses are negatively related to school attendance.
+Jensen (2000) examines differences in school enrollment rates after rainfall shocks and compares them to unaffected areas, finding that investments in children’s education do decrease in areas that have experienced either heavy or low rainfall.
+
+The impacts of weather shocks could potentially be seen some time after their occurrence.
+Marchetta et al (2019) address this by including lagged effects in their model, using a weather variable lagged by time to show the potential for a delayed effect on child outcomes.
+The reasoning behind this, in their case, was that households would observe the impact of the rainy season prior to enrolling in school before making the decision to send their children to school or not.
+Marchetta et al (2019) seek to answer the question of how the duration of the negative effects of weather shocks on education by testing how test scores are impacted after the shock.
+Instead of measuring only school attendance, including test scores in the analysis allows for the possibility of seeing patterns in lowered test scores after the weather event.
+This again suggests the potential of lagged impacts.Shah and Steinberg (2017) found differing effects on enrollment from distinct weather events; droughts led to an increase in school attendance and floods decreased it.
+The authors gave a few different explanations for why they might be finding these results.
+Rainfall could be linked to increased malaria and diarrhea which would keep children from attending school or the school meals being served more during drought seasons could serve as an incentive for attending school.
+
+Negative weather shocks can impact different groups in distinct ways.
+For instance, Marchetta et al find that girls are more negatively affected by extreme weather events.
+In research on India, Zimmermann (2020) also shows that rainfall shocks affect girls’ schooling more than boys’, and younger children are more insulated from the effects of rainfall shock than older children who are more productive workers.
+This preference for educating boys does not hold in all contexts; Jensen (2000) finds no difference between boys’ and girls’ school enrollment as a result of rainfall shocks in Cote d’Ivoire.
+
+As opposed to examining the mechanisms of decreased incomes through damaged harvests, Groppo and Kraehnert (2017) explore the effects of livestock mortality on household incomes and education in Mongolia.
+Their study suggests that weather shock of harsh winters does not affect long-term education choices, as household income decreases but the children do not substitute schooling for labor.
+The mechanisms by which weather events affect child outcomes are not always tied directly to agriculture; Marchetta et al (2017) study cyclones as well as rainfall variability, which suggests that loss of (capital) and damage to property, as well as the larger economy, could play a role in the household response that is not necessarily tied to decreasing income from harm to agriculture.
+Their work also emphasizes lack of access to insurance and savings as a main determinant of the negative child outcomes.
+Teacher attendance during drought rather than rainfall seasons and impacts of migration were other possible mechanisms discussed by Shah and Steinberg (2017).
+
+The methods used in this literature vary.
+Baez et al (2017) used a regression-based analysis of household consumption and different levels of rainfall varying by location, as well as setting up a probability model of which areas would be most affected by rainfall as a robustness check.
+Marchetta et al (2019) used a bivariate probit model as their main empirical strategy and a linear probability model as a robustness check.
+Groppo and Kraehnert (2017) use a difference-in-differences approach to distinguish between areas affected by weather disasters and those that were not.
+Jensen (2000) examines the differences between enrollment rates in areas that experienced a heavy rainfall shock and areas that did not, using deviation from the long-term mean as an indication of a shock.
+The identifying assumption is made that there would have been no differential treatment of children’s schooling between geographic areas in Cote d’Ivoire without the negative rainfall shock, an assumption that is supported by examining the percentages of children enrolled in a prior year when neither area experienced a weather shock.
+Researchers are faced with the problem of ensuring to the best of their ability that the observed changes in household outcomes are not due to unobserved changes unrelated to weather patterns.
+
 # Data and Methods
 
+## Conceptual Framework
+I attempt to answer the question of how weather events impact school attendance and registration using an empirical approach.
+The theoretical arguments for the analysis are based on the ideas discussed in the literature review, that extreme weather events can impact households’ financial situations, especially in areas where agriculture is important for the local economy and people’s livelihoods.
+The connection to schooling comes from the idea that financial burdens can cause children to substitute the time they would have spent in school into participating in economic activities to help increase household income.
+In rural areas, this effect is expected to be the strongest, as there is larger participation in agriculture and many households may use all their members at one point or another to help with the agricultural activities.
+However, both the urban and rural samples are relevant to the research question because agricultural activity might be taking place in some areas classified as urban.
+Additionally, other mechanisms linking floods and droughts to school attendance may be at play, such as decreases in real income from property damage, livestock mortality, or higher food prices.
+
 ## Data
-
-The household data was taken from the social surveys database of the Bolivia National Statistical Institute, using household surveys from the years 2005-2007, 2011-2012.
-Instituto Nacional de Estadistica. (2005). Encuesta de Hogares (2005) [SAV file]. Retrieved from https://www.ine.gob.bo/index.php/banco/base-de-datos-sociales
-
-The unit of observation at the individual level, but each individual is identified mainly by the household id.
+To answer this question, I used a combination of household survey data and flood and drought occurrences data.
+The household survey data was taken from the Bolivian National Statistical Institute as part of their social surveys database.
+The household surveys used were from the years 2005-2007 and 2011-2012.
+The intermediary years (2008, 2009, 2010) were not included because the relevant education questions were not part of the 2008 and 2009 surveys, and there was no household survey taken in 2010, the census year.
+The flood and drought datasets show the number of floods and droughts that occurred per year in each of the nine departments of Bolivia.
+The final compiled dataset is a repeated cross-section as different households were randomly surveyed each year.
+The main unit of observation is the child’s ID, but the household ID is also used as a unit in order to examine household characteristics such as relations and income per person.
+For all the combined years, there are 27,921 observations.
+Table 1 lists the variables used to answer the research question, along with brief descriptions.
 
@@
                                school_level != 18 ~ 0))
-
-```
-
-```{r export summary table 2005, echo = F, message = F}
-export_summary_table(dfSummary(hh_2005))
 
@@
                                school_level != 18 ~ 0))
-```
-
-```{r exporting data table, echo = F, message = F}
-export_summary_table(dfSummary(hh_2006))
 ```
@@
         tasks = s5_16b,
-        tasks = 
         pov_line = z,
-        extr_pov = zext) %>%
-  mutate(school_meal = case_when(breakfast == 1 ~ 1,
-                                 breakfast == 2 ~ 0))
+        extr_pov = zext)
 
@@
         pov_line = z,
-        extr_pov = zext) %>%
-  mutate(school_meal = case_when(breakfast == 1 ~ 1,
-                                 breakfast == 2 ~ 1,
-                                 breakfast == 3 ~ 0))
+        extr_pov = zext)
 
@@
 
-The flood data is from the years 2005-2012.
-The unit of observation is municipality, but the data is later combined to make the unit of observation at the department level to make it comparable with the household survey data.
-
-This data can be found at the following website: http://geosinager.defensacivil.gob.bo/maps/226.
-It is associated with a map showing the flood occurrences information spatially, and to get the data in a non-spatial format, the flood layer should be downloaded.
-It may be downloaded in a variety of different formats, including csv.
-
 ```{r import flood data, echo = F, message = F}
@@
 ```
-
-The drought data also spans the years 2005-2012 and shows the frequency of droughts in each municipality by year.
-
-This data can be obtained from http://geosinager.defensacivil.gob.bo/maps/225.
-It is shown on a map with different layers, and the raw data may be downloaded in a variety of different forms, including csv.
 
@@
 ```
-
-```{r merging 2011 and 2012 subsample for school lunch analysis}
-meal_data <- data_2012 %>% 
-  rbind(data_2012, data_2011)
-  left_join(flood_total) %>%
-  left_join(drought_total) %>%
-```
-
-## Analysis Variables
-
-The variables used in the final analysis are the variables from the flood and drought summaries, showing the number of occurrences in each department per year.
-The other variables used are from the household data, showing demographic information as well as the dependent variables related to school assistance and registration.
-The variables "assist" and "register" tell whether the child is currently registered and attending school, and the variables "grade" and "literacy" give further educational information showing what grade the child is in and whether they can read and write.
-The variables "urb_rur" and "dept" give geographic information, indicating whether the child's household is located in an urban area or a rural area and in what department of the country they are living.
-Demographic variables such as age, sex, relation to the head of household, number of household members, and income can be used in the analysis as controls.
 
@@
 \hline
-Attend          & Binary variable, 1 = attends school                          \\
+Attend          & Binary variable, 1 = attends school (of registered sample)   \\
 Register        & Binary variable, 1 = registered for school                   \\
@@
 
-#Empirical Model
+## Model Specification
 
+The main econometric model used in the analysis is a probit regression model, as my three dependent variables are all binary variables with values of 0 or 1.
+Floods and droughts are the two main explanatory variables, along with the interaction terms floods_lny and droughts_lny, which capture the differing effect that weather shocks have on schooling outcomes for children of different income levels.
+These interaction terms were formed by multiplying the flood and drought variables with the lny, logged per capita monthly income, variable.
+Several control variables are inserted into the regression to control for income, age and sex of the child, sex and education level of the household head, number of other children in the household and whether the child is the child or grandchild of the household head.
 The subscript 'd' refers to the department and 't' to the year, as the probit regression utilizes both department-level and year fixed-effects.
 
-$attend_{dt} = \beta_{1}floods_{dt} + \beta_{2}droughts_{dt} + \beta_{3}floods*lny_{dt} + \beta_{4}droughts*lny_{dt} + \beta_{5}age_{dt} + \beta_{6}female_{dt} + \beta_{7}lny_{dt} + \beta_{8}secondary_{dt} + \beta_{9}child_{dt} + \beta_{10}grandchild_{dt} + \beta_{11}num\_children_{dt} + \beta_{12}hhhead\_educ_{dt} + \beta_{13}female\_hhhead_{dt} + \beta_{i} + \gamma_{t} + \epsilon_{dt}$
+$1.
+attend_{dt} = \beta_{1}floods_{dt} + \beta_{2}droughts_{dt} + \beta_{3}floods*lny_{dt} + \beta_{4}droughts*lny_{dt} + \beta_{5}age_{dt} + \beta_{6}female_{dt} + \beta_{7}lny_{dt} + \beta_{8}secondary_{dt} + \beta_{9}child_{dt} + \beta_{10}grandchild_{dt} + \beta_{11}num\_children_{dt} + \beta_{12}hhhead\_educ_{dt} + \beta_{13}female\_hhhead_{dt} + \beta_{i} + \gamma_{t} + \epsilon_{dt}$
 
-$register_{dt} = \beta_{1}floods_{dt} + \beta_{2}droughts_{dt} + \beta_{3}floods*lny_{dt} + \beta_{4}droughts*lny_{dt} + \beta_{5}age_{dt} + \beta_{6}female_{dt} + \beta_{7}lny_{dt} + \beta_{8}secondary_{dt} + \beta_{9}child_{dt} + \beta_{10}grandchild_{dt} + \beta_{11}num\_children_{dt} + \beta_{12}hhhead\_educ_{dt} + \beta_{13}female\_hhhead_{dt} + \beta_{i} + \gamma_{t} + \epsilon_{dt}$
+$2.
+register_{dt} = \beta_{1}floods_{dt} + \beta_{2}droughts_{dt} + \beta_{3}floods*lny_{dt} + \beta_{4}droughts*lny_{dt} + \beta_{5}age_{dt} + \beta_{6}female_{dt} + \beta_{7}lny_{dt} + \beta_{8}secondary_{dt} + \beta_{9}child_{dt} + \beta_{10}grandchild_{dt} + \beta_{11}num\_children_{dt} + \beta_{12}hhhead\_educ_{dt} + \beta_{13}female\_hhhead_{dt} + \beta_{i} + \gamma_{t} + \epsilon_{dt}$
 
-$work\_reason_{dt} = \beta_{1}floods_{dt} + \beta_{2}droughts_{dt} + \beta_{3}floods*lny_{dt} + \beta_{4}droughts*lny_{dt} + \beta_{5}age_{dt} + \beta_{6}female_{dt} + \beta_{7}lny_{dt} + \beta_{8}secondary_{dt} + \beta_{9}child_{dt} + \beta_{10}grandchild_{dt} + \beta_{11}num\_children_{dt} + \beta_{12}hhhead\_educ_{dt} + \beta_{13}female\_hhhead_{dt} + \beta_{i} + \gamma_{t} + \epsilon_{dt}$
+$3.
+work\_reason_{dt} = \beta_{1}floods_{dt} + \beta_{2}droughts_{dt} + \beta_{3}floods*lny_{dt} + \beta_{4}droughts*lny_{dt} + \beta_{5}age_{dt} + \beta_{6}female_{dt} + \beta_{7}lny_{dt} + \beta_{8}secondary_{dt} + \beta_{9}child_{dt} + \beta_{10}grandchild_{dt} + \beta_{11}num\_children_{dt} + \beta_{12}hhhead\_educ_{dt} + \beta_{13}female\_hhhead_{dt} + \beta_{i} + \gamma_{t} + \epsilon_{dt}$
 
-#Results
+In the first two regression equations, measuring school attendance and school registration, the omitted condition is not attending school and not registering for school, respectively.
+Thus, increases in the dependent variable indicate higher probabilities of going to school.
+Variables such as floods and droughts are expected to have negative coefficients, as I hypothesize that experiencing weather shocks will decrease the likelihood going to school.
+The age variable and the secondary school variable are both expected to have a negative coefficient, as older children would be more likely to be able to work and have a higher opportunity cost of going to school.
+A higher number of children in the household is expected to cause a decrease in the probability of attending school.
+Higher levels of income, being the child or grandchild of the household head, high levels of education of the household head, and having a female household head are all expected to be connected to higher probabilities of school attendance and registration.
+
+The third regression uses the sample of children that do not attend school, and the dependent variable is work_reason, which shows the likelihood of work being the reason why the child is not attending school.
+For the omitted condition of 0, those children listed reasons other than work as their motivation for not attending school.
+In this regression, I would expect the coefficients on the flood and drought variables to be positive, and the control variables to flip signs, as the factors that would cause a household to send a child to school have the opposite effect on the child’s likelihood of working.
+The sample for the third regression is much smaller than the other two regressions, 3,939 versus the 24,316 in the attend sample and the 27,921 in the registering sample.
+
+These three regression equations are repeated in the analysis using different samples, the total number of households and the sample that live in rural areas.
+In the lagged variable analysis, the same model is used, but floods and droughts are replaced by lag_floods and lag_droughts, which represent the number of floods or droughts in each department during the previous year.
+Including the lagged variables is a way of testing whether households’ responses to weather shocks are long-term or short-term.
+
+# Results
 
@@
 
-```{r regression rural sample, echo = F, message = F}
-
-model_assist_r <- (lm(attend ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
-
-model_register_r <- lm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural)
-
-model_reason_r <- (lm(work_reason ~ floods + flood_secondary + droughts + drought_secondary + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural))
-
-```
-
-```{r generate linear regression table rural sample, echo = F, results = "asis"}
-stargazer::stargazer(list(model_assist_r, model_register_r, model_reason_r), 
-               title = "Rural sample linear regression results", 
-               label = "fullResults12",
-               header = F,
-               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
-```
-
-# Discussion of Data
-
-T
+#Discussion

### 2020-05-09 final_seminar_paper.Rmd
adding more text from word

@@
 One frequently cited response to negative weather shocks that decrease income is an increase in child labor to supplement the household income and a decrease in school attendance.
-This mechanism is especially relevant in agricultural areas, but can also be seen in more urban areas if the weather events cause a loss of property, injury, or displace people.
+This mechanism is especially relevant in agricultural areas but can also be seen in more urban areas if the weather events cause a loss of property, injury, or displace people.
 This study explores the relationship between flood and drought occurrences and school attendance in Bolivia.
-I find that ..."
+Contrary to my hypothesis, I find that floods and droughts actually have positive impacts on school outcomes, a relationship that might indicate that weather events negatively impact agricultural harvests and decrease the demand for children’s labor, thereby allowing them to go to school."
 header-includes:
@@
 # Introduction
+
+Weather events have been linked to changes in household income, education and working behaviors, especially in developing countries and areas where people are particularly reliant on agriculture.
+In the 21st century, with the threat of climate change, exploration of these connections is even more crucial.
+In developmental economics literature, the mechanisms by which weather events affect households’ livelihoods tend to be agricultural, linking adverse precipitation or temperature shocks to decreases in income.
+This paper seeks to answer the question of whether high occurrences of floods and droughts in a region have an effect on children’s education, specifically in Bolivia.
+
+A key reason for choosing Bolivia as a case study for this research question is its extensive climactic diversity.
+Despite this high level of internal precipitation and temperature variation, the ecological zones correspond fairly well to political boundaries of the country, which is divided into nine departments.
+Three departments fall into each ecological category, which are mountains (La Paz, Potosi, and Oruro), valleys (Cochabamba, Tarija, Chuquisaca), and plains (Santa Cruz, Beni, Pando).
+Figure 1 shows the three regional climate divisions and the departmental boundaries, with the mountains in the west, the plains in the north and east, and the valleys in the south-central part of the country.
 
@@
 Table 1 lists the variables used to answer the research question, along with brief descriptions.
+
+Table 2 provides summary statistics for the variables used in the analysis using the full sample of children, and Table 3 shows the summary statistics for children living in rural areas.
+The rural variable in this analysis is based on whether the households live in a municipality with fewer than 2000 people.
+The total sample contains 24,316 observations while the rural sample is considerably smaller, with 9,804 observations.
+Many of the statistics have similar values in both samples, but it is notable more children are registered for school in the total sample than in the rural sample, as evidenced by the higher mean on the register variable.
+The average household income is also lower for children in the rural sample, while the percentage of households below the poverty line is higher.
+Overall, most of the children ages 5-17 do attend school, with the high mean of the attend variable (0.986).
 
@@
 
+The following graphs plot the flood variables and the drought variables over the sample years to capture the variation between departments and within the same department over time.
+Figure 2 shows the flood occurrences over time, and Figure 3 shows the droughts.
+Many of the departments do not actually have much variation between the years, but some events stand out in the graph and are clearly extreme shocks.
+There is no evidence of a positive or a negative time trend in these variables.
+Comparing the two graphs, it is evident that in many of the departments have high counts of floods in the same year as low counts of droughts, instead of experiencing both shocks at once in different parts of the region.
+This is reassuring in terms of being able to assume that the departments are a decent measure of location, as some are quite large, but evidently not so internally ecologically diverse that it confuses the analysis.
+The main challenge of this data is the unit of location by which the observations are grouped, as the department level is the most disaggregated measure.
+Yet households on one side of a department many not be experiencing the full effects of the weather shocks than households in another location within the same department.
+
 ```{r creating line graphs for variation within flood variables, echo = F, message = F}
@@
 #Discussion
+
+The flood results ended up being contrary to the hypothesis discussed in the conceptual framework section, the theory supported by the literature review that weather events will have detrimental impacts on children’s education via the mechanism of agricultural damages requiring increased child labor.
+As the results were highly significant, it suggests that in the Bolivian context during the years from 2005-2012, there was another relationship between floods and droughts and schooling.
+The alternate theory is that weather events have negative effects on agricultural activities in an area, which then decreases the demand for labor in agriculture sectors.
+If children’s labor is no longer required if crops are destroyed, the opportunity cost of going to school lessens considerably and they may become more likely to register and attend.
+The significance and direction of the coefficient on the interaction terms linking floods with droughts in the main probit regression also lend credence to that theory.
+
+The drought results for the rural sample did correspond to the argument that weather events decrease income and increase the opportunity cost of school attendance as there was a significant negative relationship between droughts and school attendance.
+The result in Table 4 assumes that the effect of floods or droughts on schooling will be seen in the same year.
+This is a more convincing assumption for floods since they are a sudden-onset weather event that tend to have instantaneous effects, but as droughts are slow-onset their negative effects on people’s livelihoods might not be seen for months or even years, depending on how long they last.
+This issue is addressed by the inclusion of the lagged effects variables in Tables 6 and 7.
+These differing results could also be due to data limitations.
+One main limitation of this dataset is the fact that the flood and drought occurrences are measured at department level.
+Since there may be instances in which floods only affect one small area within a large department, the data cannot capture the fact that some people who are counted as being affected by this weather event by virtue of being in the same department might not be in the vicinity of the affected area at all.
+However, there is also the possibility that weather events in one part of the department can have consequences on other areas, through the mechanisms of higher food prices due to agricultural damage or limitations in human travel flows.

### 2020-05-09 final_seminar_paper.Rmd
Rmd doc with Line 21 error

@@
+---
+output:
+  pdf_document: default
+  html_document: default
+---
 ---

### 2020-05-09 final_seminar_paper.Rmd
Fixing the YAML

--- 2020-05-09 final_seminar_paper.Rmd
@@
     keep_tex: true
+---
 

### 2020-05-10 final_seminar_paper.Rmd
at the point where everything's good except tables are somehow now after the discussion section instead of before and the discussion section is not showing up as a section, # visible

@@
 
-```{r setup, echo = F, message = F}
+```{r setup, echo = F, warning = F, message = F}
 knitr::opts_chunk$set(results = 'asis', cache = F)
@@
 
-```{r import household data 2005, echo = F, message = F}
+```{r import household data 2005, echo = F, warning = F, message = F}
 data_2005 <- read_dta("Raw data/seminar_paper_data/2005_hhsurvey.dta") %>% 
@@
 
-```{r import household data 2006, echo = F, message = F}
+```{r import household data 2006, echo = F, warning = F, message = F}
 data_2006 <- read_dta("Raw data/seminar_paper_data/2006_hhsurvey.dta")  %>% 
@@
 
-```{r import household data 2007, echo = F, message = F}
+```{r import household data 2007, echo = F, warning = F, message = F}
 data_2007 <- read_dta("Raw data/seminar_paper_data/2007_hhsurvey.dta")  %>% 
@@
 
-```{r import household data 2011, echo = F, message = F}
+```{r import household data 2011, echo = F, warning = F, message = F}
 data_2011 <- read_dta("Raw data/seminar_paper_data/2011_hhsurvey.dta")  %>% 
@@
 
-```{r household data 2012, echo = F, message = F}
+```{r household data 2012, echo = F, warning = F, message = F}
 data_2012 <- read_dta("Raw data/seminar_paper_data/2012_hhsurvey.dta")  %>% 
@@
 
-```{r combine household datasets, echo = F, message = F}
+```{r combine household datasets, echo = F, warning = F, message = F}
 all_years <- hh_2005 %>% 
@@
 
-```{r import flood data, echo = F, message = F}
+```{r import flood data, echo = F, warning = F, message = F}
 flood <- read_csv(file.path("Raw data", "recurrencia_inundacion2002_2012bol.csv")) %>% 
@@
 
-```{r importing droughts data, echo = F, message = F}
+```{r importing droughts data, echo = F, warning = F, message = F}
 #drought data set 
@@
 
-``` {r merging datasets, echo = F, message = F}
+``` {r merging datasets, echo = F, warning = F, message = F}
 total_data<- all_years %>% 
@@
 
-```{r creating head of household subset, echo = F, message = F}
+```{r creating head of household subset, echo = F, warning = F, message = F}
 
@@
 
-```{r creating children subset, echo = F, message = F}
+```{r creating children subset, echo = F, warning = F, message = F}
 
@@
 
-total_children %>%
-      group_by(child, grandchild) %>%
-      summarise(n = n())
-
-total_children %>%
-   group_by(work_reason, money_reason) %>%
-   summarise(n = n())
-
 ```
 
-```{r combining household head characteristics with children dataset, echo = F, message = F}
+```{r combining household head characteristics with children, echo = F, warning = F, results = 'hide', message = F}
 hh_head_2 <- hh_head %>%
@@
   summarise(n = n())
+#figure out how to get rid of the table
 
@@
 
-```{r regression variable specifications, echo = F, message = F}
+```{r regression variable specifications, echo = F, warning = F, message = F}
 
@@
 
-reg_data %>%
-  group_by(dept_factor) %>%
-  summarise(n = n())
+```
 
-reg_data %>%
-   group_by(attend, register) %>%
-   summarise(n = n())
-
-```
 ##Variable Descriptions
@@
 
-```{r create summary statistic table, echo = F, results = 'asis'}
+```{r create summary statistic table, echo = F, message = F, warning = F, results = 'asis'}
 #creating a table with the main explanatory variables
@@
 
-```{r probit regression total sample, echo = F, message = F}
+```{r probit regression total sample, echo = F, warning = F, message = F}
 
@@
 
-```{r generate probit regression table, echo = F, results = "asis"}
+```{r generate probit regression table, echo = F, message = F, results = "asis"}
 stargazer::stargazer(list(probitMod_assist, probitMod_register, probitMod_work), 
@@
 
-```{r probit regression rural sample, echo = F, message = F}
+```{r probit regression rural sample, echo = F, warning = F, message = F}
 
@@
 
-```{r generate probit regression table rural sample, echo = F, results = "asis"}
+```{r generate probit regression table rural sample, echo = F, message = F, results = "asis"}
 stargazer::stargazer(list(probitMod_assist_r, probitMod_register_r, probitMod_work_r), 
@@
 
-```{r probit regression urban sample, echo = F, message = F}
+```{r probit regression urban sample, echo = F, warning = F, message = F}
 
@@
 The following tables show the results with the lagged variables for floods and droughts.
-```{r probit regression total sample lagged variables, echo = F, message = F}
+
+```{r probit regression total sample lagged variables, echo = F, warning = F, message = F}
 
@@
 
-```{r generate probit regression table lagged variables, echo = F, results = "asis"}
+```{r generate probit regression table lagged variables, echo = F, message = F, results = "asis"}
 stargazer::stargazer(list(probitMod_assist_lt, probitMod_register_lt, probitMod_work_lt), 
@@
 
-```{r probit regression rural sample lagged, echo = F, message = F}
+```{r probit regression rural sample lagged, echo = F, warning = F, message = F}
 
@@
 
-```{r generate probit regression table rural sample lagged, echo = F, results = "asis"}
+```{r probit regression table rural sample lagged, message= F, echo = F, results = "asis"}
 stargazer::stargazer(list(probitMod_assist_lr, probitMod_register_lr, probitMod_work_lr), 
@@
 
-##Robustness Checks
+## Robustness Checks
 
@@
 
-```{r probit regression total sample extr, echo = F, message = F}
+```{r probit regression total sample extr, echo = F, warning = F, message = F}
 
@@
 
-```{r generate probit regression table extr, echo = F, results = "asis"}
+```{r generate probit regression table extr, echo = F, message = F, results = "asis"}
 stargazer::stargazer(list(probitMod_assistx, probitMod_registerx, probitMod_workx), 
@@
 ```
-
-The following results show the linear regression models.
 
@@
 
-```{r generate regression table, echo = F, results = "asis"}
+```{r generate regression table, echo = F, message = F, results = "asis"}
 stargazer::stargazer(list(model_assist_t, model_register_t, model_reason_t), 
@@
 However, there is also the possibility that weather events in one part of the department can have consequences on other areas, through the mechanisms of higher food prices due to agricultural damage or limitations in human travel flows.
+
+# References
+Baez, J.E., Lucchetti, I., Genoni, M. E., & Salazar, M. (2017). Gone with the storm: Rainfall shocks and household wellbeing in Guatemala. Journal of Development Studies, 53(8), 1253-1271.
+
+Groppo, V., and K. Kraehnert. 2017. The Impact of Extreme Weather Events on Education. Journal of Population Economics 30 (2): 433–72
+
+Jensen, R. (2000). Agricultural volatility and investments in children. American Economic Review Paper and Proceedings, 90(2), 399-404.
+
+Marchetta, F., Sahn, D., Tiberti, L. (2019). The Role of Weather on Schooling and Work of Young Adults in Madagascar. American Journal of Agricultural Economics, 101(4), 1203-1227.
+
+Shah, M., & Steinberg, B.M. (2017). Drought of opportunities: Contemporaneous and long-term impacts of rainfall shocks on human capital. Journal of Political Economy, 125(2), 527-561.
+
+Zimmermann, L. (2020). Remember When It Rained—Schooling Responses to Shocks in India. World Development, 126.

### 2020-05-10 final_seminar_paper.Rmd
where the r document is at right now...won't read again

@@
 ```
+
+```{r, include=FALSE}
+options(tinytex.verbose = TRUE)
+```
+
 # Introduction
@@
 
+```{r probit regression rural sample extr, echo = F, warning = F, message = F}
+
+probitMod_assist_rx <- glm(attend ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+
+probitMod_register_rx <- glm(register ~ floods + flood_pov + droughts + drought_pov + age + female + extr + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+
+probitMod_work_rx <- glm(work_reason ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural, family=binomial(link="probit"))
+
+```
+
+```{r generate probit regression table rural sample extr, echo = F, results = "asis"}
+stargazer::stargazer(list(probitMod_assist_rx, probitMod_register_rx, probitMod_work_rx), 
+               title = "Rural sample probit regression results (poverty measure)", 
+               label = "fullResults9",
+               header = F,
+               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov","age", "female", "extr", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
+
+```
+
 ```{r linear regressions total sample, echo = F, message = F}

### 2020-05-10 data_appendix_all_years.Rmd
Final copy of data appendix

@@
 
-```{r regressions total sample}
-
-#example code:
-#model_levels <- lm(recreation_visits ~ svr_median + svr_p10 + month_code*unit_code, data = regression_data)
-
-#model_simple <- lm(assist ~ floods + droughts + year, data = reg_data)
-#summary(model_simple)
-model_assist <- lm(assist ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
-summary(model_assist)
-model_register <- lm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + rural + year_factor, data = reg_data)
-summary(model_register)
-model_reason_t <- (lm(work_reason ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data))
-summary(model_reason_t)
-
-```
-
-```{r generate regression table, echo = F, results = "asis"}
-stargazer::stargazer(list(model_assist, model_register), 
-               title = "Full sample regression results", 
-               label = "fullResults",
-               header = F,
-               keep = c("floods", "droughts"))
-```
-```{r probit regression total sample}
-
-probitMod_assist <- glm(assist ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data, family=binomial(link="probit"))
-
-probitMod_register <- glm(register ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data, family=binomial(link="probit"))
-
-probitMod_work <- glm(work_reason ~ floods + droughts + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data, family=binomial(link="probit"))
-
-summary(probitMod_assist)  # model summary
-summary(probitMod_register)
-summary(probitMod_work)
-```
-
-```{r generate probit regression table, echo = F, results = "asis"}
-stargazer::stargazer(list(probitMod_assist, probitMod_register, probitMod_work), 
-               title = "Full sample probit regression results", 
-               label = "fullResults",
-               header = F,
-               keep = c("floods", "droughts"))
-
-```
-
-```{r regression rural sample}
-
-rural <- subset(reg_data, rural == 1, select = c("hh_id", "id", "age", "female", "literacy", "dept", "rural", "income", "assist", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason"))
-
-model_assist_r <- (lm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = rural))
-summary(model_assist_r)
-model_register_r <- lm(register ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = rural)
-summary(model_register_r)
-model_reason <- (lm(work_reason ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = rural))
-summary(model_reason)
-
-```
-
-```{r regression urban sample}
-
-urban <- subset(reg_data, rural == 0, select = c("hh_id", "id", "age", "female", "literacy", "dept", "rural", "income", "assist", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor"))
-
-model_assist_u <- lm(assist ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data)
-summary(model_assist_u)
-model_register_u <- lm(register ~ floods + droughts + age + female + income + child + grandchild + num_children + hhhead_educ + female_hhhead + year_factor, data = reg_data)
-summary(model_register_u)
-```
-
 # Discussion of Data

### 2020-05-10 Garcia-Israel_final_seminar_paper.Rmd
Final seminar paper code and text

--- 2020-05-10 final_seminar_paper.Rmd
@@
 This study explores the relationship between flood and drought occurrences and school attendance in Bolivia.
-Contrary to my hypothesis, I find that floods and droughts actually have positive impacts on school outcomes, a relationship that might indicate that weather events negatively impact agricultural harvests and decrease the demand for children’s labor, thereby allowing them to go to school."
+Contrary to my hypothesis, I find that floods and droughts actually have positive impacts on school outcomes, a relationship that might indicate that weather events negatively impact agricultural harvests and decrease the demand for children’s labor, thereby allowing them to go to school.
+The impact of droughts on school attendance is negative, indicating that the hypothesis holds for slow-onset weather events in rural areas."
 header-includes:
@@
 ```
+
 # Introduction
@@
 A key reason for choosing Bolivia as a case study for this research question is its extensive climactic diversity.
+The country’s borders include parts of the Andes mountains and the Amazon rainforest, as well as grasslands and deciduous forests (Seiler et al 2013).
 Despite this high level of internal precipitation and temperature variation, the ecological zones correspond fairly well to political boundaries of the country, which is divided into nine departments.
-Three departments fall into each ecological category, which are mountains (La Paz, Potosi, and Oruro), valleys (Cochabamba, Tarija, Chuquisaca), and plains (Santa Cruz, Beni, Pando).
-Figure 1 shows the three regional climate divisions and the departmental boundaries, with the mountains in the west, the plains in the north and east, and the valleys in the south-central part of the country.
+Three departments fall into each ecological category: mountains (La Paz, Potosi, and Oruro), valleys (Cochabamba, Tarija, Chuquisaca), and plains (Santa Cruz, Beni, Pando).
+Figure A1 in the appendix shows the three regional climate divisions and the departmental boundaries, with the mountains in the west, the plains in the north and east, and the valleys in the south-central part of the country.
+
+While floods tend to occur in the lowland plains, they are also possible in the alpine regions.
+Floods and droughts also generally occur in specific seasons far apart from one another: January-March is the wet season for most of the country while June-August has the most drought occurrences.
+With increasing climate change, Bolivia is predicted to become warmer and experience more droughts overall (Seiler et al 2013).
+
+Over the study period of 2005-2012, the percentage of Bolivian men involved in agricultural activities ranged from 37 to 28 percent, making Bolivia an interesting country to study, as it has a mix of rural areas with subsistence and other agriculture as well as a rapid urbanization trend (ILO Estimates).
+
+This paper is structured as follows: the next section contains a review of literature, section III explains the data and methodology, section IV describes the results, section V is the discussion, and section VI concludes.
 
@@
 ## Data
+
 To answer this question, I used a combination of household survey data and flood and drought occurrences data.
@@
 The intermediary years (2008, 2009, 2010) were not included because the relevant education questions were not part of the 2008 and 2009 surveys, and there was no household survey taken in 2010, the census year.
-The flood and drought datasets show the number of floods and droughts that occurred per year in each of the nine departments of Bolivia.
+The flood and drought datasets show the number of floods and droughts that occurred per year in each department of Bolivia.
 The final compiled dataset is a repeated cross-section as different households were randomly surveyed each year.
@@
 
-##Variable Descriptions
-
 \begin{table}
@@
 The main econometric model used in the analysis is a probit regression model, as my three dependent variables are all binary variables with values of 0 or 1.
-Floods and droughts are the two main explanatory variables, along with the interaction terms floods_lny and droughts_lny, which capture the differing effect that weather shocks have on schooling outcomes for children of different income levels.
-These interaction terms were formed by multiplying the flood and drought variables with the lny, logged per capita monthly income, variable.
+*Floods* and *droughts* are the two main explanatory variables, along with the interaction terms *floods_lny* and *droughts_lny*, which capture the differing effect that weather shocks have on schooling outcomes for children of different income levels.
+These interaction terms were formed by multiplying the *flood* and *drought* variables with the *lny*, logged per capita monthly income, variable.
 Several control variables are inserted into the regression to control for income, age and sex of the child, sex and education level of the household head, number of other children in the household and whether the child is the child or grandchild of the household head.
@@
 Variables such as floods and droughts are expected to have negative coefficients, as I hypothesize that experiencing weather shocks will decrease the likelihood going to school.
-The age variable and the secondary school variable are both expected to have a negative coefficient, as older children would be more likely to be able to work and have a higher opportunity cost of going to school.
+The *age* variable and the secondary school variable are both expected to have a negative coefficient, as older children would be more likely to be able to work and have a higher opportunity cost of going to school.
 A higher number of children in the household is expected to cause a decrease in the probability of attending school.
@@
 
-The third regression uses the sample of children that do not attend school, and the dependent variable is work_reason, which shows the likelihood of work being the reason why the child is not attending school.
+The third regression uses the sample of children that do not attend school, and the dependent variable is *work_reason*, which shows the likelihood of work being the reason why the child is not attending school.
 For the omitted condition of 0, those children listed reasons other than work as their motivation for not attending school.
@@
 These three regression equations are repeated in the analysis using different samples, the total number of households and the sample that live in rural areas.
-In the lagged variable analysis, the same model is used, but floods and droughts are replaced by lag_floods and lag_droughts, which represent the number of floods or droughts in each department during the previous year.
+In the lagged variable analysis, the same model is used, but *floods* and *droughts* are replaced by *lag_floods* and *lag_droughts*, which represent the number of floods or droughts in each department during the previous year.
 Including the lagged variables is a way of testing whether households’ responses to weather shocks are long-term or short-term.
@@
 # Results
+
+Table 4 presents the results from the full sample, which includes both urban and rural locations.
+The first regression with the dependent variable *attend* and the second, with dependent variable *register*, have similar results.
+The flood variables were significant for the both, with positive coefficients, indicating that increases in floods lead to small increases in the probability of children attending or registering for school.
+The interaction terms between floods and income were also highly significant, and the coefficients were negative, suggesting that the slope of the relationship between floods and school attendance becomes flatter as household income increases.
+Thus, floods do not have as large of an impact on richer families’ school attendance.
+The coefficient on *floods* had a larger magnitude in the regression testing school attendance than in the regression for registration, which could point to floods having short-term rather than long-term effects on behavior.
+Similar results for floods occurred for the rural sample, as seen in Table 5.
 
@@
 
+The *droughts* variable was only significant in the rural sample, but its coefficient had a negative sign, indicating that for every increase in drought occurrence, the probability of attending or registering for school decreased.
+The coefficient on the interaction term between droughts and income was also highly significant but positive, showing that higher levels of income decrease the slope of the negative relationship between droughts and school attendance, but not school registration.
+
+All of the control variables were significant in at least one of the education outcomes regressions.
+*Age* had a negative impact on school attendance but a positive effect on registration.
+The *female* variable had a negative coefficient for probability of registering at school, while the variable indicating the presence of a female household head had a positive coefficient.
+Increases in years of the household head’s education also caused a higher probability of registration and attendance.
+Being the child or the grandchild of the household head positively impacted the probability of going to school, as did being registered for secondary school as opposed to primary school.
+
+The flood and drought coefficients were not significant for the third regression, with dependent variable *work_reason*.
+However, for the full sample in Table 4, the control variables that were significant are telling: *female*, *child*, and *grandchild* had negative coefficients, while the variable indicating the number of children in the household had a positive coefficient.
+The female household head variable was highly significant and had a positive coefficient.
+These results could suggest that children in households with female heads might be working because of the absence of working-age men, and that female children were less likely to be absent from school due to work reasons.
+However, none of these results are significant in the rural sample in Table 5.
+
 ```{r probit regression rural sample, echo = F, warning = F, message = F}
@@
 
-```{r probit regression urban sample, echo = F, warning = F, message = F}
-
-urban <- subset(reg_data, rural == 0, select = c("hh_id", "id", "age", "female", "literacy", "dept_factor", "rural", "lny", "attend", "register", "floods", "droughts", "year", "child", "grandchild", "num_children", "female_hhhead", "hhhead_literacy", "hhhead_educ", "year_factor", "work_reason", "secondary", "valleys", "mountains", "plains", "dept", "extr", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov", "flood_lny", "drought_lny", "lag_flood_lny", "lag_drought_lny", "lag_f_secondary", "lag_d_secondary", "lag_floods", "lag_droughts"))
-
-probitMod_assist_u <- glm(attend ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
-
-probitMod_register_u <- glm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
-
-probitMod_work_u <- glm(work_reason ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + secondary + hhhead_educ + female_hhhead + dept_factor + year_factor, data = urban, family=binomial(link="probit"))
-
-```
-
-The following tables show the results with the lagged variables for floods and droughts.
+Tables 6 and 7 examine the effect of the lagged flood and drought variables on school outcomes, showing whether weather events that occurred one year ago have a present impact.
+The coefficient on *lag_floods* was highly significant and positive again for both the total and the rural sample, showing that areas that experienced floods in the previous year also have higher probabilities of school attendance and registration.
+The negative coefficient on the interaction term between income and floods again points to a decrease in the slope of the relationship between floods and school outcomes.
+The coefficient on lagged droughts was significant and positive only for the *register* regression in the total sample.
+In the rural sample, however, it was significant and positive for the attending school variable.
+The change in the sign of the *droughts* coefficient shows that people might respond more slowly to droughts in terms of their behavioral changes in education.
 
@@
                header = F,
-               keep = c("floods", "droughts", "lag_flood_lny", "lag_drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
+               keep = c("floods", "droughts", "lag_flood_lny", "lag_drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"),
+               keep.stat = c("n", "adj.rsq"))
 
@@
 ```
-
 ## Robustness Checks
 
-The following results show a different income specification, the inclusion of the extreme poverty variable instead of the log of income.
+As a robustness check, I change the model specification to linear to see what effect that has on the results and how influenced they are by the probit model specification.
+Tables 8 and 9 show the results for the linear regression model, with the same regressions as the probit model results: *attend*, *register*, *work_reason*.
+Again, these results had similar signs and levels of significance to the probit regression model, indicating that these results are not due to the probit model specification.
+*Floods* was significant in both the rural and total samples, showing the same pattern of higher likelihood of school attendance and registration in years with high flood occurrences.
+The interaction terms *flood_lny* and *drought_lny* were highly significant and negative, showing that floods and droughts would have a more moderate impact on households at higher income levels as opposed to poorer households who would likely see an exacerbated effect of the schooling outcome relationships.
+However, the *droughts* variable was not at all significant in the either sample with the OLS regression.
+For the total sample, the adjusted R-squared was 0.026 for the regression testing school attendance and 0.095 for the regression testing school registration.
+The adjusted R-squared was 0.027 for the *attend* regression and 0.08 for the *register* regression in the rural sample.
+These numbers are quite low, indicating that there are other factors contributing to the variation in the dependent variables that are not being captured in the model.
+The F-statistics were all highly significant at the 1% level, indicating that the inclusion of the independent variables in the model are contributing to explaining the variation in the dependent variables.
 
-```{r probit regression total sample extr, echo = F, warning = F, message = F}
+```{r linear regressions total sample, echo = F, message = F}
 
-probitMod_assistx <- glm(attend ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+model_assist_t <- lm(attend ~ floods + flood_lny + droughts + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
 
-probitMod_registerx <- glm(register ~ floods + flood_pov + droughts + drought_pov + age + female + extr + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+model_register_t <- lm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
 
-probitMod_workx <- glm(work_reason ~ floods + flood_secondary + flood_pov + droughts + drought_secondary + drought_pov + age + female + extr + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data, family=binomial(link="probit"))
+model_reason_t <- (lm(work_reason ~ floods + flood_lny + droughts + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data))
 
@@
 
-```{r generate probit regression table extr, echo = F, message = F, results = "asis"}
-stargazer::stargazer(list(probitMod_assistx, probitMod_registerx, probitMod_workx), 
-               title = "Full sample probit regression results (poverty measure)", 
-               label = "fullResults8",
-               header = F,
-               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_pov", "drought_pov","age", "female", "extr", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
+```{r linear regressions rural sample, echo = F, message = F}
 
-```
+model_ar <- lm(attend ~ floods + flood_lny + droughts + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural)
 
-```{r linear regressions total sample, echo = F, message = F}
-
-model_assist_t <- lm(attend ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
-
-model_register_t <- lm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data)
-
-model_reason_t <- (lm(work_reason ~ floods + flood_secondary + flood_lny + droughts + drought_secondary + drought_lny + age + female + lny + secondary + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = reg_data))
+model_rr <- lm(register ~ floods + flood_lny + droughts + drought_lny + age + female + lny + child + grandchild + num_children + hhhead_educ + female_hhhead + dept_factor + year_factor, data = rural)
 
@@
 ```{r generate regression table, echo = F, message = F, results = "asis"}
-stargazer::stargazer(list(model_assist_t, model_register_t, model_reason_t), 
+stargazer::stargazer(list(model_assist_t, model_register_t), 
                title = "Full sample linear regression results", 
@@
                header = F,
-               keep = c("floods", "droughts", "flood_secondary", "drought_secondary", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead"))
+               keep = c("floods", "droughts", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead")) 
 ```
 
-#Discussion
+```{r generate regression table rural, echo = F, message = F, results = "asis"}
+stargazer::stargazer(list(model_ar, model_rr), 
+               title = "Rural sample linear regression results", 
+               label = "fullResults12",
+               header = F,
+               keep = c("floods", "droughts", "flood_lny", "drought_lny","age", "female", "lny", "secondary", "child", "grandchild", "num_children", "hhhead_educ", "female_hhhead")) 
+```
+
+# Discussion
 
@@
 If children’s labor is no longer required if crops are destroyed, the opportunity cost of going to school lessens considerably and they may become more likely to register and attend.
-The significance and direction of the coefficient on the interaction terms linking floods with droughts in the main probit regression also lend credence to that theory.
+The significance and direction of the coefficient on the interaction terms linking floods with income in the main probit regression also lend credence to that theory.
 
@@
 This issue is addressed by the inclusion of the lagged effects variables in Tables 6 and 7.
-These differing results could also be due to data limitations.
+For the flood variables, the lagged sample showed similar results as the current year sample, suggesting a positive relationship again between *floods* and the *register* and *attend* variables.
+The story with the droughts is more complex, especially in the rural areas.
+The results for the rural sample without the lag on *droughts* indicate that when a department is faced with high occurrences of drought in a given year, children will be less lightly to attend school.
+With the lag, the relationship becomes positive.
+This could point to a cumulative effect of having droughts last for a longer time.
+In that case, it could make sense that prolonged weather shocks might decrease the opportunity cost of going to school, but in the first year of a drought, agricultural crops might not be as affected.
+
+The results from the interaction terms were especially interesting as they were highly significant and pointed to an exacerbated effect of weather shocks on school outcomes when income levels were lower.
+This follows the idea that poorer households will be more likely to have children engaged in agricultural activities, who would then be affected the most by weather shocks.
+In the case of floods, children at lower income levels would be more likely to go to school, while in the case of droughts, children in rural areas at lower income levels would be more likely to be absent from school.
+
+The *work_reason* regression results were not statistically significant, probably due to the very small number of observations.
+As the main regression specification was a probit model, more observations are needed for that specific subsample, since probit models run best with higher amounts of observations to determine likelihood results.
+Yet the significance of the control variables suggests a story where children in households with many children and with female heads of household end up having a higher likelihood of missing school for work-related reasons.
+This interpretation follows the hypothesis of higher numbers of children putting a strain on income and raising the opportunity cost of schooling.
+
+In the robustness checks section, it was clear that the OLS results corresponded fairly well to the main probit regression results, as the significance, sign, and magnitude of the coefficients were comparable for all but the *droughts* variable in the rural sample.
+The significance of the F-statistic showed that the inclusion of the explanatory variables to the model improved the explanation of variation in the dependent variables.
+However, the adjusted R-squared statistics were low for all the samples, pointing to missing information in the model that could not be captured by this data.
+The adjusted R-squared values were slightly higher for the regressions with dependent variable *register*.
+
+On that note, the differing results from the hypothesis could also be due to data limitations.
 One main limitation of this dataset is the fact that the flood and drought occurrences are measured at department level.
@@
 
+# Conclusion
+
+In sum, the question of the relationship between negative weather events and schooling outcomes is difficult to answer, due to constraints of data as well as the potential for conflicting stories and mechanisms that are applicable to certain groups or locations but not others.
+The case of Bolivia is especially interesting due to the diverse climates within the country and even within departments, but this variation can sometimes be detrimental to the analysis when the data is not sufficiently disaggregated.
+Although the theoretical mechanism I anticipated was that increases in floods and droughts would decrease the probability of school attendance and school registration and increase the probability of missing school due to work reasons, the significant results for floods contradicted that theory.
+
+For the rural sample, however, droughts were found to have a negative impact on school attendance, which suggests that households have different levels of resilience to different types of natural disasters.
+The threat of climate change making Bolivia more vulnerable to droughts could exacerbate this effect.
+An alternative to the theoretical framework I hypothesized is the idea that these weather shocks actually enable the children to go to school more frequently because their labor is no longer needed for agricultural activities at times when harvests are bad.
+Further exploration is needed to fully isolate these effects and determine which stories occur at which times.
+
 # References
+
 Baez, J.E., Lucchetti, I., Genoni, M. E., & Salazar, M. (2017). Gone with the storm: Rainfall shocks and household wellbeing in Guatemala. Journal of Development Studies, 53(8), 1253-1271.
 
+GeoSinager. (2016). Mapa Recurrencias de Inundaciones Gestiones 2002-2012. [CSV file]. Retrieved from http://geosinager.defensacivil.gob.bo/maps/226
+
+GeoSinager. (2016). Mapa Recurrencias de Sequias Gestiones 2002-2012. [CSV file]. Retrieved from http://geosinager.defensacivil.gob.bo/maps/225
+
 Groppo, V., and K. Kraehnert. 2017. The Impact of Extreme Weather Events on Education. Journal of Population Economics 30 (2): 433–72
+
+Instituto Nacional de Estadistica. (2005). Encuesta de Hogares (2005) [SAV file]. Retrieved from https://www.ine.gob.bo/index.php/banco/base-de-datos-sociales
+
+International Labour Organization, ILOSTAT database.
+“Employment in agriculture, male (% of male employment).
+Data retrieved in September 2019.
+https://data.worldbank.org/indicator/SL.AGR.EMPL.MA.ZS?locations=BO&view=chart
 
@@
 
+Seiler, Christian., Hutjes, Ronald W. A., & Kabat, Pavel. (2013). Climate Variability and Trends in Bolivia. Journal of Applied Meteorology and Climatology, 52, 130-146.
+
 Shah, M., & Steinberg, B.M. (2017). Drought of opportunities: Contemporaneous and long-term impacts of rainfall shocks on human capital. Journal of Political Economy, 125(2), 527-561.
